---
title: "CU detailed outputs: Flow and Temperature"
output:
  html_document:
    df_print: paged
    toc: yes
date: today
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(kableExtra)

```


```{r, comment=NA, warning=FALSE,message=FALSE, results="asis"}

theme_custom <- function(base_size = 14) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = rel(0.85), face = "bold"),
      axis.text = element_text(size = rel(0.70), face = "bold"),
      #axis.line = element_line(color = "black", arrow = arrow(length = unit(0.3, "lines"), type = "closed")),
      axis.line = element_line(color = "black"),
      # La légende
      legend.title = element_text(size = rel(0.85), face = "bold"),
      legend.text = element_text(size = rel(0.70), face = "bold"),
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", colour = NA),
      # Les étiquettes dans le cas d'un facetting
      strip.background = element_rect(fill = "#17252D", color = "#17252D"),
      strip.text = element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
    )
}

theme_set(theme_custom())
```

## Overview

The outputs below provide indicators of climate exposure for a subset of conservation units of Pacific salmon in the Fraser Basin. The indicators are based on outputs from the Thermalscapes model of August mean temperature and the PCIC gridded hydrologic model.
Outputs use a historic period of 1981-2000 for Thermalscapes and 1971-2000 for the PCIC model, and mid-century period of 2041-2060 for Thermalscapes and 2041-2070 for PCIC. For each CU and time period, the average indicator value for streams or grid cells within the CU boundary is calculated and plotted.

Output is subsetted to accessible streams only, as modelled by BC Fish Pass.

**Flow Indicators (PCIC):**\
peakQday - date of maximum peak flow in the year (day of year)\
lowQ05 - number of days per year with low flow events below 95th percentile from historic period\

**Temperature Indicators:**\
Aug_T_proj - average August mean temperatures, weighted by stream length\
prop18proj - proportion of streams with August mean temperature above 18 degrees\
Aug_T_z    - z-score of mid-century August mean temperature compared to historic period\



```{r, results='asis'}

#choose subset of CUs to create plots for
for(i in 1:n.CUs) {
  
  cat("##", cu_run$cuid[i]," - ", cu_run$cuname[i], " - ", cu_run$Species_simple[i], "\n")
  
  kable(all_CU_stats[i,]) %>%
    kable_classic()
  
  source(file.path("2e_fw_CU_plots.R"))

  cat("\n")
  cat("\n")
  
}

```