---
title: "Climate Vulnerability Indicators for Salmon:
        PRELIMINARY Freshwater Outputs"
output:
  html_document:
    df_print: paged
    toc: yes
date: today
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(kableExtra)

#set to true to make output for FAZ. If False, default of CU outputs will be used
do_FAZ <- F


```


```{r, comment=NA, warning=FALSE,message=FALSE, results="asis"}

theme_custom <- function(base_size = 14) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = rel(0.85), face = "bold"),
      axis.text = element_text(size = rel(0.70), face = "bold"),
      #axis.line = element_line(color = "black", arrow = arrow(length = unit(0.3, "lines"), type = "closed")),
      axis.line = element_line(color = "black"),
      # La légende
      legend.title = element_text(size = rel(0.85), face = "bold"),
      legend.text = element_text(size = rel(0.70), face = "bold"),
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", colour = NA),
      # Les étiquettes dans le cas d'un facetting
      strip.background = element_rect(fill = "#17252D", color = "#17252D"),
      strip.text = element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
    )
}

theme_set(theme_custom())
```

## Overview

This document contains preliminary outputs of freshwater indicators for conservation units (CUs) of Pacific Salmon in the Fraser Basin. Outputs currently show comparisons between a historic and mid-century time period using ensemble model predictions.

Models used in the analysis include:  

** Thermalscapes **: A statistical downscaled model of August mean temperature for streams in British Columbia [https://www.tandfonline.com/doi/abs/10.1080/07011784.2023.2267028].

** 7DEC Stream Temperature Model **: A model using similar inputs as Thermalscapes to predict the highest weekly 7 day equivalent constant temperature of streams in BC. [https://onlinelibrary.wiley.com/doi/10.1111/fwb.14300]

** PCIC Variable Infiltration Capacity with Glacier (VIC-GL) Gridded Model **: A gridded hydrologic model that simulates daily streamflow and temperature for the Fraser Basin in 30km2 grid cells.

** Cumulative Threat Score **: A model that combines multiple stressors to calculate a cumulative threat score for streams in the Fraser Basin.

** BC FishPass **: A model that predicts the accessibility of streams to salmonids based on fish observations and barriers to fish passage. [https://smnorris.github.io/bcfishpass/]

Most plots are subsetted to accessible streams only, as modelled by BC Fish Pass.



```{r, results='asis'}

if(do_FAZ == TRUE) n.iter <- n.FAZ else n.iter <- n.CUs

#choose subset of CUs to create plots for
for(i in 1:n.CUs) {
    
  # p_sp <- cu_run$Species_simple[i]
  # if(i ==1) cat("## ", p_sp, "\n")
  # if(i >1)
  #   if(p_sp != cu_run$Species_simple[i-1]) cat("## ", p_sp, "\n")

  if(do_FAZ == FALSE) cat("### ", cu_run$cuid[i]," - ", cu_run$cuname[i], " - ", cu_run$Species_simple[i], "\n")
  else if (do_FAZ == TRUE) cat("### ", FAZ_Fr$FAZ_Name[i], "\n")
  
  # Subset CU or FAZ boundary
  if(do_FAZ == FALSE) {
  cu_boundary_i <- cu_boundary[cu_boundary$CUID == cuid[i], ]
  FWA_cu <- st_crop(FWA_Fr_high, cu_boundary_i)
  # Subset CU migration path
  path_CU <- path_list[[i]]
  
  #subset nuseds observations
  nuseds_CU <- nuseds_Fr[nuseds_Fr$FULL_CU_IN == cu_run$FULL_CU_IN[i],]
  
  fw_indies_CU <- fw_acc_indies[stream_cu_picks[,i],] #%>%
    #filter(!is.na(Tw8_0_00_0))
  } else if (do_FAZ == TRUE) {
  cu_boundary_i <- FAZ_Fr[i,]
  FWA_cu <- st_crop(FWA_Fr_high, cu_boundary_i)
  
  nuseds_CU <- nuseds_Fr
  fw_indies_CU <- fw_acc_indies[stream_FAZ_picks[,i],]
  }
  
  PCIC_Aug_CU_hist <- st_crop(PCIC_month, fw_indies_CU) %>%
    filter(month(time) == 8, year(time) == 1985) 
  PCIC_Aug_CU_proj <- st_crop(PCIC_month, fw_indies_CU) %>%
    filter(month(time) == 8, year(time) == 2055) 
  
  PCIC_CU_diff <- PCIC_Aug_CU_proj %>%
    st_set_dimensions(3, values = as.POSIXct("1985-08-15", "UTC"), names = "time")
  
  PCIC_CU_diff <- PCIC_CU_diff - PCIC_Aug_CU_hist
  PCIC_CU_prop <- PCIC_CU_diff / PCIC_Aug_CU_hist
  
  PCIC_day_CU <-  st_crop(PCIC_day, cu_boundary_i) %>% 
    as_tibble()
  
  PCIC_day_summary <- PCIC_day_CU %>%
    group_by(time) %>%
    dplyr::summarize(tw_mean = mean(tw_day, na.rm=T),
                     tw_q10 = quantile(tw_day, c(0.10), na.rm = T),
                     tw_q90 = quantile(tw_day, c(0.90), na.rm = T),
                     flow_mean = mean(flow_day, na.rm = T),
                     flow_q10 = quantile(flow_day, c(0.10), na.rm = T),
                     flow_q90 = quantile(flow_day, c(0.90), na.rm = T))%>%
    mutate(year = as.factor(year(time)), 
           month = month(time),
           day = yday(time))
  
  # if(p_sp == "Chinook") p_sp_pick <- 1  #choose which species to plot observation points
  # if(p_sp == "Coho") p_sp_pick <- 2 
  # if(p_sp == "Sockeye") p_sp_pick <- 3
  # if(p_sp == "Pink" | p_sp == "Chum") p_sp_pick <- 0
  cat("#### ", "CU Boundary Outputs \n")
  
  if(do_FAZ == FALSE) cu_boundary_table <- all_CU_stats[i,]
  else cu_boundary_table <- all_FAZ_stats[i,]
  
  cu_boundary_table <- cu_boundary_table %>%
    mutate(length_sum = length_sum/1000) %>%
    select(n_streams, length_sum, ThiPI_9_45_3, Tw_rate, MADprop_8_diff,
           MADprop_win_diff, peakQday_diff, CT_anad_mean) %>%
    rename("Number of stream segments" = n_streams,
           "Total stream length (km)" = length_sum,
           "7DEC High T 2041-2060" = ThiPI_9_45_3,
           "Temperature Rate of Change (deg C/10y)" = Tw_rate,
           "Change in August flow as %MAD" = MADprop_8_diff,
           "Change in winter flow as %MAD" = MADprop_win_diff,
           "Change in peak flow day"       = peakQday_diff,
           "Cumulative threat score"       = CT_anad_mean
           )
  
  
  # ts_gcms <- fw_indies_CU %>%
  #   pivot_longer(cols = c(Tw8_9_45_3, Tw8_1_45_3, Tw8_2_45_3, Tw8_3_45_3, Tw8_4_45_3, Tw8_5_45_3, Tw8_6_45_3),
  #                 names_to = "model",
  #                 values_to = "value")
  # 
  # ts_models <- fw_indies_CU %>%
  #   pivot_longer(cols = c(Tw8_9_45_3, Tav_9_45_3, Risk20_mod_len),
  #                 names_to = "model",
  #                 values_to = "value")
  
  
  #gt(cu_boundary_table) + ggtitle("CU Boundary Statistics")
  print(kable(cu_boundary_table))
  
  cu_acc_p <- ggplot() +
    geom_sf(data = st_zm(FWA_cu), color = "grey", alpha = 0.6) + 
    geom_sf(data = fw_indies_CU, aes(color = model_access_salmon)) +
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
    geom_sf(data = nuseds_CU, aes(color = SPECIES), alpha = 0.6) +
    # {switch(p_sp_pick, geom_sf(data = chin_sp, color = "darkgreen", alpha = 0.6),
    #         geom_sf(data = coho_sp, color = "darkblue", alpha = 0.6),
    #         geom_sf(data = sockeye_sp, color = "darkred", alpha = 0.6))} +
    coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
             ylim = st_bbox(cu_boundary_i)[c(2,4)]) +
    labs(subtitle = paste("CU ", cu_run$cuname[i]), colour = "BC FishPass Accessibility") +
    theme_minimal()
  
  ct_anad_p <- ggplot() +
    geom_sf(data = fw_indies_CU, aes(color = CT_anad), linewidth = 1.) +
    scale_color_viridis(limits = c(0,5.5)) + 
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
    coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
             ylim = st_bbox(cu_boundary_i)[c(2,4)]) +
    labs(subtitle = paste("Cumulative stressor score (anadromous)"), color = "Threat Score") +
    theme_minimal()
  
  ts_hist_p <- ggplot() +
    geom_sf(data = fw_indies_CU, aes(color = Tw8_0_00_0), linewidth = 1.) +
    scale_color_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                          limits = c(5,25)) + 
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
    coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
             ylim = st_bbox(cu_boundary_i)[c(2,4)]) +
    labs(subtitle = paste("August T, Thermalscapes, 1981-2000"), color = "T (deg C)") +
    theme_minimal()
  
  ts_proj_p <- ggplot() +
    #geom_sf(data = st_zm(FWA_cu), color = "grey", alpha = 0.4) + 
    #geom_stars(data = PCIC_Aug_CU_proj["tw_month"]) +
    #scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) + 
    geom_sf(data = fw_indies_CU, aes(color = Tav_9_45_3), linewidth = 1.) +
    scale_color_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                          limits = c(5,25)) + 
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
    coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
             ylim = st_bbox(cu_boundary_i)[c(2,4)]) +
    labs(subtitle = paste("Mean 7DEC, 2041-2060"), color = "T (deg C)") +
    theme_minimal()
  
  # ggplot() + 
  #   geom_histogram(data = fw_indies_CU, aes(x = ThiPI_9_45_3), bins = 30, colour = "green")
  #   #geom_histogram(data = fw_indies_CU, aes(x = Tw8_9_45_3), bins = 30, colour = "blue") 
  
  max_hist_flow <- max(fw_indies_CU$mean_flow_m3s_17_1, na.rm = T)
  max_proj_flow <- max(fw_indies_CU$mean_17_40, na.rm = T)
  max_plot_flow <- max(max_hist_flow, max_proj_flow)
  
  flow_hist_p <- ggplot() +
    geom_sf(data = fw_indies_CU, aes(color = mean_flow_m3s_17_1), linewidth = 1.) +
    scale_color_viridis(limits = c(0, max_plot_flow)) + 
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
    coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
             ylim = st_bbox(cu_boundary_i)[c(2,4)]) +
    labs(subtitle = paste("Mean annual discharge, 1981-2000"), color = "Flow (m3/s)") +
    theme_minimal()
  
  flow_proj_p <- ggplot() +
    geom_sf(data = fw_indies_CU, aes(color = mean_17_40), linewidth = 1.) +
    scale_color_viridis(limits = c(0, max_plot_flow)) + 
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
    coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
             ylim = st_bbox(cu_boundary_i)[c(2,4)]) +
    labs(subtitle = paste("Mean annual discharge, 2041-2060"), color = "Flow (m3/s)") +
    theme_minimal()
  
  tPCIC_hist_p <- ggplot() +
    geom_stars(data = PCIC_Aug_CU_hist["tw_month"]) +
    scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                         limits = c(5,25)) + 
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
    #geom_sf(data = st_zm(FWA_cu), color = "black", alpha = 0.6) + 
    coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
             ylim = st_bbox(cu_boundary_i)[c(2,4)]) + 
    labs(subtitle = "1970-2000 August stream temperature", fill = "T (deg C)")
  
  tPCIC_proj_p <- ggplot() +
    geom_stars(data = PCIC_Aug_CU_proj["tw_month"]) +
    scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                         limits = c(5,25)) + 
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
    #geom_sf(data = st_zm(FWA_cu), color = "black", alpha = 0.6) + 
    coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
             ylim = st_bbox(cu_boundary_i)[c(2,4)]) + 
    labs(subtitle = "2041-2070 August stream temperature", fill = "T (deg C)")
  
  # temp <- PCIC_Aug_CU_hist$tw_month %>%
  #   as.vector() %>%
  #   na.omit()
  # ggplot() +
  #   geom_histogram(aes(x = temp), bins = 30, colour = "red")
  
  max_hist_flow <- max(PCIC_Aug_CU_hist$flow_month, na.rm = T)
  max_proj_flow <- max(PCIC_Aug_CU_proj$flow_month, na.rm = T)
  max_plot_flow <- max(max_hist_flow, max_proj_flow)
  
  qPCIC_hist_p <- ggplot() + 
    geom_stars(data = PCIC_Aug_CU_hist["flow_month"]) +
    scale_fill_viridis(limits = c(0, max_plot_flow)) +
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
    #geom_sf(data = st_zm(FWA_cu), color = "black", alpha = 0.6) + 
    coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
             ylim = st_bbox(cu_boundary_i)[c(2,4)]) + 
    labs(subtitle = "August flow 1970-2000", fill = "Flow (m3/s)")
  
  qPCIC_proj_p <- ggplot() + 
    geom_stars(data = PCIC_Aug_CU_proj["flow_month"]) +
    scale_fill_viridis(limits = c(0, max_plot_flow)) + 
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
    #geom_sf(data = st_zm(FWA_cu), color = "black", alpha = 0.6) + 
    coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
             ylim = st_bbox(cu_boundary_i)[c(2,4)]) + 
    labs(subtitle = "August flow 2041-2070", fill = "Flow (m3/s)")
  
  
  ###plotting of Thermalscapes values across GCMs
  # ts_stream_p <- ggplot() +
  #   geom_violin(data = ts_gcms, aes(x = model, y = value, fill = model), draw_quantiles = c(0.5)) +
  #   scale_fill_discrete(labels=c('CanESM2', 'CSIRO-Mk3-6-0', 'GFDL-CM3', 
  #                                'HadGEM2-ES', 'MIROC-ESM', 'MPI-ESM-LR', 'ensemble mean')) +
  #   labs(fill = "Global Climate Model", subtitle = "Thermalscapes August mean temperature 2041-2060")
  
  # ct_stream_p <- ggplot() +
  #   geom_histogram(data = fw_indies_CU, aes(x = CT_anad))
  
  
  timing_s <- cu_timing$sp_start[cu_timing$cuid == cuid[i]]
  timing_e <- cu_timing$sp_end[cu_timing$cuid == cuid[i]]
  
  
  ### plotting of hydrological regime
  hydro_day_plot <- ggplot(data = PCIC_day_summary, aes(x = as.Date("2019-12-31") + yday(time), y = flow_mean, ymin = flow_q10, ymax = flow_q90, colour = year)) +
    geom_point() +
    geom_line() +
    geom_ribbon(alpha = 0.1) +
    geom_vline(xintercept = as.Date("2019-12-31") + timing_s, linetype = "dashed") +
    geom_vline(xintercept = as.Date("2019-12-31") + timing_e, linetype = "dashed") +
    geom_text(aes(x = as.Date("2019-12-31") + timing_e, y = max(flow_mean), label = "Spawning period"), size = 3) +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    scale_colour_discrete(labels=c('1970-2000', '2041-2070')) +
    labs(subtitle = "PCIC daily flow within CU boundary accessible cells (80% range)", colour = "Year range",
         x = "Date", y = "Daily Flow")
  
  tw_day_plot <- ggplot(data = PCIC_day_summary, aes(x = as.Date("2019-12-31") + yday(time), y = tw_mean, ymin = tw_q10, ymax = tw_q90, colour = year)) +
    geom_point() +
    geom_line() +
    geom_ribbon(alpha = 0.1) +
    geom_vline(xintercept = as.Date("2019-12-31") + timing_s, linetype = "dashed") +
    geom_vline(xintercept = as.Date("2019-12-31") + timing_e, linetype = "dashed") +
    geom_text(aes(x = as.Date("2019-12-31") + timing_e, y = max(tw_mean), label = "Spawning period"), size = 3) +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    scale_colour_discrete(labels=c('1970-2000', '2041-2070')) +
    labs(subtitle = "PCIC daily temperature within CU boundary accessible cells (80% range)", colour = "Year range",
         x = "Date", y = "Daily Temperature")
  
  #printcu_acc_p, nrow = 1, ncol = 1)
  #grid.arrange(ts_hist_p, ts_proj_p, nrow = 1, ncol = 2)
  print(cu_acc_p)
  print(ct_anad_p)
  print(ts_hist_p)
  print(ts_proj_p)
  print(flow_hist_p)
  print(flow_proj_p)
  #print(ts_stream_p)
  
  
  #grid.arrange(tPCIC_hist_p, tPCIC_proj_p, nrow =1, ncol =2)
  #grid.arrange(qPCIC_hist_p, qPCIC_proj_p, nrow =1, ncol =2)
  grid.arrange(hydro_day_plot, tw_day_plot, nrow = 2, ncol = 1)
  
  
  
  if(do_FAZ == FALSE) {
  
  cat("\n\n")
  cat("#### ", "CU Migration Path Outputs \n")
  
  
  cu_upstream_table <- all_CU_stats[i,] %>%
    select(timing_start, timing_end,  migr_dist, Tmigr_hist_mean, Tmigr_proj_mean, Qmigr_hist_mean, Qmigr_proj_mean, Qmigr_propdiff) %>%
    rename("River entry timing start (2.5%)" = timing_start,
           "River entry timing end (97.5%)" = timing_end,
           "Migration distance (km)" = migr_dist,
           "Mean T 1971-2000" = Tmigr_hist_mean,
           "Mean T 2041-2070 (T_mig)" = Tmigr_proj_mean,
           "Mean flow 1971-2000" = Qmigr_hist_mean,
           "Mean flow 2041-2070" = Qmigr_proj_mean,
           "% change in flow (Q_mig)" = Qmigr_propdiff)
  
  #gt(cu_boundary_table) + ggtitle("CU Boundary Statistics")
  print(kable(cu_upstream_table))
  
  ### Upstream Migration Plots
  timing_s <- cu_timing$rt_start[cu_timing$cuid == cuid[i]]
  timing_e <- cu_timing$rt_end[cu_timing$cuid == cuid[i]]
  
  PCIC_migr_hist_CU <- st_crop(PCIC_day, path_CU) %>%
    filter(yday(time) >= timing_s, yday(time) <= timing_e, year(time) == 1985) %>%
    aggregate(by = "year", FUN = mean)
  
  PCIC_migr_proj_CU <- PCIC_day[path_CU] %>%
    filter(yday(time) >= timing_s, yday(time) <= timing_e, year(time) == 2055) %>%
    aggregate(by = "year", FUN = mean)
  
  migr_hist_T_plot <- ggplot() +
    geom_stars(data = PCIC_migr_hist_CU["tw_day"]) +
    scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                         limits = c(5,25)) + 
    theme_minimal() +
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.2) +
    geom_sf(data = st_zm(path_CU)) + 
    labs(subtitle = paste("PCIC temp 1970-2000" ,"days",timing_s, "-", timing_e))
  
  migr_proj_T_plot <- ggplot() +
    geom_stars(data = PCIC_migr_proj_CU["tw_day"]) +
    scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                         limits = c(5,25)) + 
    theme_minimal() +
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.2) +
    geom_sf(data = st_zm(path_CU)) + 
    labs(subtitle = paste("PCIC temp 2041-2070" ,"days",timing_s, "-", timing_e))
  
  max_hist_flow <- max(PCIC_migr_hist_CU$flow_day, na.rm = T)
  max_proj_flow <- max(PCIC_migr_proj_CU$flow_day, na.rm = T)
  max_plot_flow <- max(max_hist_flow, max_proj_flow)
  
  migr_hist_Q_plot <- ggplot() +
    geom_stars(data = PCIC_migr_hist_CU["flow_day"]) +
    scale_fill_viridis(limits = c(0, max_plot_flow)) + 
    theme_minimal() +
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.2) +
    geom_sf(data = st_zm(path_CU)) + 
    labs(subtitle = paste("PCIC flow 1970-2000:" ,"days",timing_s, "-", timing_e))
  
  migr_proj_Q_plot <- ggplot() +
    geom_stars(data = PCIC_migr_proj_CU["flow_day"]) +
    scale_fill_viridis(limits = c(0, max_plot_flow)) + 
    theme_minimal() +
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.2) +
    geom_sf(data = st_zm(path_CU)) + 
    labs(subtitle = paste("PCIC flow 2041-2070:" ,"days",timing_s, "-", timing_e))
  
  grid.arrange(migr_hist_T_plot, migr_proj_T_plot, nrow =1, ncol =2)
  grid.arrange(migr_hist_Q_plot, migr_proj_Q_plot, nrow =1, ncol =2)

  }
  
  cat("\n")
  cat("\n")
  
}

```
