---
title: "CU Profile"
output:
  html_document:
    toc: true
    number-sections: true
params:
  FULL_CU_IN: "CK-3"

---

```{r custom theme, echo = FALSE, message = F, warning = F}

knitr::opts_chunk$set(message = F, echo = F, warning = F)

theme_cvis <- function(base_size = 14) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = rel(0.85), face = "bold"),
      axis.text = element_text(size = rel(0.70), face = "bold"),
      axis.line = element_line(color = "black"),
      # La légende
      legend.title = element_text(size = rel(0.85), face = "bold"),
      legend.text = element_text(size = rel(0.70), face = "bold"),
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", colour = NA),
      # Les étiquettes dans le cas d'un facetting
      strip.background = element_rect(fill = "#17252D", color = "#17252D"),
      strip.text = element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
    )
}
theme_set(theme_cvis())
```

```{r subset CU}

cu_run_i <- cu_run[cu_run$FULL_CU_IN == params$FULL_CU_IN,]

cuid_i <- cu_run_i$cuid
sp_pick <- cu_run_i$Species_simple
sp_pick_abb <- cu_run_i$spp

```


# Climate Vulnerability Indicators For `r cu_run$CU_NAME[i]`

This report summarizes climate vulnerability information for a conservation unit of Pacific salmon.

## Summary of Vulnerability Indicators

The following tables show the raw (untransformed) and standardized indicator values. Standardized values are between 0 and 1, with 1 indicating a higher risk score. The mean value of indicators across all Fraser basin CUs for the species are shown for comparison of relative risk. 

```{r indicators summary, results = "asis"}

STD_all_i <- STD_all[STD_all$cuid == cuid[i],]
CVIS_all_i <- CVIS_all[CVIS_all$cuid == cuid[i],]

CVIS_STD_all_i <- CVIS_STD_all[CVIS_STD_all$cuid == cuid[i],]

STD_all_spec <- STD_all %>%
  group_by(STD_Species_simple) %>%
  summarise(across(STD_FW_SPN_EXP_projT_9:last_col(), \(x) mean(x, na.rm = T)))

CVIS_all_spec <- CVIS_all %>%
  group_by(Species_simple) %>%
  summarise(across(FW_SPN_EXP_projT_9:last_col(), \(x) mean(x, na.rm = T)))

dem_table <- names(select(CVIS_all_i, contains("DEM"))) %>%
  bind_cols(t(select(CVIS_all_i, contains("DEM")))) %>%
  bind_cols(t(select(STD_all_i, contains("DEM")))) %>%
  bind_cols(t(select(CVIS_all_spec, contains("DEM")))) %>%
  bind_cols(t(select(STD_all_spec, contains("DEM"))))
names(dem_table) <- c("Indicator", "CU_Raw", "CU_Std", "Species_Raw", "Species_Std")

spn_table <- names(select(CVIS_all_i, contains("FW_SPN"))) %>%
  bind_cols(t(select(CVIS_all_i, contains("FW_SPN")))) %>%
  bind_cols(t(select(STD_all_i, contains("FW_SPN")))) %>%
  bind_cols(t(select(CVIS_all_spec, contains("FW_SPN")))) %>%
  bind_cols(t(select(STD_all_spec, contains("FW_SPN"))))
names(spn_table) <- c("Indicator", "CU_Raw", "CU_Std", "Mean_Species_Raw", "Mean_Species_Std")
spn_table <- spn_table %>%
  mutate(above_mean = if_else(CU_Std > Mean_Species_Std, T, F))

mig_table <- names(select(CVIS_all_i, contains("FW_MIG"))) %>%
  bind_cols(t(select(CVIS_all_i, contains("FW_MIG")))) %>%
  bind_cols(t(select(STD_all_i, contains("FW_MIG")))) %>%
  bind_cols(t(select(CVIS_all_spec, contains("FW_MIG")))) %>%
  bind_cols(t(select(STD_all_spec, contains("FW_MIG"))))
names(mig_table) <- c("Indicator", "CU_Raw", "CU_Std", "Mean_Species_Raw", "Mean_Species_Std")
mig_table <- mig_table %>%
  mutate(above_mean = if_else(CU_Std > Mean_Species_Std, T, F))

dem_table %>%
  kbl(caption = "Demographic indicators", digits = 2) %>%
  kable_classic()
  
spn_table %>%
  kbl(caption = "Freshwater spawning to rearing indicators", digits = 2) %>%
  kable_classic()

colors <- c("CU" = "blue", "Species Mean" = "black", "FALSE" = "forestgreen", "TRUE" = "darkred")
ggplot(spn_table, aes(x=Indicator)) +
  geom_segment( aes(xend=Indicator, y=Mean_Species_Std, yend=CU_Std, color = above_mean), size = 2) +
  geom_point( aes(y=CU_Std, color = "CU"), size=3) +
  geom_point( aes(y=Mean_Species_Std, color = "Species Mean"), size=3) +
  coord_flip()+
  labs(x = "",
       y = "Standardized Value",
       color = "Legend") +
  scale_color_manual(values = colors) +
  ylim(0,1)


mig_table %>% 
  kbl(caption = "Freshwater upstream migration indicators", digits = 2) %>%
  kable_classic()


ggplot(mig_table, aes(x=Indicator)) +
  geom_segment( aes(xend=Indicator, y=Mean_Species_Std, yend=CU_Std, color=above_mean),  size = 2) +
  geom_point( aes(y=CU_Std, color = "CU"), size=4) +
  geom_point( aes(y=Mean_Species_Std, color = "Species Mean"), size=4) +
  coord_flip()+
  labs(x = "",
       y = "Standardized Value",
       color = "Legend") +
  scale_color_manual(values = colors) +
  ylim(0,1)

```



## Conservation Unit Demographic Overview
 
### Status, abundance and enhancement
```{r demographics summary , echo = FALSE, warning = FALSE, message = FALSE}

dem_details <- cu_run_i %>%
  select(FULL_CU_IN, CU_NAME, 
         gen_length, 
         WSP_population_status, Most_Recent_Generational_Average,
         SEP_avg_annual_releases_actual, SEP_primary_prod_objective)

dem_details %>%
  kbl(digits = 2) %>%
  kable_classic()


```


### Timing and life history

Timing by life history stage is taken from Wilson and Peacock (2025) [https://cdnsciencepub.com/doi/10.1139/cjfas-2024-0213].

Data quality is rated on a scale of 1 to 5 based on number of observations and their recentness, with 1 being the highest quality data.

```{r timing summary, echo = FALSE}

# Extract spawning and migration details
timing_i <- cu_timing_long %>%
  filter(cuid == cuid_i) %>%
  arrange(desc(life_stage))

timing_wide_i <- timing_i %>%
  pivot_wider(names_from = life_stage, values_from = c(start, peak, end, dat_qual)) %>%
  select(c(FULL_CU_IN, culabel, oe_age, peak_freshwater_migration, peak_ocean_entry, peak_run_timing, peak_spawning)) 

timing_wide_i %>%
  kbl() %>%
  kable_classic()

sp_timing_s <- round(cu_timing_Fr$sp_start[cu_timing_Fr$cuid == cuid_i])
sp_timing_e <- round(cu_timing_Fr$sp_end[cu_timing_Fr$cuid == cuid_i])
sp_timing_peak <- round(cu_timing_Fr$sp_end[cu_timing_Fr$cuid == cuid_i])
rt_s <- round(cu_timing_Fr$rt_start[cu_timing_Fr$cuid == cuid_i])
rt_e <- round(cu_timing_Fr$rt_end[cu_timing_Fr$cuid == cuid_i])
rt_peak <- round(cu_timing_Fr$rt_peak[cu_timing_Fr$cuid == cuid_i])
oe_s <- round(cu_timing_Fr$oe_start[cu_timing_Fr$cuid == cuid_i])
oe_e <- round(cu_timing_Fr$oe_end[cu_timing_Fr$cuid == cuid_i])
oe_peak <- round(cu_timing_Fr$oe_peak[cu_timing_Fr$cuid == cuid_i]) 

#ridgeline timing plot
# ggplot(timing_i, aes(x = date, y = life_stage, fill = life_stage)) +
#   geom_density_ridges() +
#   theme_ridges() + 
#   theme(legend.position = "none")

#timing plot
colors <- c("peak" = "maroon", "95th perc" = "darkblue")
timing_p <- ggplot( timing_i, aes(x=life_stage, xend=life_stage, 
                                  y=as.Date("2000-01-01") + start, 
                                  yend=as.Date("2000-01-01") + end), color="grey") +
  geom_segment(color="grey", linewidth = 4) +
  geom_point( aes(y=as.Date("2000-01-01") + start, color="95th perc"), size=3 ) +
  geom_point( aes(y=as.Date("2000-01-01") + end, color="95th perc"), size=3 ) +
  geom_point( aes(x=life_stage, y=as.Date("2000-01-01") + peak, color="peak"), size=3 ) +
  geom_text( aes(x=life_stage, y=as.Date("2000-01-01") + 350, label = paste("Data \n Quality:", dat_qual)), size=3) +
  scale_y_date(date_breaks = "1 month", date_labels = "%b", limits =   c(as.Date("2000-01-01"), as.Date("2000-12-31"))) +
  coord_flip()+
  xlab("Life Stage") +
  ylab("Date") +
  labs(subtitle = paste("Ocean Entry Age: ", timing_wide_i$oe_age),
       color = "Legend")
  
print(timing_p)

```


## Freshwater Information

### Spawning to rearing
The CU boundary is used to subset streams and migration paths for the CU.
```{r spawning to rearing, echo = FALSE, warnings = F, message = F}

cu_boundary_i <- cu_boundary[cu_boundary$CUID == cuid_i, ]
FWA_cu <- st_crop(FWA_Fr_high, cu_boundary_i)

colnames(stream_cu_picks) == cuid_i
stream_cu_sub <- stream_cu_picks[,colnames(stream_cu_picks) == cuid_i]
amod_CU <- fw_amod[stream_cu_sub,]

#subset nuseds observations
nuseds_CU <- nuseds_Fr[nuseds_Fr$FULL_CU_IN == cu_run$FULL_CU_IN[i],]

amod_CU <- amod_CU %>%
  mutate(nuseds_presence = replace_na(str_detect(nuseds_sp, sp_pick), F))

# get species-specific ENM reaches
reaches_ENM_sp <- get(paste0("reaches_ENM_", sp_pick_abb))
pick_st <- lengths(st_intersects(reaches_ENM_sp, cu_boundary_i)) > 0
reaches_ENM_cu <- reaches_ENM_sp[pick_st,] %>%
  mutate(Fav_diff = Fav_f.9_45_3 - Fav_f.0_00_1)

spn_stats_CU <- all_spn_stats[all_spn_stats$cuid == cuid_i,] %>%
  select(c(CU_NAME, FULL_CU_IN, cuid, n_streams, length_sum, avg_length, avg_order, cu_area))

spn_stats_CU %>%
  kbl() %>%
  kable_classic()


PCIC_day_CU <-  st_crop(PCIC_day, cu_boundary_i) %>% 
  as_tibble()

PCIC_day_summary <- PCIC_day_CU %>%
  group_by(time) %>%
  dplyr::summarize(waterTemperature_mean = mean(waterTemperature, na.rm=T),
                   waterTemperature_q10 = quantile(waterTemperature, c(0.10), na.rm = T),
                   waterTemperature_q90 = quantile(waterTemperature, c(0.90), na.rm = T),
                   discharge_mean = mean(discharge, na.rm = T),
                   discharge_q10 = quantile(discharge, c(0.10), na.rm = T),
                   discharge_q90 = quantile(discharge, c(0.90), na.rm = T))%>%
  mutate(year = as.factor(year(time)), 
         month = month(time),
         day = yday(time))


## plots
order_p <- ggplot(amod_CU) +
  geom_histogram(aes(x = STREAM_ORDER, fill = nuseds_presence)) +
  labs(fill = "NuSEDS site")

cu_acc_p <- ggplot() +
    geom_sf(data = st_zm(FWA_cu), color = "grey", alpha = 0.6) + 
    geom_sf(data = amod_CU, aes(color = model_access_salmon)) +
    geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
    geom_sf(data = nuseds_CU, aes(color = SPECIES), alpha = 0.6) +
    coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
             ylim = st_bbox(cu_boundary_i)[c(2,4)]) +
    labs(subtitle = paste(cu_run_i$CU_NAME), colour = "BC FishPass")


  ### plotting of hydroloy
  hydro_day_plot <- ggplot(data = PCIC_day_summary, aes(x = as.Date(paste0(hist_ystart, "-01-01")) + yday(time),
                                                       y = discharge_mean, ymin = discharge_q10, ymax = discharge_q90,
                                                       colour = year)) +
    geom_ribbon(alpha = 0.1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_s, 
               linetype = "dashed", colour = "forestgreen", size = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_e, 
               linetype = "dashed", colour = "forestgreen", size = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_peak, 
               linetype = "dotted", colour = "forestgreen", size = 1) +
    geom_text(aes(x = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_peak, y = 0), 
              label = "Spawning", colour = "forestgreen", size = 3) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_s, linetype = "dashed", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_e, linetype = "dashed", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_peak, linetype = "dotted", linewidth = 1) +
    geom_text(aes(x = as.Date(paste0(hist_ystart, "-01-01")) + rt_peak, y = 0), 
              label = "Run timing", colour = "black", size = 3) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + oe_s, 
               linetype = "dashed", colour = "goldenrod", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + oe_e, 
               linetype = "dashed", colour = "goldenrod", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + oe_peak, 
               linetype = "dotted", colour = "goldenrod", linewidth = 1) +
    geom_text(aes(x = as.Date(paste0(hist_ystart, "-01-01")) + oe_peak, y = 0), 
              label = "Ocean Entry", colour = "goldenrod", size = 3) +
        geom_point() +
    geom_line() +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    scale_colour_discrete(labels=c(paste0(hist_ystart,"-",hist_ystart+20), paste0(proj_ystart,"-", proj_ystart+20))) +
    labs(subtitle = "PCIC daily flow within CU boundary accessible cells (80% range)", colour = "Year range",
         x = "Date", y = "Daily Flow")
  
  tw_day_plot <- ggplot(data = PCIC_day_summary, aes(x = as.Date(paste0(hist_ystart, "-01-01")) + yday(time), y = waterTemperature_mean, ymin = waterTemperature_q10, ymax = waterTemperature_q90, colour = year)) +
    geom_ribbon(alpha = 0.1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_s, 
               linetype = "dashed", colour = "forestgreen", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_e,
               linetype = "dashed", colour = "forestgreen", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_peak,
               linetype = "dotted", colour = "forestgreen", linewidth = 1) +
    geom_text(aes(x = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_peak, y = 0), 
              label = "Spawning", colour = "forestgreen", size = 3) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_s, 
               linetype = "dashed", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_e, 
               linetype = "dashed", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_peak, 
               linetype = "dotted", linewidth = 1) +
    geom_text(aes(x = as.Date(paste0(hist_ystart, "-01-01")) + rt_peak, y = 0), 
              label = "Run timing", colour = "black", size = 3) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + oe_s, 
               linetype = "dashed", colour = "goldenrod", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + oe_e, 
               linetype = "dashed", colour = "goldenrod", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + oe_peak, 
               linetype = "dotted", colour = "goldenrod", linewidth = 1) +
    geom_text(aes(x = as.Date(paste0(hist_ystart, "-01-01")) + oe_peak, y = 0), 
              label = "Ocean Entry", colour = "goldenrod", size = 3) +
    geom_point() +
    geom_line() +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    scale_colour_discrete(labels=c(paste0(hist_ystart,"-",hist_ystart+20), paste0(proj_ystart,"-", proj_ystart+20))) +
    labs(subtitle = "PCIC daily temperature within CU boundary accessible cells (80% range)", colour = "Year range",
         x = "Date", y = "Daily Temperature")


ct_anad_p <- ggplot() +
  geom_sf(data = amod_CU, aes(color = CT_anad), linewidth = 1.) +
  scale_color_viridis(limits = c(0,5.5)) + 
  geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
  coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
           ylim = st_bbox(cu_boundary_i)[c(2,4)]) +
  labs(subtitle = paste("Cumulative stressor score (anadromous)"), color = "Threat Score")

ct_anad_h <- ggplot(amod_CU) +
  geom_histogram(aes(x = CT_anad, fill = nuseds_presence)) +
  geom_vline(aes(xintercept = mean(CT_anad, na.rm = T)), color = "red", linetype = "dashed") +
  geom_vline(aes(xintercept = median(CT_anad, na.rm = T)), color = "blue", linetype = "dashed") +
  geom_text(aes(x = mean(CT_anad, na.rm = T), y = 20, label = "mean"), color = "red") +
  geom_text(aes(x = median(CT_anad, na.rm = T), y = 0, label = "median"), color = "blue") +
  labs(fill = "NuSEDS site")

ts_hist_p <- ggplot() +
  geom_sf(data = amod_CU, aes(color = Tw8_0_00_1), linewidth = 1.) +
  scale_color_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                        limits = c(5,25)) + 
  geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
  coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
           ylim = st_bbox(cu_boundary_i)[c(2,4)]) +
  labs(subtitle = paste("August T, Thermalscapes, 2001-2020"), color = "T (deg C)")

ts_hist_h <- ggplot(amod_CU) +
  geom_histogram(aes(x = Tw8_0_00_1, fill = nuseds_presence)) +
  geom_vline(aes(xintercept = mean(Tw8_0_00_1, na.rm = T)), color = "red", linetype = "dashed") +
  geom_vline(aes(xintercept = median(Tw8_0_00_1, na.rm = T)), color = "blue", linetype = "dashed") +
  geom_text(aes(x = mean(Tw8_0_00_1, na.rm = T), y = 20, label = "mean"), color = "red") +
  geom_text(aes(x = median(Tw8_0_00_1, na.rm = T), y = 0, label = "median"), color = "blue") +
  labs(fill = "NuSEDS site")

ts_proj_p <- ggplot() +
  geom_sf(data = amod_CU, aes(color = Tw8_9_45_3), linewidth = 1.) +
  scale_color_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                        limits = c(5,25)) + 
  geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
  coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
           ylim = st_bbox(cu_boundary_i)[c(2,4)]) +
  labs(subtitle = paste("August T, Thermalscapes, 2041-2060"), color = "T (deg C)")

ts_proj_h <- ggplot(amod_CU) +
  geom_histogram(aes(x = Tw8_9_45_3, fill = nuseds_presence)) +
  geom_vline(aes(xintercept = mean(Tw8_9_45_3, na.rm = T)), color = "red", linetype = "dashed") +
  geom_vline(aes(xintercept = median(Tw8_9_45_3, na.rm = T)), color = "blue", linetype = "dashed") +
  geom_text(aes(x = mean(Tw8_9_45_3, na.rm = T), y = 20, label = "mean"), color = "red") +
  geom_text(aes(x = median(Tw8_9_45_3, na.rm = T), y = 0, label = "median"), color = "blue") +
  labs(fill = "NuSEDS site")

### plot indicators

rateT_p <- ggplot() +
  geom_sf(data = amod_CU, aes(color = FW_SPN_EXP_rateT_9), linewidth = 1.) +
  scale_color_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                        limits = c(0,1)) +
  geom_sf(data = cu_boundary_i, color = "black", alpha = 0.3) +
  coord_sf(xlim = st_bbox(cu_boundary_i)[c(1,3)],
           ylim = st_bbox(cu_boundary_i)[c(2,4)]) +
  labs(subtitle = paste("Rate of August mean temperature change, 01-20 to 41-60"), color = "deg C/decade") +
  theme_cvis()

rateT_h <- ggplot(amod_CU) +
  geom_histogram(aes(x = FW_SPN_EXP_rateT_9, fill = nuseds_presence)) +
  geom_vline(aes(xintercept = mean(FW_SPN_EXP_rateT_9, na.rm = T)), color = "red", linetype = "dashed") +
  geom_vline(aes(xintercept = median(FW_SPN_EXP_rateT_9, na.rm = T)), color = "blue", linetype = "dashed") +
  annotate("text", x = mean(amod_CU$FW_SPN_EXP_rateT_9, na.rm = T), y = 20, label = "mean", color = "red") +
  annotate("text", x = median(amod_CU$FW_SPN_EXP_rateT_9, na.rm = T), y = 20, label = "median", color = "blue") 

### ENM favourability within CU boundary
ENM_proj_p <- ggplot() +
  geom_sf(data = reaches_ENM_cu, aes(color = Fav_f.9_45_3)) + 
  geom_sf(data = cu_boundary_i, color = "black", alpha = 0.2) +
  scale_color_viridis()

ENM_diff_p <- ggplot() +
  geom_sf(data = reaches_ENM_cu, aes(color = Fav_diff)) + 
  geom_sf(data = cu_boundary_i, color = "black", alpha = 0.2) +
  scale_color_viridis()

ENM_proj_h <- ggplot() +
  geom_histogram(data = reaches_ENM_cu, aes(x = Fav_f.9_45_3)) 

ENM_diff_h <- ggplot() +
  geom_histogram(data = reaches_ENM_cu, aes(x = Fav_diff))


order_p
cu_acc_p
hydro_day_plot / tw_day_plot
ct_anad_p + ct_anad_h
ts_hist_p + ts_hist_h
ts_proj_p + ts_proj_h
rateT_p + rateT_h
ENM_proj_p + ENM_proj_h
ENM_diff_p + ENM_diff_h

```


### Upstream migration

```{r upstream migration, echo = FALSE}

# Subset CU migration path

path_CU <- path_list[names(path_list) == cuid_i]
path_CU <- path_CU[[1]]

path_bbox <- st_bbox(path_CU)
boundary_bbox <- st_bbox(cu_boundary_i)
plot_bbox <- st_bbox(c(xmin = min(path_bbox[1], boundary_bbox[1]), ymin = min(path_bbox[2], boundary_bbox[2]), 
               xmax = max(path_bbox[3], boundary_bbox[3]), ymax = max(path_bbox[4], boundary_bbox[4])), crs = st_crs(4269))

#coast crop 
coast_crop <- st_transform(bc_coast, crs = 4269) %>%
  st_make_valid() %>%
  st_crop(plot_bbox)


PCIC_migr_CU <- PCIC_day[path_CU] 

PCIC_migr_hist_CU <- PCIC_migr_CU %>%
  filter(yday(time) >= rt_s, yday(time) <= sp_timing_peak, year(time) == hist_ystart) %>%
    aggregate(by = "year", FUN = mean)
PCIC_migr_proj_CU <- PCIC_migr_CU %>%
  filter(yday(time) >= rt_s, yday(time) <= sp_timing_peak, year(time) == proj_ystart) %>%
    aggregate(by = "year", FUN = mean)


PCIC_migr_day_CU <-  as_tibble(PCIC_migr_CU) %>%
  filter(!is.na(waterTemperature))

PCIC_migr_day_summary <- PCIC_migr_day_CU %>%
  group_by(time) %>%
  dplyr::summarize(waterTemperature_mean = mean(waterTemperature, na.rm=T),
                   waterTemperature_q10 = quantile(waterTemperature, c(0.10), na.rm = T),
                   waterTemperature_q90 = quantile(waterTemperature, c(0.90), na.rm = T),
                   discharge_mean = mean(discharge, na.rm = T),
                   discharge_q10 = quantile(discharge, c(0.10), na.rm = T),
                   discharge_q90 = quantile(discharge, c(0.90), na.rm = T))%>%
  mutate(year = as.factor(year(time)), 
         month = month(time),
         day = yday(time))



#----------------------- Migration Plots-------------------------------------------

migr_day_p <- ggplot(data = PCIC_migr_day_summary, aes(x = as.Date(paste0(hist_ystart, "-01-01")) + yday(time), y = waterTemperature_mean, ymin = waterTemperature_q10, ymax = waterTemperature_q90, colour = year)) +
  geom_ribbon(alpha = 0.1) +
  geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_s, 
             linetype = "dashed", colour = "forestgreen", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_e, 
               linetype = "dashed", colour = "forestgreen", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_peak, 
               linetype = "dotted", colour = "forestgreen", linewidth = 1) +
    geom_text(aes(x = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_peak, y = 0), 
              label = "Spawning", colour = "forestgreen", size = 3) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_s, 
               linetype = "dashed", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_e, 
               linetype = "dashed", linewidth = 1) +
    geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_peak, 
               linetype = "dotted", linewidth = 1) +
    geom_text(aes(x = as.Date(paste0(hist_ystart, "-01-01")) + rt_peak, y = 0), 
              label = "Run timing", colour = "black", size = 3) +
    geom_point() +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  scale_colour_discrete(labels=c(paste0(hist_ystart,"-",hist_ystart+20), paste0(proj_ystart,"-", proj_ystart+20))) +
  labs(subtitle = "PCIC average temperature over CU migration path (80% range)", colour = "Year range",
       x = "Date", y = "Daily Temperature")

flow_day_p <- ggplot(data = PCIC_migr_day_summary, aes(x = as.Date(paste0(hist_ystart, "-01-01")) + yday(time), y = discharge_mean, ymin = discharge_q10, ymax = discharge_q90, colour = year)) +
  geom_ribbon(alpha = 0.1) +
  geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_s, 
             linetype = "dashed", colour = "forestgreen", linewidth = 1) +
  geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_e, 
             linetype = "dashed", colour = "forestgreen", linewidth = 1) +
  geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_peak, 
             linetype = "dotted", colour = "forestgreen", linewidth = 1) +
  geom_text(aes(x = as.Date(paste0(hist_ystart, "-01-01")) + sp_timing_peak, y = 0), 
            label = "Spawning", colour = "forestgreen", size = 3) +
  geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_s, 
             linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_e, 
             linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = as.Date(paste0(hist_ystart, "-01-01")) + rt_peak, 
             linetype = "dotted", linewidth = 1) +
  geom_text(aes(x = as.Date(paste0(hist_ystart, "-01-01")) + rt_peak, y = 0), 
            label = "Run timing", colour = "black", size = 3) +
    geom_point() +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  scale_colour_discrete(labels=c(paste0(hist_ystart,"-",hist_ystart+20), paste0(proj_ystart,"-", proj_ystart+20))) +
  labs(subtitle = "PCIC average discharge over CU migration path (80% range)", colour = "Year range",
       x = "Date", y = "Discharge")

#merge_path <- st_union(path_CU, cu_boundary_i)

# path_plot <- ggplot() +
#   geom_sf(data = cu_boundary_i, color = "black", alpha = 0.2) +
#   geom_sf(data = path_CU) +
#   geom_sf(data = bc_coast, color = "black", alpha = 0.2) +
#   coord_sf(xlim = st_bbox(st_union(path_CU, cu_boundary_i))[c(1,3)],
#            ylim = st_bbox(st_union(path_CU, cu_boundary_i))[c(2,4)]) +
#   theme_cvis() +
#   labs(subtitle = paste("CU migration path"))

migr_hist_T_plot <- ggplot() +
  geom_stars(data = PCIC_migr_hist_CU["waterTemperature"]) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                       limits = c(5,25)) + 
  geom_sf(data = coast_crop, alpha = 0.1) +
  geom_sf(data = cu_boundary_i, color = "black", alpha = 0.1) +
  geom_sf(data = st_zm(path_CU)) + 
  labs(subtitle = "PCIC temp 2001-2020")
 
migr_proj_T_plot <- ggplot() +
  geom_stars(data = PCIC_migr_proj_CU["waterTemperature"]) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                       limits = c(5,25)) + 
  geom_sf(data = coast_crop, alpha = 0.1) +
  geom_sf(data = cu_boundary_i, color = "black", alpha = 0.1) +
  geom_sf(data = st_zm(path_CU)) + 
  labs(subtitle = "PCIC temp 2041-2060")
 
migr_day_p / flow_day_p
migr_hist_T_plot
migr_proj_T_plot

```

