---
title: "CVIS Overview Report"
output:
  html_document:
    toc: true
    number-sections: true
    toc_float: true
    toc_depth: 3

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

theme_cvis <- function(base_size = 14) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = rel(0.85), face = "bold"),
      axis.text = element_text(size = rel(0.70), face = "bold"),
      axis.line = element_line(color = "black"),
      # La légende
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.background = element_rect(fill = "transparent", colour = NA),
      # Les étiquettes dans le cas d'un facetting
      strip.background = element_rect(fill = "#17252D", color = "#17252D"),
      strip.text = element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
    )
}
theme_set(theme_cvis())
```


### Overview

This report summarizes climate vulnerability indicators for conservation units of Pacific salmon in the Fraser basin. Preliminary outputs of freshwater and demographic indicators are shown across a subset of conservation units. 

The following is a preliminary list of indicators and their abbreviation code. Not all indicators are currently included in this report:

```{r}
indicator_list <- read_csv(here("processed_data", "CVIS_list.csv"))

kbl(indicator_list) %>%
  kable_styling("striped", full_width = F) %>%
  column_spec(1, bold = T)
```



### Data sources

**BCFishPass** - Freshwater fish habitat accessibility model used to subset accessible streams. [https://smnorris.github.io/bcfishpass/index.html]
**CU boundaries** - spatial boundaries for CUs based on Freshwater Adaptive Zones. Taken from PSF data library [<https://data.salmonwatersheds.ca/result?datasetid=104>]\
**PCIC gridded model output** - indicators from the Salmon Climate Impacts portal [<https://pacificclimate.org/analysis-tools/salmon-climate-impacts-portal>]
**PSF CU-Specific Life History Timing** - timing of freshwater life history events for each CU, used to subset PCIC model outputs for upstream migration timing [https://datadryad.org/dataset/doi:10.5061/dryad.wm37pvmwx]



## Overview of all indicators
The following plots show standardized values of indicators across all conservation units. The values are standardized to a 0-1 scale, where 0 is the lowest value and 1 is the highest value for each indicator. 
```{r}

ggplot(STD_spn_long, aes(x = indicator, y = CU_NAME)) +
  geom_tile(aes(fill = value), color = "gray") +
  geom_text(aes(label = round(value,1)), color = "black", size = 2.5) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                        limits = c(0,1)) + 
  ylab(NULL) +
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Spawning to rearing indicators")

ggplot(STD_mig_long, aes(x = indicator, y = CU_NAME)) +
  geom_tile(aes(fill = value), color = "gray") +
  geom_text(aes(label = round(value,1)), color = "black", size = 2.5) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                        limits = c(0,1)) + 
  ylab(NULL) +
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Migration indicators")

ggplot(STD_dem_long, aes(x = indicator, y = CU_NAME)) +
  geom_tile(aes(fill = value), color = "gray") +
  geom_text(aes(label = round(value,1)), color = "black", size = 2.5) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
                        limits = c(0,1)) + 
  ylab(NULL) +
  theme(text = element_text(size = 10)) + 
    labs(title = "Demographic indicators")

```


## Freshwater spawning to rearing

### Correlation of indicators
```{r}
cor_stats <- cor(CVIS_spn %>% select(-c(cuid, CU_NAME, FULL_CU_IN, Species_simple)), use = "pairwise.complete.obs")
corrplot(cor_stats,  method = "number", tl.col = "black")

```

### Projected Temperature (projT)
The following plots show untransformed and standardized indicator values. 
```{r}

p1 <- ggplot(CVIS_STD_all) +
  geom_point(aes(x=SPN_EXP_projT_9, y=STD_SPN_EXP_projT_9, colour = Species_simple)) +
  geom_line(data = SIM_STD_all, aes(x=SPN_EXP_projT_9, y=STD_SPN_EXP_projT_9), linetype = 2) +
  ylim(0,1)

h1 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=SPN_EXP_projT_9)) +
  xlab("Raw values") +
  ylab(NULL)

h2 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=STD_SPN_EXP_projT_9)) +
  xlab("Standardized") +
  ylab(NULL)

cu_boundary_show <- cu_boundary %>%
  filter(CUID %in% cuid) %>%
  left_join(select(CVIS_STD_all, -CU_NAME), by = c("CUID" = "cuid"))

m1 <- ggplot(CVIS_STD_all) +
  geom_sf(data = cu_boundary_show, aes(fill = STD_SPN_EXP_projT_9), alpha = 0.3) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) + 
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) + 
  theme_minimal()

h1 + p1 + h2
m1



```

### Rate of temperature change (rateT)
```{r}

p1 <- ggplot(CVIS_STD_all) +
  geom_point(aes(x=SPN_EXP_rateT_9, y=STD_SPN_EXP_rateT_9, colour = Species_simple)) +
  geom_line(data = SIM_STD_all, aes(x=SPN_EXP_rateT_9, y=STD_SPN_EXP_rateT_9), linetype = 2) +
  ylim(0,1)

h1 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=SPN_EXP_rateT_9)) +
  xlab("Raw values") +
  ylab(NULL)

h2 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=STD_SPN_EXP_rateT_9)) +
  xlab("Standardized") +
  ylab(NULL)

m1 <- ggplot(CVIS_STD_all) +
  geom_sf(data = cu_boundary_show, aes(fill = STD_SPN_EXP_rateT_9), alpha = 0.3) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) + 
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) + 
  theme_minimal() +
  labs(fill = "Standardized Trate")

h1 + p1 + h2
m1

```



### Increased winter flows (winQ)
```{r}

p1 <- ggplot(CVIS_STD_all) +
  geom_point(aes(x=SPN_EXP_winQ, y=STD_SPN_EXP_winQ, colour = Species_simple)) +
  geom_line(data = SIM_STD_all, aes(x=SPN_EXP_winQ, y=STD_SPN_EXP_winQ), linetype = 2) + 
  ylim(0,1)

h1 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=SPN_EXP_winQ)) +
  xlab("Raw values") +
  ylab(NULL)

h2 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=STD_SPN_EXP_winQ)) +
  xlab("Standardized") +
  ylab(NULL)

m1 <- ggplot(CVIS_STD_all) +
  geom_sf(data = cu_boundary_show, aes(fill = STD_SPN_EXP_winQ), alpha = 0.3) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) + 
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) + 
  theme_minimal()

h1 + p1 + h2
m1

```


### Decreased summer flows (augQ)
```{r}

p1 <- ggplot(CVIS_STD_all) +
  geom_point(aes(x=SPN_EXP_augQ, y=STD_SPN_EXP_augQ, colour = Species_simple)) +
  geom_line(data = SIM_STD_all, aes(x=SPN_EXP_augQ, y=STD_SPN_EXP_augQ), linetype = 2) +
  ylim(0,1)

h1 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=SPN_EXP_augQ)) +
  xlab("Raw values") +
  ylab(NULL)

h2 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=STD_SPN_EXP_augQ)) +
  xlab("Standardized") +
  ylab(NULL)

m1 <- ggplot(CVIS_STD_all) +
  geom_sf(data = cu_boundary_show, aes(fill = STD_SPN_EXP_augQ), alpha = 0.3) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) + 
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) + 
  theme_minimal()

h1 + p1 + h2
m1

```



### Change in peak flow day (peakQ)
```{r}

p1 <- ggplot(CVIS_STD_all) +
  geom_point(aes(x=SPN_EXP_peakQday, y=STD_SPN_EXP_peakQday, colour = Species_simple)) +
  geom_line(data = SIM_STD_all, aes(x=SPN_EXP_peakQday, y=STD_SPN_EXP_peakQday), linetype = 2) +
  ylim(0,1)

h1 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=SPN_EXP_peakQday)) +
  xlab("Raw values") +
  ylab(NULL)

h2 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=STD_SPN_EXP_peakQday)) +
  xlab("Standardized") +
  ylab(NULL)

m1 <- ggplot(CVIS_STD_all) +
  geom_sf(data = cu_boundary_show, aes(fill = STD_SPN_EXP_peakQday), alpha = 0.3) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) + 
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) + 
  theme_minimal()

h1 + p1 + h2
m1

```

### Sensitivity - Cumulative threat score (CT_anad)
```{r}

p1 <- ggplot(CVIS_STD_all) +
  geom_point(aes(x=SPN_SEN_CT_anad, y=STD_SPN_SEN_CT_anad, colour = Species_simple)) +
  geom_line(data = SIM_STD_all, aes(x=SPN_SEN_CT_anad, y=STD_SPN_SEN_CT_anad), linetype = 2) +
  ylim(0,1)

h1 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=SPN_SEN_CT_anad)) +
  xlab("Raw values") +
  ylab(NULL)

h2 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=STD_SPN_SEN_CT_anad)) +
  xlab("Standardized") +
  ylab(NULL)

m1 <- ggplot(CVIS_STD_all) +
  geom_sf(data = cu_boundary_show, aes(fill = STD_SPN_SEN_CT_anad), alpha = 0.3) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) + 
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) + 
  theme_minimal()

h1 + p1 + h2
m1

```


### Sensitivity - Freshwater residence time (dur)
```{r}

p1 <- ggplot(CVIS_STD_all) +
  geom_point(aes(x=SPN_SEN_dur, y=STD_SPN_SEN_dur, colour = Species_simple)) +
  geom_line(data = SIM_STD_all, aes(x=SPN_SEN_dur, y=STD_SPN_SEN_dur), linetype = 2) +
  ylim(0,1)

h1 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=SPN_SEN_dur)) +
  xlab("Raw values") +
  ylab(NULL)

h2 <- ggplot(CVIS_STD_all) +
  geom_histogram(aes(x=STD_SPN_SEN_dur)) +
  xlab("Standardized") +
  ylab(NULL)

m1 <- ggplot(CVIS_STD_all) +
  geom_sf(data = cu_boundary_show, aes(fill = STD_SPN_SEN_dur), alpha = 0.3) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) + 
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) + 
  theme_minimal()

h1 + p1 + h2
m1

```


### Freshwater rearing to spawning overall vulnerability

```{r}

p1 <- ggplot(CVIS_STD_all) +
  geom_point(aes(x=STD_SPN_EXP_AVG, y=STD_SPN_SEN_AVG, colour = Species_simple)) +
  xlim(0,1) + 
  ylim(0,1)

m1 <- ggplot(CVIS_STD_all) +
  geom_sf(data = cu_boundary_show, aes(fill = STD_SPN_EXP_AVG), alpha = 0.3) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) + 
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) + 
  labs(title = "Average standardized exposure") + 
  theme_minimal()

m2 <- ggplot(CVIS_STD_all) +
  geom_sf(data = cu_boundary_show, aes(fill = STD_SPN_SEN_AVG), alpha = 0.3) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) + 
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) + 
  labs(title = "Average standardized sensitivity") + 
  theme_minimal()

m3 <- ggplot(CVIS_STD_all) +
  geom_sf(data = cu_boundary_show, aes(fill = STD_SPN_EXPxSEN), alpha = 0.3) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) + 
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) +
  labs(title = "Exposure x sensitivity") + 
  theme_minimal()

p1
m1
m2
m3
```



