---
title: 'Climate Vulnerability Indicators for Salmon: A Quantitative Framework for Evaluating the Relative Vulnerability of Pacific Salmon Conservation Units'
subtitle: 'DRAFT - not for further distribution without permission of the authors'
author: "Cory Lagasse, Josie Iacarella, Colin Bailey, Michael Arbeider, Timothy Healy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:  
    toc: true  
    number-sections: true  
    toc_float: true  
    toc_collapsed: true
    toc_depth: 3
    theme: flatly
    css: "CVIS_style.css"
    fig_caption: true
  word_document:
    toc: true
    toc_depth: '3'
bibliography: references.json
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, results = "asis")

# Base setup
source(here::here("code", "0_setup.R"))
load(file.path(paths$output, "scoring_results.Rdata"))
load(file.path(paths$params, "indicator_tables.Rdata"))

default_rcp <- "45"
default_period <- "3"
default_gcm <- "9"
# dsmodel_baseline <- c("station", "vicgl", "tscapes", "ensemble")

report_dat <- all_std_long %>%
  left_join(tbl_standardize %>% select(abbrev, dsmodel_baseline_ind = dsmodel_baseline), by = c("indicator" = "abbrev")) %>%
  filter(
    dsmodel == dsmodel_baseline_ind,
    period_code %in% periods_use,
    rcp %in% rcp_vec
  )
```

# Introduction

Climate change is having profound effects on Pacific salmon populations throughout their range, contributing to the declines of many stocks while introducing greater variability and uncertainty into assessments and planning. Although the significance of climate change for Pacific salmon is well established, its effects have been and will continue to be variable for the 5 species, 69 stock management units (SMUs) and over 450 conservations units (CUs) of Pacific salmon in Canada. The salmonscape of the future will be different than the past, with serious implications for the people, societies, and ecosystems of which salmon are an integral part.

A wealth of scientific research informs our understanding of future changes to freshwater and marine ecosystems, and how Pacific salmon are likely to respond to those changes [@crozier2023]. While many of the mechanisms of climate change vulnerability are understood, at least a conceptual level, there are few broad-scale assessments that evaluate the relative or absolute climate vulnerability of Pacific salmon across multiple species and wide geographic areas (e.g. Crozier et al. 2019). This gap is particularly notable across the Canadian distribution of Pacific salmon, which spans from the southern to the northern half of their range and includes a great diversity of life history types and ecosystems.

Climate change vulnerability assessments (CCVAs) can help bridge the gap between our scientific understanding and application towards resource management decisions and planning. Pacific salmon are a vitally important species, resource, and symbol within Canada, and in recent years have been the target of hundreds of millions of dollars in investments to support stock rebuilding, habitat restoration, fisheries monitoring, and hatchery enhancement. The success of these activities will be influenced by climate changes, which may lead to further declines in productivity for some stocks or outweigh the positive influences of rebuilding work. While there are many factors to be considered in prioritizing investments and making resource management decisions for Pacific salmon, assessment frameworks that are scientifically robust, transparent, and reproducible are more likely to be trusted and facilitate the incorporation of climate change considerations into decision-making [@boyce2022; @rangwala2021].

In this report, we present an indicator-based approach for assessing climate change risk factors across the Pacific salmon life-cycle. The framework incorporates recently developed, downscaled climate models in marine and freshwater domains [@weller2024; @peña2024], cumulative habitat stressors models [@murray2024], population status assessments [@csas2024], life history timing [@wilson2025], and environmental niche models [@iacarella2023] into 17 quantitative indicators. The indicators are grounded in ecological theory and represent biological mechanisms of climate change responses, including demographics, physiology, and genetic adaptation [@urban2016; @pacifici2015; @boyce2022].

We apply the climate vulnerability indicators for salmon (CVIS) framework to 5 species and 50 CUs of Pacific salmon within the Fraser River basin, an area of high life history diversity with variation in recent responses to climate change. Climate change is recognized as a prime threat to many populations within the Fraser basin, particularly for Fraser River sockeye and Interior Fraser Chinook, however, to date most assessments considering climate vulnerability have focused on specific CUs, species, or watersheds [@doutaz2023; @dionne2023]. Furthermore, many threats assessments and other attempts to evaluate climate vulnerability have been centered around expert judgement and categorical traits, limiting comparability across assessments.

By quantifying multiple mechanisms of climate vulnerability across a large basin, we aim to better understand differences in climate change risk at the individual CU, species, and basin level. In particular, we investigate which CUs are most likely to be at risk from climate change during their freshwater rearing, upstream migration, and nearshore marine life stages, and which factors are most important in differentiating that risk among CUs and species. We also quantify the sensitivity of risk rankings across different time periods (mid-century versus end-of-century), emissions scenarios, global climate models, and scoring frameworks to understand the influence of different methods and objectives for assessment. Ultimately, our framework provides a means for identifying patterns of climate change vulnerability across spatial areas and demographic units, and testing the sensitivity of those patterns to different models and assumptions to help identify the drivers of climate vulnerability.

# Methods Overview

The CVIS framework consists of 17 quantitative indicators (Table 1) classified into five categories based on mechanism and life-stage: freshwater spawning and rearing, upstream migration, nearshore marine, demographics, and genetics. Every indicator uses a quantitative metric, such as the mean stream temperature or percent change in flow, to represent risk. The indicator approach allows for flexibility in inputs and metrics of risk; various model types can be used to derive statistical summaries of risk depending on the mechanism to assess. Variation and uncertainty in indicator values are represented using 80% quantile ranges of global climate models where relevant.

```{r}

tbl_ind_report %>%
  arrange(Category) %>%               # ensure groups are contiguous
  gt(groupname_col = "Category") %>%  # turn Category into row-group headers
  # Make the header (column labels) bold
  tab_style(
    style = cell_text(weight = "bold", align = "center"),
    locations = cells_column_labels()
  ) %>%
  # Center the row-group headers
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_row_groups()
  ) %>%
  tab_options(
    table.font.size = "small",
    data_row.padding = px(1),
    row_group.padding = px(2),
    row_group.border.top.color = "gray75",
    row_group.border.bottom.color = "gray75"
  )



```

The framework focuses on CUs of Pacific salmon as the assessment unit. As the unit of biodiversity for Pacific salmon in Canada, CUs represent groupings based on life history characteristics, freshwater and marine adaptive zones, and genetic structure [@Holtby2007]. The number of populations and spatial distribution of CUs is variable, with some CUs representing just a single population with fewer than a thousand spawners per year while other CUs include dozens of populations and millions of spawners per year on average. Despite this variability, conservation at the CU level is critical for maintaining biodiversity under the Wild Salmon Policy, and the CU is an important unit for assessment that is used to determine aggregate stock status [@csas2024].

**Indicator Standardization**\
All indicators are standardized using response functions to translate raw values to a scale from 0 to 1, where 0 represents lowest risk and 1 is the highest risk relative to other CUs. Standardization facilitates easier interpretation and comparison of indicators, and allows the representation of risk based on stressor-response relationships and thresholds [@boyce2022]. Where relevant and defensible, physiological or ecological thresholds, such as optimal temperature limits, are used to parameterize standardization functions. This allows risk scores to be zeroed for environmental changes and other factors in situations where they are likely to remain within safe limits or have a positive influence on productivity.

**Overall Vulnerability Scoring**

Individual indicators represent a single mechanism or metric of climate vulnerability, but overall vulnerability is determined by the cumulative effects of multiple mechanisms and stressors that act within and across life stages. We apply multiple scoring methods for overall vulnerability to explore different causal relationships and interactions among individual indicators and categories. Comparing among scoring methods allows us to evaluate how relationships among indicators, weighting of indicators, and weighting of categories affect the relative vulnerability rankings among CUs. We summarize results and variation across four scoring methods that each reflect different assumptions for determining overall vulnerability.

**Emissions scenarios and time period**\
CCVAs can assess vulnerability for a particular time period or multiple time periods, as well as different emissions scenarios depending upon assessment objectives [@foden2013]. Our indicators that evaluate environmental change are based primarily on the CMIP5 suite of global climate models and results are presented focusing on the mid-century period under representative concentration pathway (RCP) 4.5, representing moderate emission reductions compared to business as usual. The influence of different emissions scenarios and time periods are assessed by evaluating changes in relative vulnerability rankings across RCP 4.5 and 8.5, and for mid-century (2041-2060) versus end-of-century (2081-2100) time periods.

# Freshwater Spawning to Rearing

#### Data Sources

**BCFishPass** - Freshwater fish habitat accessibility model used to subset accessible streams. [<https://smnorris.github.io/bcfishpass/index.html>]\
**CU boundaries** - spatial boundaries for CUs based on Freshwater Adaptive Zones. Taken from PSF data library [<https://data.salmonwatersheds.ca/result?datasetid=104>] **Cumulative Threats** - Cumulative threats to anadromous salmon stream habitat in the Fraser Basin [<https://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2024/2024_041-eng.pdf>]\
**Thermalscapes Stream Temperatures** - Statistically downscaled projections of August stream temperature developed across British Columbia [<https://datadryad.org/dataset/doi:10.5061/dryad.bzkh189fk>]

**Environmental Niche Models** - Models of current and future habitat favourability for British Columbia and the Western US [<https://cdnsciencepub.com/doi/10.1139/cjfas-2023-0096>]

**PCIC Variable Infiltration Capacity With Glaciers (VIC-GL) gridded model** - Dynamically downscaled projections of stream temperature and flow from 1980 to 2099 for gauged watersheds within British Columbia [<https://pacificclimate.org/analysis-tools/salmon-climate-impacts-portal>]

**PSF CU-Specific Life History Timing** - timing of life history events for determining freshwater residency time [<https://datadryad.org/dataset/doi:10.5061/dryad.wm37pvmwx>]

**August flow projections at gauge stations** - Preliminary model outputs of statistically downscaled flow projections for gauge stations in British Columbia. Developed by Sacha Ruzzante et al.

#### Methods

Vulnerability indicators for freshwater spawning and rearing life stages represent environmental change to streams, as well as cumulative threats to stream habitats and freshwater residency times. The British Columbia Freshwater Atlas was used as the stream network with spatial delineation to Pacific salmon conservation units based upon conservation unit boundaries. The stream network was limited to accessible streams with channel width and gradients that meet criteria for intrinsic habitat potential for each Pacific salmon species using the [BC Fishpass accessibility model](https://smnorris.github.io/bcfishpass/02_model_access.html) and [habitat model](https://smnorris.github.io/bcfishpass/03_model_habitat_linear.html). For sockeye salmon, the habitat model also includes criteria for proximity to lakes. This filtering process removed smaller and inaccessible streams while retaining all stream segments with observed occurrences and modelled habitat potential.

Environmental change indicators included projected August stream temperatures, rate of change in August stream temperature, percent change in August flows, percent change in flows from November to January, and change in environmental niche model habitat suitability. Each indicator was calculated as the mean value across all stream segments assigned to the conservation unit, weighted by stream segment length. Temperature indicators (projected and rate of change) were based on a statistical downscaling approach fitted to streams throughout British Columbia [@weller2024]. Hydrologic indicators were based on the Pacific Climate Impacts Consortium gridded hydrological model that was further downscaled to the Fraser basin stream network [@iacarella2025]. For the indicator of percent change in August flows, an alternative set of projections using a statistical downscaling approach fitted to hydrometric stations in BC (Ruzzante in prep) was also applied to compare variation in responses (to be further described in an appendix).

Variation in environmental change indicators for each CU was summarized across global climate model projections and spatial dimensions using 80% quantile ranges. Spatial variation quantiles were weighted by stream length and characterized the range in responses for accessible streams with intrinsic habitat potential using the ensemble mean of climate model projections. Global climate model quantiles represented the variation of the weighted mean for streams within the CU boundary across all included global climate models (generally 6).

### i. Projected August stream temperature (tw8proj) {.tabset .tabset-pills}

**Metric:** The average of August stream temperatures across all years of the projection period (2041-2060 or 2081-2100) over all accessible streams with modelled habitat potential contained in the conservation unit (CU) boundary. Streams are weighted by stream segment length when calculating the average value. Projections from representative concentration pathways (RCP) 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

For additional description of indicator see [A.1 Projected August stream temperature (tw8proj)](#a1.-projected-august-stream-temperature-tw8proj)

```{r projT plots}
# fw_indicators <- tbl_indicators$abbrev[tbl_indicators$type %in% c("fwR")]

for (rcp_pick in rcp_vec)
{
  cat("#### ", rcp_pick, "\n\n", sep = "")

  filter_std <- report_dat %>%
    filter(
      rcp == rcp_pick,
      period_code == default_period
    )

  make_indicator_plots(filter_std,
    ind_pick = "tw8proj"
  )

  cat("\n\n")
}
```

### ii. Rate of change in August temperature (tw8rate) {.tabset .tabset-pills}

**Metric:** The decadal rate of increase in August mean stream temperature from a baseline period of 1981-2000 to the projection period (2041-2060 or 2081-2100) averaged over all accessible streams with modelled habitat potential contained in the CU boundary. Streams are weighted by stream segment length when calculating the average value. Projections from RCP 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

For additional description of indicator see [A.2 Rate of increase in August stream temperature (tw8rate)](#a.2-rate-of-increase-in-august-stream-temperature-tw8rate)

```{r projrate}
for (rcp_pick in rcp_vec)
{
  cat("#### ", rcp_pick, "\n\n", sep = "")

  filter_std <- report_dat %>%
    filter(
      rcp == rcp_pick,
      period_code == default_period,
      dsmodel %in% dsmodel_baseline
    )

  make_indicator_plots(filter_std,
    ind_pick = "tw8rate"
  )

  cat("\n\n")
}
```

### iii. Proportional decrease in August flows (lowQpdelta) {.tabset .tabset-pills}

**Metric:** The change in average August flow from a baseline period of 1981-2010 to the projection period (2041-2060 or 2081-2100), expressed as a proportion of the average historical flow and averaged over all accessible streams with modelled habitat potential contained in the CU boundary. Streams are weighted by stream segment length when calculating the average value. Projections from RCP 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

For additional description of indicator see [A.3 Proportional decrease in August flows (lowQpdelta)](#a.3-proportional-decrease-in-august-flows-lowqpdelta)

```{r}
for (rcp_pick in rcp_vec)
{
  cat("#### ", rcp_pick, "\n\n", sep = "")

  filter_std <- report_dat %>%
    filter(
      rcp == rcp_pick,
      period_code == default_period,
      dsmodel %in% dsmodel_baseline
    )

  make_indicator_plots(filter_std,
    ind_pick = "flow8pdelta",
    spatial_pal_dir = 1
  )

  cat("\n\n")
}
```

### iv. Proportional increase in November to January flows (highQpdelta) {.tabset .tabset-pills}

Metric: The change in average flow for the months of November, December and January from a baseline period of 1981-2010 to the projection period (2041-2060 or 2081-2100), expressed as a proportion of the average historical flow and averaged over all accessible streams with modelled habitat potential contained in the CU boundary. Streams are weighted by stream segment length when calculating the average value. Projections from RCP 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

For additional description of indicator see [A.4 Proportional increase in November to January flows (highQpdelta)](#a.4-proportional-increase-in-november-to-january-flows-highqpdelta)

```{r}
for (rcp_pick in rcp_vec)
{
  cat("#### ", rcp_pick, "\n\n", sep = "")

  filter_std <- report_dat %>%
    filter(
      rcp == rcp_pick,
      period_code == default_period,
      dsmodel %in% dsmodel_baseline
    )

  make_indicator_plots(filter_std,
    ind_pick = "flow18pdelta"
  )

  cat("\n\n")
}
```

### v. Ecological Niche Model change in habitat favourability (favchange) {.tabset .tabset-pills}

**Metric:** The change in habitat favourability from the baseline period (1981-2000) to the projection period (2041-2060 or 2081-2100), calculated as the difference between future and baseline favourability values and averaged over all accessible streams within the CU boundary. Habitat favourability is derived from environmental niche models (ENMs) fitted to salmon occurrence data and environmental predictors, providing a metric of environmental suitability that ranges from 0 to 1. The future favourability values use an ensemble mean of six GCMs from RCP 4.5 and 8.5 scenarios for the environmental predictors.

For additional description of indicator see [A.5 Ecological Niche Model change in habitat favourability (favchange)](#a.5-ecological-niche-model-change-in-habitat-favourability-favchange)

```{r}
for (rcp_pick in rcp_vec)
{
  cat("#### ", rcp_pick, "\n\n", sep = "")

  filter_std <- report_dat %>%
    filter(
      rcp == rcp_pick,
      period_code == default_period,
      dsmodel %in% dsmodel_baseline
    )

  make_indicator_plots(filter_std,
    ind_pick = "favchange",
    spatial_pal_dir = 1
  )

  cat("\n\n")
}
```

### vi. Cumulative threats to stream habitat (CT)

**Metric:** A composite index of threats from human activity to stream habitat, calculated using a normalized additive model that combines nine threat categories including flow alteration, in-stream habitat destruction, riparian disturbance, nutrients, pollution, and sedimentation. The indicator is calculated as the mean cumulative threat score across all accessible streams with modelled habitat potential in the CU boundary, weighted by stream segment length.

For additional description of indicator see [A.6 Cumulative threats to stream habitat (ct)](#a.6-cumulative-threats-to-stream-habitat-ct)

```{r}

make_indicator_plots(report_dat,
  ind_pick = "cthr"
)
```

### vii. Freshwater residency time

**Metric:** The duration in days that juvenile salmon reside in freshwater from the start of egg incubation to ocean entry. For each CU, residency time is calculated using as the number of days from peak spawn timing to peak ocean entry timing, plus the typical number of years spent in freshwater (0 or 1).

For additional description of indicator see [A.7 Freshwater residency time (fwres)](#a.7-freshwater-residency-time-fwres)

```{r}

make_indicator_plots(report_dat,
  ind_pick = "fwres",
  make_spatial = F
)
```

# Upstream Migration

#### Data Sources

**PCIC Variable Infiltration Capacity With Glaciers (VIC-GL) gridded model** - Dynamically downscaled projections of stream temperature and flow from 1980 to 2099 for gauged watersheds within British Columbia [<https://pacificclimate.org/analysis-tools/salmon-climate-impacts-portal>]\
**PSF CU-Specific Life History Timing** - timing of freshwater life history events for each CU used for run timing and spawn timing [<https://datadryad.org/dataset/doi:10.5061/dryad.wm37pvmwx>]

**NUSEDS (New Spawning Escapement Database)** - CU spawning sites used to determine migration end points [<https://open.canada.ca/data/en/dataset/c48669a3-045b-400d-b730-48aafe8c5ee6>]

#### Methods

Indicators for the upstream migration life stage represent changes to stream temperatures and flows, and distance to spawning sites. For each CU, a migration path from river mouth to each NUSEDS spawning sites was created using the BC Freshwater Atlas stream network. Migration distance for each path was calculated as the length of all stream segments from river mouth to spawning site, and averaged across all NuSEDS sites to get an average migration distance for each CU.

To estimate temperatures and flow during upstream migration, an upstream movement model was applied to stream segments within the migration path and used to subset days of the year and spatial grid cells from the VIC-GL model. The model began with stream segments within 100km of the river mouth at the start of run timing. The front end of this spatial subset was moved upstream at a rate equal to the distance to spawning sites divided by the number of days between spawn timing start and run timing start. This assumes that the spawn timing start coincides with arrival to the spawning sites. The tail end of the migration was modelled to arrive at the spawning grounds at the peak spawn date, and moved upstream from the river mouth at a rate equal to the distance to spawning sites divided by the number of days between spawn timing peak and run timing end.

### i. Projected temperature during migration {.tabset .tabset-pills}

**Metric:** The average of water temperatures during the upstream migration across all years of the projection period (2041-2060 or 2081-2100) estimated along the migration path from river entry to spawning sites using an upstream migration model. The model uses a moving migration window to represent the thermal exposure of CUs starting at the river mouth on the run timing start date and terminating at spawning sites on the peak spawning date. Projections from representative concentration pathways (RCP) 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

For additional description of indicator see [A.8 Projected temperature during upstream migration (migrT)](#a.8-projected-temperature-during-upstream-migration-migrt)

```{r}
for (rcp_pick in rcp_vec)
{
  cat("#### ", rcp_pick, "\n\n", sep = "")

  filter_std <- report_dat %>%
    filter(
      rcp == rcp_pick,
      period_code == default_period
    )

  make_indicator_plots(filter_std,
    ind_pick = "migrTproj",
    make_spatial = F
  )

  cat("\n\n")
}
```

### ii. Proportional change in flow during upstream migration {.tabset .tabset-pills}

**Metric:** The proportional change in flow during the upstream migration from the baseline period (1981-2000) to the projection period (2041-2060 or 2081-2100) estimated along the migration path from river entry to spawning sites using an upstream migration model. The model uses a moving migration window to represent the average flow experienced by CUs starting at the river mouth on the run timing start date and terminating at spawning sites on the peak spawning date. Projections from representative concentration pathways (RCP) 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

For additional description of indicator see [A.9 Proportional change in flow during upstream migration (migrQ)](#a.9-proportional-change-in-flow-during-upstream-migration-migrq)

```{r}
for (rcp_pick in rcp_vec)
{
  cat("#### ", rcp_pick, "\n\n", sep = "")

  filter_std <- report_dat %>%
    filter(
      rcp == rcp_pick,
      period_code == default_period
    )

  make_indicator_plots(filter_std,
    ind_pick = "migrQpdelta",
    make_spatial = F
  )

  cat("\n\n")
}
```

### iii. Upstream migration distance {.tabset .tabset-pills}

**Metric:** The distance of the migration path from river mouth to spawning sites. For CUs with multiple spawning locations, the average of the migration distance to all spawning sites is used.

For additional description of indicator see [A.10 Upstream migration distance (migrdist)](#a.10-upstream-migration-distance-migrdist)

```{r}

make_indicator_plots(report_dat,
                     ind_pick = "migrdist",
                     make_spatial = F
)

```

# Nearshore Marine

#### Data Sources

**PSF CU-Specific Life History Timing** - ocean entry timing used to determine projected SST [<https://datadryad.org/dataset/doi:10.5061/dryad.wm37pvmwx>]

**SST using Quantile Delta Mapping** - CMIP6 earth system model SST values upscaled to a higher resolution using a quantile delta mapping bias correction method. Data set currently in prep for publication by ECCC (Dosser pers comm.).

**Marine Adaptive Zones** - Developed as part of CU definitions by Holtby and Ciruna (2007), adjusted to cover the marine regions.

**HOTSSea -** hindcast of the Salish Sea used for the historical climatology of SST [<https://gmd.copernicus.org/articles/18/211/2025/>]. Currently used for sensitivity analysis.

**SalishSeaCast** - Regional oceanographic model of the Salish Sea used for projected SST between 2046-2065. Currently used for sensitivity analysis.

**Cumulative Impacts -** Marine habitat cumulative impacts [<https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2024/2024_024-eng.html>]

#### Methods

Marine indicators represent changes to sea surface temperature and cumulative habitat impacts within marine adaptive zones surrounding the period of marine entry for each conservation unit. Sea surface temperature projections from CMIP6 models were averaged for the two months prior peak ocean entry to 2 months following peak ocean entry. Historical climatologies were estimated using the same method but averaged from the HOTSSea model for better representation of historic SSTs.

The spatial domain over which SSTs were averaged was the Strait of Georgia marine adaptive zone. The same spatial domain was applied to all CUs from the Fraser basin as they all share a common ocean entry point.

### i. Projected sea surface temperature near ocean entry {.tabset .tabset-pills}

**Metric:** The projected sea surface temperature (SST) across all years of the projection period (2041-2060 or 2081-2100) throughout the marine adaptive zone (MAZ) for the CU over a 6-month period surrounding the peak month of ocean entry. Projections from RCP 4.5 and 8.5 are included with an ensemble mean calculated from 28 global climate models.

For additional description of indicator see [A.11 Projected nearshore sea surface temperature (SSTproj)](#a.11-projected-nearshore-sea-surface-temperature-sstproj)

```{r}
for (rcp_pick in rcp_vec)
{
  cat("#### ", rcp_pick, "\n\n", sep = "")

  filter_std <- report_dat %>%
    filter(
      rcp == rcp_pick,
      period_code == default_period
    )

  make_indicator_plots(filter_std,
    ind_pick = "SSTproj",
    standardized_plots = F,
    make_spatial = F
  )

  cat("\n\n")
}
```

### ii. Rate of change in sea surface temperature near ocean entry {.tabset .tabset-pills}

**Metric:** The decadal rate of increase in SST from a baseline period to the projection period (2041-2060 or 2081-2100) throughout the marine adaptive zone (MAZ) for the CU over a 6-month period surrounding the peak month of ocean entry. Projections from RCP 4.5 and 8.5 are included with an ensemble mean calculated from 28 global climate models.

For additional description of indicator see [A.12 Rate of increase in nearshore sea surface temperatures (SSTrate)](#a.12-rate-of-increase-in-nearshore-sea-surface-temperatures-sstrate)

```{r}
for (rcp_pick in rcp_vec)
{
  cat("#### ", rcp_pick, "\n\n", sep = "")

  filter_std <- report_dat %>%
    filter(
      rcp == rcp_pick,
      period_code == default_period
    )

  make_indicator_plots(filter_std,
    ind_pick = "SSTrate",
    standardized_plots = F,
    make_spatial = F
  )

  cat("\n\n")
}
```

### iii. Cumulative impacts to nearshore marine habitat (CImpact)

**Metric:** The average cumulative impact score of nearshore habitats averaged throughout the MAZ for a CU. Cumulative impact scores are calculated for multiple habitat categories within the Pacific region including pelagic, benthic, eelgrass, and kelp forests.  Stressors included as part of the additive cumulative impact model include aquaculture, fishing, invasive species, nutrient inputs, pollution inputs, and shipping [@murray2024].

For additional description of indicator see [A.13 Cumulative impacts to nearshore marine habitat (CImpact)](#a.13-cumulative-impacts-to-nearshore-marine-habitat-cimpact)

```{r}

make_indicator_plots(report_dat,
  ind_pick = "CImpact",
  standardized_plots = F,
  make_spatial = F
)
```

# Demographics

#### Data Sources

**WSP Status Assessments** - Recent assessments for Fraser basin conservation units using methods described in CSAS 2024 [<https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ScR-RS/2024/2024_004-eng.html>]

#### Methods

Demographic indicators are derived from spawner abundance data and recent Wild Salmon Policy assessments of CU status in the Fraser Basin. The CU status indicator represents the categorical status designation (Red, Amber, Green) using the Rapid Status Assessment Approach in 2022/23 [@pestal2023]. The recent generational average of spawners is derived from the same data set used to assess status and determined based on the geometric average of the number of spawners over the most recent generation with available estimates.

### i. Conservation Status

**Metric:** Recent Wild Salmon Policy (WSP) conservation status of a CU categorized as either Red, Amber, or Green. Status was assessed using the Rapid Status Assessment Approach that uses a Yes/No decision tree considering the number of recent spawners, trends in abundance, data type (absolute or relative abundance), and biological benchmarks (Pestal et al. 2023; CSAS 2024). The input data, thresholds and appropriate metrics, and final status of Fraser CUs in 2022/2023 were evaluated and agreed upon by expert judgement throughout the Rapid Status process.

For additional description of indicator see [A.14 Conservation status (CUstatus)](#a.14-conservation-status-custatus)

```{r}

make_indicator_plots(report_dat,
  ind_pick = "CUstatus",
  standardized_plots = T,
  make_spatial = F
)

cat("\n\n")
```

### ii. Recent generational average of spawners (CUnmat)

**Metric:** The geometric average of natural origin spawners (male and female) over the most recent generation with available estimates. Sockeye spawners are further refined to only include effective female spawners and not total male and female. For cyclic sockeye CUs where abundance is dominated by a single-year on the cycle line, the number of spawners from the most recent dominant year is used instead of the average over all years of a generation.

For additional description of indicator see [A.15 Recent generational average of spawners (Cunmat)](#a.15-recent-generational-average-of-spawners-cunmat)

```{r}

make_indicator_plots(report_dat,
  ind_pick = "CUnmat",
  standardized_plots = F,
  make_spatial = F
)

cat("\n\n")
```

# Genetics

#### Data Sources

**MGL Adaptive Genetic Variation Analysis** - All genetic indicator data was provided by the PBS Molecular Genetics Lab as part of a study on adaptive genetic variation of Pacific salmon (Healy et al, in prep)

#### Methods

Genetic indicators assess the genetic differences among populations to evaluate the raw material for future selection (heterozygosity) and the differences between current genetic composition and optimal genetic composition under future conditions (genomic offset). The genetic indicators described here come from a genetic analysis of Pacific salmon populations across their North American range (Healy et al, in prep), however, only populations within the Fraser basin are currently included. The analysis uses low-coverage sequencing to evaluate genome-wide variation of individuals within populations. For each population, 7-8 individuals were sequenced using samples from the DNA archive of the Pacific Biological Station Genetic Stock Identification program.

Measures of genetic diversity and genomic offset are not comparable across species, therefore, each indicator was standardized based on the range of values within a species, rather than variation across all species used for the other indicators. Genetic indicators are currently available for sockeye and Chinook salmon CUs. Coho salmon will also be added in the future.

### i. Heterozygosity

**Metric:** Heterozygosity is a common metric of genetic diversity within populations that quantifies the proportion of loci that have different alleles. Higher heterozygosity is considered indicative of greater adaptive potential to changing environmental conditions because there is more raw material upon which selection can act. Heterozygosity is measured at the individual level as the proportion of heterozygous sites, which are then averaged among all samples to get a population average value. Where multiple populations have been measured, the average across populations is taken to determine CU heterozygosity.

For additional description of indicator see [A.16 Heterozygosity (hetzyg)](#a.16-heterozygosity-hetzyg)

```{r}


make_indicator_plots(report_dat,
  ind_pick = "hetzyg",
  standardized_plots = F,
  make_spatial = F
)
```

### ii. Genomic offset

**Metric:** The genomic offset is calculated as the Euclidean distance between a population's current genetic composition and the genetic composition predicted to be optimal under future climate conditions. This distance represents the expected allele frequency change required for the population to track projected climate change, in this case based on 19 bioclim variables [<https://www.worldclim.org/data/bioclim.html>] in freshwater habitat under the RCP 8.5 scenario. Where multiple populations have been genotyped within a CU, the average genomic offset of populations was used to provide the indicator value for the CU.

For additional description of indicator see [A.17 Genomic offset (genoff)](#a.17-genomic-offset-genoff)

```{r}


make_indicator_plots(report_dat,
  ind_pick = "genoff",
  standardized_plots = F,
  make_spatial = F
)
```

# Overall Vulnerability

### Methods

Individual indicators represent one type of mechanism or trait related to climate vulnerability, however, to better understand overall vulnerability to climate change we combine standardized scores within and across indicator categories. Various methods are used in CCVAs and risk assessments for determining overall risk, such as additive or multiplicative scores, threshold approaches (Fisheries and Oceans Canada 2019), and logic models (Crozier et al. 2019), with the selection of method depending on the analytic framework and the putative mechanisms through which individual risk factors combine to determine overall risk.

For the CVIS framework, we explore four different methods that represent alternative hypotheses for how indicators and vulnerability within categories will combine to determine the overall vulnerability to climate change. Specifically, we use 1) a sum-of-averages approach that averages the values of all indicators within a category and then sums the category scores to obtain an overall vulnerability score. This gives equal weight to each category regardless of how many indicators it contains, balancing the influence of the different life stage groupings and the genetics and demographic categories. We also use 2) an average-of-all-indicators method that ignores the category type and gives all indicators an equal weighting. In addition, we use 3) a sum-of-cubic means approach that takes the cubic mean of indicator values to get a category score, then adds the category scores to determine the overall score. A cubic mean places greater weighting on high indicator values when determining category scores compared to an averaging approach. The final method is 4) a logic model that bins the standardized indicator scores into three risk categories of low, moderate, and high using fixed thresholds of 0.33 and 0.66, then determines overall vulnerability using logic rules based on the number of indicators within each category.

### Category Scores {.tabset .tabset-pills}

```{r fw-indicators, fig.width=12, fig.height=10, dpi=300, out.width="100%", fig.cap="Freshwater spawning and rearing vulnerability indicators across Fraser River salmon conservation units."}
for (category_pick in unique(tbl_indicators$long_type))
{
  cat("#### ", category_pick, "\n\n", sep = "")

  type_pick <- tbl_indicators$category[tbl_indicators$long_type == category_pick][1]

  fw_p <- indicator_tile_plot(
    all_std_long,
    indicators_choose = tbl_indicators$abbrev[tbl_indicators$category == type_pick],
    indicators_metadata = tbl_indicators,
    rcp_pick = default_rcp,
    period_pick = as.character(default_period),
    dsmodel_pick = dsmodel_baseline
  )
  print(fw_p)
  cat("\n\n")
}
```

### Species Scores {.tabset .tabset-pills}

```{r species scores, fig.width=10, fig.height=8, dpi=300, out.width="100%"}
for (species_pick in unique(cu_run$SPECIES_NAME))
{
  cat("#### ", species_pick, "\n\n", sep = "")

  p <- species_category_tile_plot_from_scores(
    all_std_long = all_std_long,
    scores_tidy = scores_tidy,
    species_pick = species_pick,
    species_palette = species_palette,
    indicators_metadata = tbl_indicators,
    rcp_pick = default_rcp,
    period_pick = as.character(default_period),
    dsmodel_pick = dsmodel_baseline
  )

  print(p)
  cat("\n\n")
}
```

### Total Scores {.tabset .tabset-pills}

The following plots show total scores and CU vulnerability ranks across indicators.

```{r combined scores, fig.width=10, fig.height=8, dpi=300, out.width="100%"}
for (rcp_pick in unique(scores_tidy$rcp))
{
  cat("#### ", rcp_pick, "\n\n", sep = "")

  # Format scores_tidy to work with indicator_tile_plot
  cat_scores_as_ind <- scores_tidy %>%
    filter(
      rcp == rcp_pick, period_code == default_period,
      method == "avg", category %in% c("fwrs", "migr", "mar", "dem", "gen")
    ) %>%
    mutate(indicator = paste0("avg_", category), std_value = score)

  overall_scores_as_ind <- scores_tidy %>%
    filter(
      rcp == rcp_pick, period_code == default_period,
      method == "catavg", category == "all"
    ) %>%
    mutate(indicator = "catavg", std_value = score)

  combined_for_tile <- bind_rows(cat_scores_as_ind, overall_scores_as_ind)

  # For GCMs, default to 9 (ensemble mean)
  if ("gcm" %in% names(combined_for_tile)) {
    combined_for_tile <- combined_for_tile %>% filter(gcm == "9" | gcm == 0 | gcm == min(gcm))
  }

  p <- indicator_tile_plot(combined_for_tile,
    indicators_choose = c("avg_fwrs", "avg_migr", "avg_mar", "avg_dem", "avg_gen", "catavg"),
    overall_score_cols = "catavg",
    category_name = "Average vulnerability by category and total vulnerability"
  )

  print(p)
  cat("\n\n")
}
```

# Sensitivity Analysis

One test for the utility of an indicator is the correlation with other indicator values, and how uncertainty propagates from models into vulnerability rankings. Here we present granular uncertainty and multivariate risk profile analysis.

```{r load_sens_data, include=FALSE}
library(reshape2)
library(patchwork)
# Load new Rdata objects
load(file.path(paths$output, "sensitivity_analysis_4b_directional.Rdata"))
load(file.path(paths$output, "risk_drivers_4d.Rdata"))

# Unpack results
contributions <- sensitivity_analysis_4b_directional$contributions
mrd_stats <- sensitivity_analysis_4b_directional$mrd_stats
cor_list <- sensitivity_analysis_4b_directional$cor_matrix
categories <- sensitivity_analysis_4b_directional$categories

red_flag_summary <- risk_drivers_4d$red_flag_summary
variation_summary <- risk_drivers_4d$variation_summary
pca_output <- risk_drivers_4d$pca_output
pca_result <- risk_drivers_4d$pca_result

source_colors <- sens_source_palette
```

### Granular Uncertainty Contributions {.tabset .tabset-pills}

#### Relative Contribution
```{r relative_contribution, fig.width=10, fig.height=6, dpi=300, out.width="100%"}
prop_long <- contributions %>%
  select(FULL_CU_IN, category, starts_with("prop_")) %>%
  pivot_longer(cols = starts_with("prop_"), names_to = "source", values_to = "proportion") %>%
  mutate(source = str_remove(source, "prop_"))

prop_summary <- prop_long %>%
  group_by(category, source) %>%
  summarise(mean_prop = mean(proportion, na.rm = TRUE), .groups = "drop")

p_prop <- ggplot(prop_summary, aes(x = category, y = mean_prop, fill = source)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = source_colors) +
  labs(
    title = "Mean Relative Contribution of Uncertainty Sources",
    subtitle = "Average across all CUs per category",
    x = "Evaluation Category",
    y = "Proportion of Total Absolute Deviation",
    fill = "Uncertainty Source"
  ) +
  theme_cvis() +
  coord_flip()

print(p_prop)
```

#### Directional Influence
```{r directional_influence, fig.width=14, fig.height=9, dpi=300, out.width="100%"}
raw_long <- contributions %>%
  select(FULL_CU_IN, category, starts_with("raw_dev_")) %>%
  pivot_longer(cols = starts_with("raw_dev_"), names_to = "source", values_to = "raw_deviation") %>%
  mutate(source = str_remove(source, "raw_dev_"))

p_direction <- ggplot(raw_long, aes(x = source, y = raw_deviation, fill = source)) +
  geom_boxplot(alpha = 0.7, outlier.size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  facet_wrap(~category) +
  scale_fill_manual(values = source_colors) +
  labs(
    title = "Directional Influence of Scenarios and Methods",
    subtitle = "Distribution of (Scenario Score - Baseline Score) across CUs",
    x = "Variation Source",
    y = "Score Change (Directional)",
    fill = "Source"
  ) +
  theme_cvis() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(p_direction)
```

#### Mean Rank Displacement
```{r mean_rank_displacement, fig.width=12, fig.height=8, dpi=300, out.width="100%"}
p_mrd <- ggplot(mrd_stats, aes(x = reorder(source, MRD), y = MRD, fill = source)) +
  geom_bar(stat = "identity") +
  facet_wrap(~category, scales = "free_x") +
  scale_fill_manual(values = source_colors) +
  labs(
    title = "Mean Rank Displacement",
    subtitle = "Average shift in vulnerability rank compared to baseline",
    x = "Scenario Shift",
    y = "Mean Displacement (Ranks)",
    fill = "Source"
  ) +
  theme_cvis() +
  coord_flip()

print(p_mrd)
```

### Risk Profile Drivers {.tabset .tabset-pills}

#### PCA Risk Profile
```{r pca_risk_profile, fig.width=11, fig.height=8, dpi=300, out.width="100%"}
loadings <- as.data.frame(pca_result$rotation[, 1:2])
loadings$indicator <- rownames(loadings)

scale_fact <- max(abs(pca_output$PC1)) / max(abs(loadings$PC1)) * 0.7
loadings <- loadings %>%
  mutate(
    x_end = PC1 * scale_fact,
    y_end = PC2 * scale_fact
  )

p_base_pca <- ggplot(pca_output, aes(x = PC1, y = PC2)) +
  geom_text(aes(label = FULL_CU_IN), size = 2.2, vjust = 1.6, alpha = 0.5, fontface = "bold") +
  geom_segment(
    data = loadings %>% filter(abs(PC1) > 0.15 | abs(PC2) > 0.15),
    aes(x = 0, y = 0, xend = x_end, yend = y_end),
    arrow = arrow(length = unit(0.2, "cm")), color = "grey40"
  ) +
  geom_text(
    data = loadings %>% filter(abs(PC1) > 0.15 | abs(PC2) > 0.15),
    aes(x = x_end * 1.1, y = y_end * 1.1, label = indicator),
    size = 3.5, color = "black", fontface = "bold"
  ) +
  labs(
    subtitle = paste(
      "Variance Explained: PC1 =",
      round(pca_result$sdev[1]^2 / sum(pca_result$sdev^2) * 100, 1), "% | PC2 =",
      round(pca_result$sdev[2]^2 / sum(pca_result$sdev^2) * 100, 1), "%"
    ),
    x = "PC1", y = "PC2"
  ) +
  theme_cvis()

p_pca_clusters <- p_base_pca +
  geom_point(aes(color = cluster, shape = SPECIES_NAME), size = 3, alpha = 0.7) +
  scale_color_brewer(palette = "Set1", name = "Risk Cluster") +
  labs(title = "CU Risk Profile PCA - Clusters")

print(p_pca_clusters)
```

#### Red Flag Frequency
```{r red_flag_freq, fig.width=10, fig.height=7, dpi=300, out.width="100%"}
p_drivers <- ggplot(
  red_flag_summary %>% head(15),
  aes(x = reorder(name, n_red_flags), y = n_red_flags, fill = category)
) +
  geom_col() +
  geom_text(aes(label = paste0(round(pct_red_flag * 100), "%")),
    hjust = -0.1, size = 3, fontface = "bold"
  ) +
  coord_flip(clip = "off") +
  scale_fill_scico_d(palette = "roma", name = "Category") +
  labs(
    title = "Red Flag Frequency by Indicator",
    subtitle = paste0("Number of CUs where indicator exceeds red flag threshold (score >= 0.66)"),
    x = NULL, y = "Count of CUs with Red Flag"
  ) +
  theme_cvis() +
  theme(plot.margin = margin(5, 40, 5, 5))

print(p_drivers)
```


### Indicator Variation Propagation {.tabset .tabset-pills}

```{r indicator_variation_plots, fig.width=14, fig.height=10, dpi=300, out.width="100%"}
rcp_ref <- "45"
period_ref <- "0"

baseline_vals <- all_std_long %>%
  filter(
    rcp == rcp_ref, period_code == period_ref, gcm == "9",
    dsmodel %in% dsmodel_baseline
  ) %>%
  group_by(indicator) %>%
  summarise(baseline_mean = mean(std_value, na.rm = TRUE), .groups = "drop")

future_dat <- all_std_long %>%
  filter(
    gcm == "9", period_code != "0",
    dsmodel %in% dsmodel_baseline
  )

future_agg <- future_dat %>%
  group_by(indicator, category, rcp, period_code) %>%
  summarise(
    mean = mean(std_value, na.rm = TRUE),
    qlowgcm = quantile(std_value, qlgcm, na.rm = TRUE),
    qhighgcm = quantile(std_value, qhgcm, na.rm = TRUE),
    .groups = "drop"
  )

plot_dat <- future_agg %>%
  left_join(baseline_vals, by = "indicator") %>%
  mutate(
    scenario_label = paste0("RCP ", rcp, " P", period_code),
    direction = if_else(mean > baseline_mean, "High Risk", "Low Risk")
  ) %>%
  filter(!is.na(qlowgcm) | !is.na(qhighgcm))

categories_to_plot <- unique(plot_dat$category)

colors_lolli <- c(
  "GCM Range" = "grey85",
  "RCP 45" = "#33a02c",
  "RCP 85" = "#e31a1c"
)

for (cat in categories_to_plot) {
  cat_dat <- plot_dat %>% filter(category == cat)
  cat_desc <- cat_label_map[cat]
  if (is.na(cat_desc) || is.null(cat_desc)) cat_desc <- cat

  p <- ggplot(cat_dat, aes(y = scenario_label)) +
    geom_segment(aes(x = qlowgcm, xend = qhighgcm, yend = scenario_label),
      color = colors_lolli["GCM Range"], linewidth = 6, alpha = 0.6
    ) +
    geom_vline(aes(xintercept = baseline_mean), linetype = "dashed", color = "grey40") +
    geom_segment(aes(x = baseline_mean, xend = mean, yend = scenario_label, color = rcp),
      linewidth = 1.2
    ) +
    geom_point(aes(x = mean, fill = rcp), shape = 21, size = 4, color = "white", stroke = 1) +
    facet_wrap(~indicator, scales = "free_x", ncol = 3) +
    scale_color_manual(values = c("45" = colors_lolli["RCP 45"], "85" = colors_lolli["RCP 85"])) +
    scale_fill_manual(values = c("45" = colors_lolli["RCP 45"], "85" = colors_lolli["RCP 85"])) +
    labs(
      title = paste0("Indicator Uncertainty Propagation: ", cat_desc),
      subtitle = "Lollipops: Change from baseline mean; Grey bars: 10th-90th GCM Range",
      x = "Raw Indicator Value (Physical Units)",
      y = "Future Scenario",
      fill = "Scenario (RCP)", color = "Scenario (RCP)"
    ) +
    theme_cvis() +
    theme(
      strip.text = element_text(face = "bold"),
      legend.position = "bottom",
      panel.grid.major.y = element_line(color = "grey95")
    )

  cat("#### ", cat_desc, "\n\n", sep = "")
  print(p)
  cat("\n\n")
}
```

# Appendix A - Description of Indicators

### A.1 Projected August stream temperature (tw8proj) {#a1.-projected-august-stream-temperature-tw8proj}

**Metric:** The average of August stream temperatures across all years of the projection period (2041-2060 or 2081-2100) over all accessible streams with modelled habitat potential contained in the conservation unit (CU) boundary. Streams are weighted by stream segment length when calculating the average value. Projections from representative concentration pathways (RCP) 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

**Explanation:** August mean stream temperature is widely used as an index of stream thermal regimes and is strongly correlated with other temperature metrics, such as average spring and fall temperatures [@isaak2018; @iacarella2024; @isaak2023]. Temperature data for August is most widely available compared to other times of the year, meaning that these models have the most empirical support compared to other time periods [@isaak2017; @weller2024]. This time of year is also when stream temperatures peak, therefore representing the most thermally stressful period for many aquatic species.

For Pacific salmon, high stream temperatures in August may impact rearing juveniles present in streams, particularly Chinook and Coho salmon with a yearling life history, through both acute mortality and indirect stress. Although modelled temperatures rarely exceed physiological limits for juvenile survival [@iacarella2024], elevated temperatures can reduce available habitat, limit growth, productivity and refuge opportunities, and impair function such as aerobic scope [mantua2010; mayer2023]. For other salmon species and life histories, August stream temperatures may not be directly encountered during rearing, but are correlated with high temperatures during adjacent time periods. Therefore, the indicator provides a metric of thermal stress that may apply for rearing and spawning in other months. The negative consequences of high temperatures for these life stages are varied but well-established. For example, elevated temperatures encountered by spawners in late summer and early fall may cause delayed spawning [@mccullough2001], higher pre-spawn mortality [@bowerman2018], increased susceptibility to pathogens, and reduced spawning success [@richter2005].

**Information Source:** Projected temperatures are taken from stream thermalscapes scenarios, a statistical stream temperature model fitted to 680,000 km of the BC stream network at approximately 400m intervals [@weller2024]. The model produces temperature predictions averaged over 20-year periods from 1981-2100 using covariates that included catchment area, elevation, glacial cover, air temperature, and precipitation fitted to August mean temperatures from 562 stations in BC. Temporal covariates were taken from six CMIP5-era global climate models. Final model performance was assessed using 10-fold cross-validation and resulted in an R^2^ of 0.79 and RMSE of 1.53 C. Streams with catchment areas below 1km2 were not modelled, however, these represent only a small fraction of accessible streams with modelled habitat potential for Pacific salmon.

**Standardization Method:** Projected temperatures are standardized from 0-1 using a scaled exponential normalization increasing from a lower threshold of 15°C according to the formula:

$$z_{i} = \frac{e^{\lambda \frac{x_{i} - Q_{0.025}}{Q_{0.975} - Q_{0.025}}} - e^{\lambda}}{e^{\lambda} - e^{-\lambda}}$$

Where $z_{i}$ is the normalized value, $x_{i}$ is the original value, $Q_{0.975}$ is the 97.5^th^ quantile value, $Q_{0.025}$ is the 2.5^th^ quantile value, and $\lambda$ is a scaling parameter used to control the curvature. For this indicator $Q_{0.025}$ was set to 15, and $\lambda$ was set to 3. An exponential function was used to reflect the expectation of non-linear scaling in risk as temperatures approach higher values that are less suitable for Pacific salmon.

Using this function, any CUs where the average projected August mean temperature was below 15°C had a standardized risk value of 0, representing the expectation of relatively low risk due to projected stream temperatures. The lower threshold of 15°C was selected in consideration of other literature reviews, studies, and water quality guidelines for Pacific salmon. Guidelines and reviews from the US Pacific Northwest and British Columbia suggest optimal temperatures for spawning, rearing, and egg incubation below 16°C [@oliver2001; @u.s.environmentalprotectionagency2003]. These guidelines are based on a weekly 7-day average of daily maximum (7DADM) temperatures and maximum weekly average temperatures (MWAT), which could easily exceed the August mean temperature during hot periods. Therefore a lower threshold of 15°C was used for the August mean temperature.

**Limitations:** The indicator's relevance varies across salmon species and life histories. While August temperatures are directly experienced by stream-type Chinook and Coho salmon rearing as juveniles, they serve primarily as an index of thermal stress during adjacent time periods for other species. In addition, most CUs of sockeye salmon in the Fraser basin rear in lakes, whose temperatures are not modelled. Habitat suitability for rearing sockeye salmon would be better represented by physical and ecological models for lakes that consider factors such as stratification, trophic state, and oxygenation.

The stream temperature model relies on statistical relationships between air temperature, precipitation and observed stream temperatures up to the present, but predictions are more uncertain when extrapolating to temperatures beyond those observed. Broad-scale models may not capture fine-scale thermal heterogeneity important for salmon behavior, and the monthly averaging obscures daily temperature fluctuations and extreme events that can cause acute mortality [@iacarella2024]. The 15°C standardization threshold represents a generalized guideline that may not reflect variation among species, populations, and life stages in thermal optima.

### A.2 Rate of increase in August stream temperature (tw8rate) {#a.2-rate-of-increase-in-august-stream-temperature-tw8rate}

**Metric:** The decadal rate of increase in August mean stream temperature from a baseline period of 1981-2000 to the projection period (2041-2060 or 2081-2100) averaged over all accessible streams with modelled habitat potential contained in the CU boundary. Streams are weighted by stream segment length when calculating the average value. Projections from RCP 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

**Explanation**: The rate of climatic warming is a fundamental component for evaluating species exposure to climate changes and is closely related to metrics such as climate velocity and time of climate emergence [@sunday2012; @brito-morales2018; @boyce2022]. Faster rates of warming are more likely to outpace a species’ ability to adapt or migrate, allow less time until exceedance of thermal thresholds or safety margins, and cause greater ecosystem disruption [@comte2017; @morley2019; @boyce2022]. August stream temperatures are a widely used index of stream thermal regime and the rate of temperature increase during this period indicate potential changes to the thermal regime and increases in thermal stress.

For Pacific salmon, the thermal regime within a watershed influences spawning and emergence timing, metabolic rates and growth, hydrology, and ecosystem composition. Warmer temperatures reduce incubation times for eggs, leading to earlier emergence and potential mismatches with primary productivity and optimal food availability, which may not shift at the same pace or extent [@healey2011; @weinheimer2017]. Higher temperatures may increase juvenile growth and productivity in streams and lakes, particularly for populations at higher latitudes and elevations [@schindler2005; @gallagher2022], but this will only favour populations where juveniles are able to adapt with higher growth rates and is contingent upon food availability matching with phenology [@crozier2008]. Greater warming may also favour the invasion of nonnative predators and competitors, such as smallmouth bass, which are predicted to expand with warmer stream temperatures [@lawrence2014].

The life history and phenology of each CU is adapted to historical conditions and most likely to correspond with the optimal fitness timing for each life-history event [@crozier2008]. As stream temperatures warm, the current equilibrium may be disrupted favouring shifts in timing that may be complementary or divergent across life history events. Ultimately, faster rates of warming limit the amount of time for evolutionary responses, increasing the likelihood of negative consequences to survival.

**Information Source:** Projected and baseline temperatures are taken from stream thermalscapes scenarios [@weller2024], as described previously for the projected august stream temperature indicator.

**Standardization Method:** The rate of increase in August stream temperature is standardized using a linear normalization function according to the formula:

$$z_{i} = \frac{x_{i} - Q_{0.025}}{Q_{0.975} - Q_{0.025}}$$

Where $z_{i}$ is the normalized value, $x_{i}$ is the original value, $Q_{0.975}$ is the 97.5^th^ quantile value, and $Q_{0.025}$ is the 2.5^th^ quantile value of the original data range. The 95% quantile range was used instead of minimum and maximum values to reduce the influence of outliers when normalizing. 

**Limitations:**

### A.3 Proportional decrease in August flows (lowQpdelta) {#a.3-proportional-decrease-in-august-flows-lowqpdelta}

**Metric:** The change in average August flow from a baseline period of 1981-2010 to the projection period (2041-2060 or 2081-2100), expressed as a proportion of the average historical flow and averaged over all accessible streams with modelled habitat potential contained in the CU boundary. Streams are weighted by stream segment length when calculating the average value. Projections from RCP 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

**Explanation:** Pacific salmon require adequate flow during their freshwater life history to migrate and access habitats, provide foraging and refuge opportunities, and sustain water quality. Reduced summer flows and drought may affect Pacific salmon through a variety of mechanisms such as reducing the amount of useable habitat and connectivity among habitats [@tonina2022], exacerbating thermal stress during high temperature periods [@vehanen2023], reducing growth by decreasing food availability [@obedzinski2018], reducing water quality and lowering dissolved oxygen concentration, and causing mortality due to stranding [@warren2015]. Although empirical relationships between low flows and salmon survival are not widely established compared to relationships with temperature, decreases in summer flows have been associated with reduced productivity for multiple populations of stream-type Chinook [@crozier2006; @warkentin2022; @connor2003] and reduced coho salmon survival [@obedzinski2018]. We focus on August flows as it represents one of the lowest flow months for many streams within BC and the month with the strongest drying trend over the last half century [@ruzzante2025]. The proportional change in August flows provides an indication of the future drying trend for streams within a CU boundary.

**Information Source:** Model outputs for this indicator are based on the models used by Iacarella et al. 2025 in development of geospatial indicators of stream cumulative threats within the Fraser River basin. Flows were first derived from the Pacific Climate Impacts Consortium (PCIC) Variable Infiltration Capacity with a glacial component (VIC-GL) model, a process-based macroscale hydrology model that projects daily stream temperature and runoff on a gridded resolution of 1/16° (approx. 30 km2) within gauged watersheds of BC [@schnorbus2022]. The VIC-GL model simulates the full water and energy balance including snowpack accumulation and melt, evapotranspiration, soil moisture dynamics, and a module that accounts for glacier retreat under warming temperatures. The gridded model predictions were then downscaled and mapped on to the FWA stream network at a monthly resolution using a model that included the VIC-GL runoff values in combination with average precipitation values for each watershed unit within the Fraser River basin [@iacarella2025].

**Standardization Method:** The standardized value for the proportional decrease in August flows is obtained using an exponential decay normalization according to the formula:

$$z_{i} = e^{-\lambda \frac{x_{i} - Q_{0.025}}{Q_{0.975} - Q_{0.025}}} - e^{-\lambda} \frac{x_{i} - Q_{0.025}}{Q_{0.975} - Q_{0.025}}$$

Where $z_{i}$ is the normalized value, $x_{i}$ is the original value, $Q_{0.975}$ is the 97.5^th^ quantile value, $Q_{0.025}$ is the 2.5^th^ quantile value, and $\lambda$ is a scaling parameter used to control the curvature. For this indicator $\lambda$ was set to 3, and $Q_{0.975}$ was set to 0 so that risk increased according to the decay function for any changes in flow below 0. An exponential normalization reflects the expectation that vulnerability scaled non-linearly with decreases in August flows, with watersheds that experience the greatest decreases in flow more exposed to drought risk and negative impacts.

**Limitations:** The indicator's interpretation varies across salmon species and their diverse life histories. Stream-type Chinook salmon that rear in freshwater for one or more years are highly sensitive to summer low flows, which limit available rearing habitat and foraging opportunities during critical growth periods [@connor2003; @crozier2006]. Coho salmon similarly depend on adequate summer flows to maintain pool habitat and connectivity to off-channel refugia [@obedzinski2018]. Other species and life histories may be affected if low flows prevent or impede access to spawning grounds, or lead to reduced spawning success or habitat availability, however, there is less empirical evidence of a relationship between low flows and productivity.

The VIC-GL hydrological model's performance in simulating low flows is generally weaker than for annual discharge, as macroscale models often struggle to accurately represent groundwater contributions and small-scale hydrological features that maintain summer flows [@wenger2010]. The PCIC VIC-GL model was parameterized using static land cover and forest cover based on observations from 2010. However, logging and wildfires have reduced forest cover since then, and further changes in forest cover and composition will occur over the coming decades. The model does not account for future changes in water withdrawals, land use, or forestry practices that could further alter flow regimes.

### A.4 Proportional increase in November to January flows (highQpdelta) {#a.4-proportional-increase-in-november-to-january-flows-highqpdelta}

**Metric:** The change in average flow for the months of November, December and January from a baseline period of 1981-2010 to the projection period (2041-2060 or 2081-2100), expressed as a proportion of the average historical flow and averaged over all accessible streams with modelled habitat potential contained in the CU boundary. Streams are weighted by stream segment length when calculating the average value. Projections from RCP 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

**Explanation:** Elevated flows during the November to January period represents the risk of increased redd scour, a well-known cause of recruitment failure for salmon populations. High flows can mobilize streambed substrates, physically dislodging eggs from redds or burying them under fine sediments [@goode2013; @montgomery1996]. Empirical studies have linked high flow events to reduced productivity in certain populations [@weinheimer2017], including Chinook salmon in the Nicola River within the Fraser basin [@warkentin2022]

Climate change is expected to increase both the magnitude and frequency of extreme precipitation events in coastal BC, creating the potential for more frequent scour events. Models for other regions show potentially severe effects on survival and recruitment under future high flow scenarios [@battin2007; @shanley2014], although impacts will depend on channel morphology, stream gradient, substrate composition, and the rate of changes in flow [@goode2013; @montgomery1996].

**Information Source:** Model outputs for this indicator are based on the same approach and models used by Iacarella et al. 2025 and described earlier for the August flow indicator.

Standardization Method: The standardized value for the proportional increase in November to January flows is obtained using a scaled exponential normalization according to the formula: $$z_{i} = \frac{e^{\lambda \frac{x_{i} - Q_{0.025}}{Q_{0.975} - Q_{0.025}}} - e^{\lambda}}{e^{\lambda} - e^{-\lambda}}$$

Where $z_{i}$ is the normalized value, $x_{i}$ is the original value, $Q_{0.975}$ is the 97.5^th^ quantile value, $Q_{0.025}$ is the 2.5^th^ quantile value, and $\lambda$ is a scaling parameter used to control the curvature. For this indicator $Q_{0.025}$ was set to 0, and $\lambda$ was set to 3. The standardization reflects the expectation that scour risk increases non-linearly with flow magnitude, as greater increases are more likely to exceed critical bed mobilization thresholds and cause substantial egg mortality [@goode2013].

**Limitations:** The indicator is not based on an estimation of frequency of high flow events, but rather the increase in mean flows during November to January period. The monthly resolution of the modelled flows may not capture changes in peak flow events that are most likely to cause redd scour. Channel morphology and the condition of riparian and upslope vegetation also play important roles in moderating responses to extreme flow events that are not represented by this indicator. The downscaling of GCM precipitation, particularly for extreme events, introduces additional uncertainty as GCMs have limited skill in reproducing the intensity and spatial patterns of atmospheric rivers and other precipitation extremes that dominate winter flooding in coastal BC [@werner2016].

There is less empirical evidence on the relationship between higher winter stream flow and salmon productivity compared to indicators for stream temperature or summer flows, particularly for salmon populations within British Columbia. The time period during which high flows may be more likely to cause redd scour and increase mortality may vary by watershed and peak flow periods generally occur during the spring months. However, we chose to focus on winter flows as egg incubation for almost all CUs occurs during this time.

### A.5 Ecological Niche Model change in habitat favourability (favchange) {#a.5-ecological-niche-model-change-in-habitat-favourability-favchange}

**Metric:** The change in habitat favourability from the baseline period (1981-2000) to the projection period (2041-2060 or 2081-2100), calculated as the difference between future and baseline favourability values and averaged over all accessible streams within the CU boundary. Habitat favourability is derived from environmental niche models (ENMs) fitted to salmon occurrence data and environmental predictors, providing a metric of environmental suitability that ranges from 0 to 1. The future favourability values use an ensemble mean of six GCMs from RCP 4.5 and 8.5 scenarios for the environmental predictors.

**Explanation:** ENMs quantify the relationship between species distributions and environmental conditions, projecting how climate change may alter habitat suitability across landscapes and over time [@pacifici2015; @foden2019]. ENMs can identify regions where climate change is expected to improve habitat conditions versus areas where deteriorating conditions may reduce population viability. By comparing current and future habitat favourability for spawning and natal rearing, this indicator captures the net effect of multiple environmental changes on these life stages, accounting for shifts in August stream temperature and annual precipitation.

**Information Source:** Favourability predictions were applied from ENMs for the spawning ranges of the five Pacific salmon species across BC, as described in Iacarella and Weller 2023. Environmental predictor variables included August stream temperature, annual precipitation, distance to ocean, and proximity to lakes (for sockeye only). Models were fitted separately for each species and a generalized boosted model was selected as the best model following cross-validation among five different model types. A baseline period of 1981-2020 was used for all model fitting, with projected August stream temperature and precipitation from 2021-2100 derived over 20 year time periods from six different GCMs.

**Standardization Method:** Changes in habitat favourability are standardized using an inverted linear normalization according to the formula: $$z_{i} = 1 - \frac{x_{i} - Q_{0.025}}{Q_{0.975} - Q_{0.025}}$$

Where $z_{i}$ is the normalized value, $x_{i}$ is the original value, $Q_{0.975}$ is the 97.5^th^ quantile value, and $Q_{0.025}$ is the 2.5^th^ quantile value of the original data range. For this indicator $Q_{0.975}$ was set to 0 so that any positive changes in favourability were assigned a risk score of 0.

**Limitations:** Projections to future conditions may require extrapolation into non-analog climates—combinations of environmental variables not present during the training period. The models do not account for potential evolutionary adaptation or rapid phenotypic plasticity that could maintain population persistence despite environmental change, nor do they incorporate demographic processes (dispersal, metapopulation dynamics) or biotic interactions (predation, competition, disease) that mediate climate impacts. ENM predictions should be interpreted within the context that (1) habitat occupancy is limited by other factors, (2) salmon must be able to migrate to and from these areas, and (3) extreme events from climate change are not captured by these models [@iacarella2023].

### A.6 Cumulative threats to stream habitat (ct) {#a.6-cumulative-threats-to-stream-habitat-ct}

**Metric:** A composite index of threats from human activity to stream habitat, calculated using a normalized additive model that combines nine threat categories including flow alteration, in-stream habitat destruction, riparian disturbance, nutrients, pollution, and sedimentation. The indicator is calculated as the mean cumulative threat score across all accessible streams with modelled habitat potential in the CU boundary, weighted by stream segment length.

**Explanation:** Climate change does not occur in isolation but rather compounds existing anthropogenic stressors on salmon habitat, creating cumulative impacts that can push populations beyond thresholds for persistence. Habitat status is a commonly included factor in climate change vulnerability assessments and often associated with species sensitivity to climate change [@wade2013; @foden2019; @crozier2019]. Stream habitats that are already in a degraded state have reduced resilience to absorb climate-driven changes in temperature, flow, and other properties [@beechie2013]. The cumulative threats score for spawning and rearing streams within CU boundaries provides a metric of pre-existing watershed impairment that contextualizes sensitivity to environmental changes represented by other indicators.

**Information Source:** Cumulative threat scores were adapted from those described by Iacarella et al. 2025 and incorporate information from multiple federal and provincial databases with spatial coverage across all streams within the Fraser River basin. For some threat categories, particularly longitudinal fragmentation (passage barriers), the cumulative threat model uses distinct methods for anadromous versus resident species but only the anadromous species threat score was included in the indicator. All threat categories were included in the indicator score, recognizing that while some categories may be more important than others in determining habitat sensitivity to climate change all of them may contribute to some extent.

**Standardization Method:** The cumulative threat score is standardized using a linear normalization according to the formula: $$z_{i} = \frac{x_{i} - Q_{0.025}}{Q_{0.975} - Q_{0.025}}$$

Where $z_{i}$ is the normalized value, $x_{i}$ is the original value, $Q_{0.975}$ is the 97.5^th^ quantile value, and $Q_{0.025}$ is the 2.5^th^ quantile value of the original data range. A simple linear function was adopted because the threat score already represents an additive index where each category has been standardized across the range of values within the Fraser River basin. The 2.5th and 97.5th percentile values were used instead of the minimum and maximum values to avoid undue influence from CUs with extreme values.

**Limitations:** The cumulative threats index assumes that stressors to stream habitats combine linearly and independently, and that this additive score represents stream sensitivity to environmental changes. In reality, the relationship between anthropogenic stressors on stream habitat and the ability of that habitat to support salmon and remain resilient to climate change is complex and non-linear. Other limitations and assumptions of the cumulative threats model are described in Iacarella et al. 2025.

The indicator's interpretation requires consideration of species-specific threat sensitivities and life history vulnerabilities. Stream-type Chinook and Coho salmon that rear in tributaries for over a year are directly exposed to the full range of instream and riparian threats, making cumulative threat scores highly relevant to these life histories, while other species and life histories may only be vulnerable to a subset of threat categories. The threat categories only represent streams and therefore do not quantify the condition of lake rearing habitats for sockeye. The CU-level averaging of threat scores obscures within-CU variation in habitat quality, which may be critical for persistence if high-quality refugia enable subpopulation persistence despite widespread habitat degradation.

### A.7 Freshwater residency time (fwres) {#a.7-freshwater-residency-time-fwres}

**Metric:** The duration in days that juvenile salmon reside in freshwater from the start of egg incubation to ocean entry. For each CU, residency time is calculated using as the number of days from peak spawn timing to peak ocean entry timing, plus the typical number of years spent in freshwater (0 or 1).

**Explanation:** Freshwater residency time is an integrative life history trait that determines the cumulative exposure to freshwater climate stressors (temperature, flow, habitat availability). While the climate vulnerability implications of residency duration are complex and context-dependent, there is ample evidence that CUs with longer freshwater residency are more sensitive to environmental changes in freshwater habitats. The freshwater residency time indicator provides a simple metric to differentiate life history types among CUs and assess sensitivity to future changes based on their temporal exposure within freshwater.

**Information Source:** Freshwater residency time is derived from a life history timing data set for Pacific salmon CUs across their Canadian range, as described in Wilson and Peacock 2025. The repository incorporates multiple data sources to characterize timing by life stage across regions and species. For spawn timing, the estimates are based primarily upon New Spawning Escapement Database (NuSEDS) data collected by aerial surveys, stream walks, and weir counts. For ocean entry timing, estimates use various collection methods including incline plane traps, rotary screw traps, weirs, and fyke nets, though most estimates for Fraser CUs are based on the lower Fraser river incline plane and rotary screw traps. Data quality estimates are provided for each CU representing the recentness and number of sites or observations for each CU.

**Standardization Method:** The freshwater residency indicator is scored using a step-function that divides residency days into three categories: immediate migrants with residency times less than 200 days, sub-yearlings with residency times greater than 200 days but less than 300 days, and yearlings with residency times greater than 300 days. Immediate migrants, which are restricted to pink and chum salmon CUs, are assigned a standardized value of 0, while sub-yearlings and yearlings were assigned values of 0.5 and 1, respectively.

**Limitations:** Data quality varies by life stage and species, with estimates of ocean entry timing generally being more uncertain than spawn timing, and estimates for most sockeye CUs being higher quality than other species [@wilson2025]. Populations with variation in freshwater residency times and life history stategies may not be adequately represented by a single residency metric, but diversity in life history strategies may be an important factor in determining adaptive capacity. The freshwater residency indicator assumes that residency duration remains relatively constant under climate change, yet there is substantial evidence that the timing of life history transitions within freshwater exhibit phenotypic plasticity and may respond adaptively to changing conditions [@crozier2008].

Many factors mediate the sensitivity of Pacific salmon to environmental change within freshwater that are not represented by a simple metric of residency time. Disentangling the sensitivity of populations to different factors is difficult without information on survival by life stage and life cycle models [@crozier2021]. We assume a monotonic relationship between freshwater residency time and climate vulnerability, however, "optimal" residency depends on the relative severity of freshwater versus marine stressors, which varies across areas and may shift as climate change progresses. Freshwater habitats that are resilient to climate changes or increase in habitat favourability with warming could serve as refuges that sustain local salmon productivity and favour life histories with longer residency times, particularly if marine conditions are poor.

### A.8 Projected temperature during upstream migration (migrT) {#a.8-projected-temperature-during-upstream-migration-migrt}

**Metric:** The average of water temperatures during the upstream migration across all years of the projection period (2041-2060 or 2081-2100) estimated along the migration path from river entry to spawning sites using an upstream migration model. The model uses a moving migration window to represent the thermal exposure of CUs starting at the river mouth on the run timing start date and terminating at spawning sites on the peak spawning date. Projections from representative concentration pathways (RCP) 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

**Explanation:** Higher temperatures can negatively affect salmon during their upstream migration through a variety of mechanisms including acute mortality due to exceedance of critical thermal limits, creation of thermal barriers that delay spawning, greater disease susceptibility, higher pre-spawn mortality, and reduced aerobic scope [@mccullough2001; @bowerman2018; @martins2011; @mayer2023]. Within the Fraser basin, where some populations migrate over 1000 kilometres upstream, mortality and reduced spawning success from higher river temperatures have been extensively studied and modelled [@crossin2004; @farrell2008; @eliason2011], and are used annually by fisheries managers to adjust escapement targets for Fraser sockeye in order to compensate for increased mortality during high temperature conditions [@patterson2016]. Mortality is generally not equal across sexes, with observed mortality two times higher for females than males [@hinch2021]. The average projected migration temperature provides an indication of the indirect and direct negative consequences of future temperatures and exceedance of thermal limits for salmon CUs.

**Information Source:** Projected water temperatures were taken from gridded outputs of the PCIC VIC-GL model. The VIC-GL model generates projections of water temperature and flow for each day of the year from 1951-2100, which were averaged over the projection periods. The model grid was intersected with the upstream migration path within the stream network to subset cells when averaging temperature. River entry and spawn timing for each CU were taken from the same life history timing data set used to assess freshwater residency time [@wilson2025] and used to determine start and end dates for the upstream migration model. The location of spawning sites, representing the terminal ends of migration paths, were taken from NuSEDS observations for each CU. Sites with old and intermittent observations were removed by filtering sites with less than 5 observations and where the most recent observation was older than 2010.

**Standardization Method:** Projected temperatures are standardized from 0-1 using a scaled exponential normalization increasing from a lower threshold of 15°C according to the formula: $$z_{i} = \frac{e^{\lambda \frac{x_{i} - Q_{0.025}}{Q_{0.975} - Q_{0.025}}} - e^{\lambda}}{e^{\lambda} - e^{-\lambda}}$$

Where $z_{i}$ is the normalized value, $x_{i}$ is the original value, $Q_{0.975}$ is the 97.5^th^ quantile value, $Q_{0.025}$ is the 2.5^th^ quantile value, and $\lambda$ is a scaling parameter used to control the curvature. For this indicator $Q_{0.025}$ was set to 15, and $\lambda$ was set to 3. For this indicator $Q_{0.025}$ was set to 15, and $\lambda$ was set to 3.

Any CUs where the projected temperature during upstream migration was below 15°C had a standardized risk value of 0, representing the expectation of relatively low risk. The lower threshold of 15°C corresponds with the optimal temperature for aerobic scope of Fraser sockeye [@eliason2011], and is a lower limit beyond which increases in mortality associated with temperature may be observed [@martins2011; @atlas2021]. There is limited evidence of negative temperature impacts on aerobic scope or migration below 15°C, even for populations of coho salmon that typically experience much lower migration temperatures [@raby2016]. An exponential function was used for standardization reflecting the expectation that risks of mortality and sub-lethal impacts increase non-linearly with temperature.

**Limitations:** The upstream migration model assumes a uniform rate of migration from river mouth to spawning sites with the first arrival to spawning sites coinciding with spawn timing start. In reality, migration speeds may be variable and some CUs may utilize cold-water refuges or hold in lakes along their upstream migration. The indicator does not consider cold-water refuges, which can decrease exposure to sub-lethal thermal stress and be particularly important in maintaining the viability of summer-run timing phenotypes [@snyder2022]. The migration model uses a fixed river entry and spawn timing for CUs based on historical observations, but shifts in run and spawn timing can occur in response to natural and anthropogenic pressures [@kovach2012; @crozier2025].

### A.9 Proportional change in flow during upstream migration (migrQ) {#a.9-proportional-change-in-flow-during-upstream-migration-migrq}

**Metric:** The proportional change in flow during the upstream migration from the baseline period (1981-2000) to the projection period (2041-2060 or 2081-2100) estimated along the migration path from river entry to spawning sites using an upstream migration model. The model uses a moving migration window to represent the average flow experienced by CUs starting at the river mouth on the run timing start date and terminating at spawning sites on the peak spawning date. Projections from representative concentration pathways (RCP) 4.5 and 8.5 are included with an ensemble mean calculated from six global climate models.

**Explanation:** Both high and low flows can create barriers to salmon migrating to spawning grounds and impact their condition through direct effects on swimming energetics and interacting effects of flow and temperature. The effects of high and low flows vary based on the watershed type and on CU migration life history. In large river systems like the Fraser, high flows increase the energetic costs of swimming, slowing down the speed of migrating salmon and potentially creating velocity barriers in river reaches with constricted banks like the Fraser Canyon [@rand2006]. This pattern seems to be the opposite in smaller watersheds where lower flows are a particular challenge as they may create passage barriers that reduce access to spawning habitats, exacerbate effects of increased temperatures, and delay migration [@atlas2021].

**Information Source:** Projected water temperatures were taken from gridded outputs of the PCIC VIC-GL model. The VIC-GL model generates projections of water temperature and flow for each day of the year from 1951-2100, which were averaged over the projection periods. The model grid was intersected with the upstream migration path within the stream network to subset cells when averaging temperature.

**Standardization Method:** This indicator is not currently included in scoring and was standardized to a value of 0 across all CUs. All Fraser CUs were projected to experience a decrease in flow during their upstream migration, but this may bring both positive and negative consequences for migration success and survival.

### A.10 Upstream migration distance (migrdist) {#a.10-upstream-migration-distance-migrdist}

**Metric:** The distance of the migration path from river mouth to spawning sites. For CUs with multiple spawning locations, the average of the migration distance to all spawning sites is used.

**Explanation:** Migration distance represents a simple metric of the difficulty and energetic requirements for upstream migration, the most energetically demanding phase of the salmon life cycle. The energetic costs of migration scale with distance as adults cease feeding prior to freshwater entry and rely on stored lipids and proteins for all metabolic demands during migration, gonadal development, and spawning. Longer upstream migrations impose greater energetic demands, are more likely to expose populations to higher temperatures or thermal barriers, and require more specialized adaptation to overcome [@farrell2008; @eliason2011]. While populations that undergo difficult, longer migrations have adaptations that allow them to survive and spawn successfully, such as higher lipid stores, fewer eggs, and smaller, more fusiform morphologies [@crossin2004], such adaptations reflect evolutionary trade-offs whose consequences may become more severe with further warming [@crozier2008].\

**Information Source:** The location of spawning sites, representing the terminal ends of migration paths, were taken from NuSEDS observations for each CU. Sites with old and intermittent observations were removed by filtering sites with less than 5 observations and where the most recent observation was older than 2010. Migration paths were created from the stream network within the Freshwater Atlas (FWA) based on downstream segment distances and watershed codes.

**Standardization Method:** Upstream migration distance is standardized using a linear normalization according to the formula: $$z_{i} = \frac{x_{i} - Q_{0.025}}{Q_{0.975} - Q_{0.025}}$$

Where $z_{i}$ is the normalized value, $x_{i}$ is the original value, $Q_{0.975}$ is the 97.5^th^ quantile value, and $Q_{0.025}$ is the 2.5^th^ quantile value of the original data range. A simple linear function was adopted reflecting the proportional scaling of energy requirements with migration distance. The 2.5th and 97.5th percentile values were used instead of the minimum and maximum values to avoid undue influence from CUs with extreme values.

**Limitations:** The indicator treats all kilometers equally, but migration path heterogeneity can impose highly variable energetic costs and create velocity or elevation barriers to migration (e.g. Hell’s Gate in the Fraser Canyon). Migration difficulty and energetic requirements depend on temperature and flow, which are assessed separately.

### A.11 Projected nearshore sea surface temperature (SSTproj) {#a.11-projected-nearshore-sea-surface-temperature-sstproj}

**Metric:** The projected sea surface temperature (SST) across all years of the projection period (2041-2060 or 2081-2100) throughout the marine adaptive zone (MAZ) for the CU over a 6-month period beginning with the month prior to ocean entry. Projections from RCP 4.5 and 8.5 are included with an ensemble mean calculated from 28 global climate models.

**Explanation:** Nearshore marine SST is a critical environmental variable for Pacific salmon that influences growth and survival by determining prey availability and quality, ecosystem composition, metabolic demands, and habitat selection. During warmer years, shifts in zooplankton composition to species with lower nutritional value are well documented within the southern range of Pacific salmon [@mackas2007; @fergusson2020]. Warmer SST may also increase competition and put Pacific salmon at a competitive disadvantage with other species such as juvenile sablefish [@daly2024]. Direct physiological effects from elevated temperatures in marine areas are more difficult to isolate and prove, but higher temperatures result in higher metabolic rates and energy requirements [@beauchamp2007; @daly2015], and influence depth and swimming behaviour [@ueno1992; @sabal2023].

At regional scales, SST trends and anomalies is one of the most important variables for explaining patterns of salmon productivity, with higher SSTs consistently associated with lower survival and productivity across populations and particularly those in southern BC and the continental United States [@mueter2002; @sharma2013; @kilduff2014]. For BC salmon, high SSTs from marine heatwaves, particularly the Blob, have been associated with poor productivity [@grant2019] and models have linked SST anomalies with reduced recruitment [@connors2020]. By representing the future SST that salmon will experience during their entry into the ocean, this indicator provides a metric of climate exposure that can be compared across regions and relative to the historical baseline. This metric shares properties with other indicators of climate risk used for marine species, such as thermal safety margins and time of climate emergence [@boyce2022].

**Information Source:** Ensemble mean and quantile ranges for SSTs were derived from 28 global climate models from the CMIP6 suite of models that were bias corrected to a higher resolution using the quantile delta mapping (QDM) method trained on the global Ocean Sea Surface Temperature and Ice Analysis (OSTIA) historical data set [@donlon2012].

MAZs used to determine SST spatial domains were adapted from those used to define salmon CUs for the Wild Salmon Policy [@Holtby2007]. These areas consist of twelve distinct ecoregions for salmon that characterize the nearshore adaptive marine environments for Pacific salmon. MAZ boundaries were slightly modified from those described in Holtby and Ciruna 2007 to correspond with boundaries from corresponding marine biogeographic regions [@dfo2009].

SSTs were compared to dynamically-downscaled, regional oceanographic model projections for validation. For the Georgia Strait MAZ, SSTs were compared to the SalishSeaCast model, developed to represent the complex and distinct oceanographic processes within the Salish Sea (Ridenour and Holdsword, in prep). For other coastal regions of BC, SSTs were compared to the BCCM model surface layer projections, with a mask layer used to exclude regions where conditions could not be adequately represented by the models [@peña2024]. Both models are dynamically downscaled from the CanESM2 global climate model from CMIP5.

**Standardization Method:** The negative relationship between SST and productivity was represented with a linear function that increased with any change in SST above historical averages. Because the indicator was averaged over a large area and time frame for each CU, the variation in responses was low and the resulting indicator was standardized across the range of values that included historical SSTs to projected SSTs. Thus, the lowest historical SST values represented a standardized score of 0, while highest future SST values represented a score of 1.

**Limitations:** The mechanisms relating nearshore SST to salmon survival are complex and mediated by ecosystem interactions, oceanography, and life history characteristics, which are not represented by this or other indicators. Projected SST represents a coarse proxy for these underlying dynamics and the underlying relationship with survival and productivity may exhibit non-stationary and non-linear responses over time.

Pacific salmon CUs were associated with marine conditions based on MAZs, a representation of the marine entry and nearshore rearing area for CUs. Within the Fraser Basin, all CUs share the Georgia Strait MAZ, however, nearshore life history strategies are diverse with some marine ecotypes spending longer and variable residency times within Georgia Strait, while others out-migrate further north relatively quickly. The exposure of CUs to changes in nearshore SST may therefore vary, and the consequences for vulnerability are not evaluated. Habitat use, distribution, and depth also vary across marine ecotypes and may determine exposure to warmer temperatures, hypoxic waters, and other negative conditions. Models of fine-scale marine distribution could provide a more detailed and realistic representation of exposure to increased SSTs.

### A.12 Rate of increase in nearshore sea surface temperatures (SSTrate) {#a.12-rate-of-increase-in-nearshore-sea-surface-temperatures-sstrate}

**Metric:** The decadal rate of increase in SST from a baseline period to the projection period (2041-2060 or 2081-2100) throughout the marine adaptive zone (MAZ) for the CU over a 6-month period beginning with the month prior to ocean entry. Projections from RCP 4.5 and 8.5 are included with an ensemble mean calculated from 28 global climate models.

**Explanation:** The rate of climatic warming is a fundamental component for evaluating species exposure to climate changes and is closely related to metrics such as climate velocity and time of climate emergence [@sunday2012; @brito-morales2018; @boyce2022]. Faster rates of warming are more likely to outpace a species’ ability to adapt or migrate, allow less time until exceedance of thermal thresholds or safety margins, and cause greater ecosystem disruption [@comte2017; @morley2019; @boyce2022]. The rate of increase in nearshore SST represents changes to the marine thermal regime during a period of critical growth and survival for salmon.

**Information Source:** Historical and projected SST values are taken from CMIP6 models bias corrected to a higher resolution using the QDM method as described for the projected nearshore SST indicator.

**Standardization Method:** The rate of increase in nearshore SST is standardized using an linear function with a range of 0 to 1 corresponding to the 95% quantile range (2.5th and 97.5th percentiles) of CU values.

**Limitations:** As described for the projected nearshore SST indicator, the mechanisms relating the rate of increase in SST to survival and climate change vulnerability are complex and not fully represented by this indicator.

### A.13 Cumulative impacts to nearshore marine habitat (CImpact) {#a.13-cumulative-impacts-to-nearshore-marine-habitat-cimpact}

**Metric:** The average cumulative impact score of nearshore habitats averaged throughout the MAZ for a CU. Cumulative impact scores are calculated for multiple habitat categories within the Pacific region including pelagic, benthic, eelgrass, and kelp forests. Stressors included as part of the additive cumulative impact model include aquaculture, fishing, invasive species, nutrient inputs, pollution inputs, and shipping [@murray2024].

**Explanation:** Estuarine and nearshore marine environments are important habitats for salmon as they complete smoltification and early marine growth. The cumulative impacts to nearshore marine habitats provides an index of stressors to these habitats that contextualizes sensitivity to environmental changes represented by other indicators. Marine habitats experiencing greater cumulative impacts are more vulnerable to tipping points and synergistic impacts from climate change [@halpern2008; @boyce2022].

**Information Source:** This indicator is derived from a cumulative impact model developed for Canadian marine ecosystems [@murray2024].

**Standardization Method:** The cumulative impacts score is standardized using a linear function with a range of 0 to 1 corresponding to the 95% quantile range (2.5th and 97.5th percentiles) of CU values. A simple linear function was adopted because the threat score already represents a standardized, additive index.

**Limitations:**

### A.14 Conservation status (CUstatus) {#a.14-conservation-status-custatus}

**Metric:** Recent Wild Salmon Policy (WSP) conservation status of a CU categorized as either Red, Amber, or Green. Status was assessed using the Rapid Status Assessment Approach that uses a Yes/No decision tree considering the number of recent spawners, trends in abundance, data type (absolute or relative abundance), and biological benchmarks [@pestal2023; @csas2024]. The input data, thresholds and appropriate metrics, and final status of Fraser CUs in 2022/2023 were evaluated and agreed upon by expert judgement throughout the Rapid Status process.

**Explanation:** Conservation status is an important metric of climate change sensitivity and adaptive capacity that is often incorporated in vulnerability assessments [@pacifici2015; @crozier2019; @foden2019; @boyce2022]. Status is a comprehensive metric that can represent multiple facets of vulnerability, such as stress already imposed on a species from factors other than climate change, rarity and demographic risk, as well as sensitivity to recent climatic changes. A poor conservation status often indicates that a species is experiencing stress from non-climatic anthropogenic stressors, reducing its resiliency to further changes and catastrophic events.

For Pacific salmon CUs, conservation status reflects the recent and historical influence of multiple stressors across marine and freshwater habitats, as well as the inherent size and extent of a CU. Smaller CUs with lower abundances are more likely to be assigned a red status because the assessment method uses a minimum abundance threshold of 1,500 spawners measured against the CUs generational geometric mean.

**Information Source:** WSP status evaluations of Fraser CUs were conducted in 2023 and 2024 using the Rapid Status Assessment Approach, which uses a decision tree interpreted and agreed upon by expert judgement. Data requirements include spawner abundances from NuSEDS or area-based Stock Assessment programs, stock-recruitment data where available, appropriate escapement sites and years selected by stock assessment experts based on data quality and quantity, and data treatments including gap filling as applicable. The appropriate application of abundance thresholds is assessed and relative-abundance benchmarks may be updated from prior integrated status assessments if updated data or methods are available. Confidence ratings of high, medium, or low are assigned as part of the assessment framework based on the decision tree node used to assign status (i.e. the method for assigning status based on data availability) , which is a corollary of confidence measures from previous integrated status processes. Higher confidence scores were assigned to status designations based on relative and absolute abundance indices rather than trend metrics. For further details on the WSP rapid status method see [@pestal2023; @csas2024].

**Standardization Method:** The status category was converted to a standardized score between 0 and 1 using a simple categorical function. A status of green was assigned a score of 0; amber status was scored as 0.5; and red status was scored as 1.

**Limitations:** The categorical status rating and standardization method are based primarily on generational averages or trends in spawner abundance and do not capture the full range of demographic characteristics that influence population condition and climate vulnerability. Other characteristics such as fecundity, the number and status of sites (populations), spawner size, and spawner age may be exhibiting different trends across CUs that are not evaluated as part of the conservation status indicator. Trends in spawners can only act as a proxy for how other biological process may affect productivity.

Data quality and confidence in the assessment score varies across CUs and is represented by a confidence rating assigned to each status assignment. A few CUs were data-deficient and not able to be assessed.

### A.15 Recent generational average of natural spawners (Cunmat) {#a.15-recent-generational-average-of-spawners-cunmat}

**Metric:** The geometric average of the number of natural-origin spawners (male and female) over the most recent generation with available estimates. Natural-origin spawners are those that are born in the wild but whose parents may or may not be of hatchery origin [@dfo2018]. For cyclic sockeye CUs where abundance is dominated by a single-year on the cycle line, the number of spawners from the most recent dominant year is used instead of the average over all years of a generation.

**Explanation:** Ecological theory and empirical evidence shows that larger, healthier populations possess demographic buffers allowing them to better withstand climate and anthropogenic pressures [@pacifici2015; @foden2019]. Smaller populations are more susceptible to stochastic events, are associated with lower genetic diversity, and have lower resilience to stressors and cumulative threats. The recent generational average of natural spawners represents the higher demographic risk for populations at lower abundances and is an important component of climate change vulnerability [@pearson2014; @wade2013].

Only natural-origin spawners were used for the generational average as they are more likely to contribute to population persistence and adaptive capacity compared to hatchery-origin spawners. Natural-origin spawners are also used for status assessments and thresholds for the WSP rapid status assessment method [@pestal2023].

**Information Source:** The generational average of spawners were taken from the WSP rapid status evaluations of Fraser CUs, as described for the conservation status indicator.

**Standardization Method:** The generational average of spawners was converted to a standardized score between 0 and 1 using an exponential decay function that increased inversely with abundance below a maximum threshold of 10,000 spawners. CUs with more than 10,000 spawners were assigned a score of 0, with the risk value increasing to the maximum score of 1 at 0 spawners. This relationship corresponds to the abundance thresholds used for WSP status assessments and the expectation of risk increasing non-linearly as abundance declines.

**Limitations:** This metric shares similarities with the conservation status indicator, but is focused specifically on recent abundances independent of any trends. For many CUs absolute abundance information that is representative of the entire CU is not available, therefore a relative index was used. Relative indices of abundance are adjusted using an expansion factor, but may nonetheless underestimate the absolute abundance due to potential error associated with the expansion method. In addition, some streams are not consistently surveyed and are not included when determining the relative index of abundance for a CU.

### A.16 Heterozygosity (hetzyg) {#a.16-heterozygosity-hetzyg}

**Metric:** Pacific salmon are diploid organisms with two copies of each nuclear chromosome in the genome. Heterozygosity is the proportion of genomic locations that contain a different allele (i.e., variant) on each copy of the chromosome within an individual. To calculate heterozygosity, genotypes are called for each genetic loci from sequencing data, and the number of loci with two variants is divided by the total number of loci in the dataset for each individual. Heterozygosity for a population is then calculated as the average across individuals.

**Explanation:** The impacts of climate change on organisms, in part, depend on the responses of the organisms to those changes. One mechanism through which organisms respond to environmental change is through adaptation. Adaptation as a result of directional selection requires selection to operate on variation that exists within the population. Climate change responses will need to occur over relative short timeframes in the context of evolution, and thus climate change adaptation is unlikely to rely primary on new mutations in a population. Consequently, the raw material that will allow adaptive responses to climate change is the genetic variation that exists in the population currently, which is known as standing genetic variation. Heterozygosity is a metric that provides an estimate of standing genetic variation, and lower heterozygosity within a population suggests lower capacity for adaptive responses [@tigano2023].

**Information Source:** The heterozygosity values presented in this assessment were determined from low-coverage whole-genome resequencing of Pacific salmon DNA. DNA was isolated from tissue samples, primarily fin, using standard methods and then sequenced on an Illumina NovaSeq X. Approximately 33M PE150 sequencing reads were generated per sample, which were then trimmed to remove adaptor content with fastp, and aligned to a reference genome corresponding to the relevant species with bwa. Alignments were filtered to remove PCR duplicates, and then genotype likelihoods at single-nucleotide polymorphisms (SNPs) were determined with angsd. Genotypes were imputed from the genotype likelihoods with beagle, and heterozygosity values were calculated by the proportion of heterozygous sites out of the total number of SNPs in the dataset.

**Standardization Method:** Heterozygosity is standardized using a linear normalization according to the formula:

$$z_{i} = \frac{x_{i} - Qs_{0.025}}{Qs_{0.975} - Qs_{0.025}}$$

Where $z_{i}$ is the normalized value, $x_{i}$ is the original value, $Qs_{0.975}$ is the 97.5^th^ quantile value, and $Qs_{0.025}$ is the 2.5^th^ quantile value of the original data range within the species only. Measures of heterozygosity are not comparable across species, therefore ranges used for standardization were restricted to observations within the same species. The 2.5th and 97.5th percentile values were used instead of the minimum and maximum values to avoid undue influence from CUs with extreme values.

**Limitations:** Heterozygosity is a common method to estimate standing genetic variation, but it relies upon called genotypes for individuals. For the datasets used here, the numbers of individuals sequenced were sufficient to support imputation of genotypes from the genotype likelihoods. However, this may not always be the case, particularly using low-coverage sequencing data. In such a case, an alternative metric of standing genetic variation should be employed (e.g., nucleotide diversity; [@layton2021]). Additionally, any estimate of standing genetic variation will be sensitive to the genomic loci that are used to generate the estimate. As a result, heterozygosity values are not directly comparable across different studies or species.

### A.17 Genomic offset (genoff) {#a.17-genomic-offset-genoff}

**Metric:** A genomic offset is an estimate of the mismatch between the genetic variation in a population of organisms under current conditions and the genetic variation that would be predicted to track a forecasted change in climate. To determine a genomic offset, first, genotype-environment associations are established using genomic data and environmental data across a species’ current range. Putatively adaptive loci are identified from these associations and then used to model genetic variation among populations as a function of environmental variation (i.e., allelic turnover in response to climate variation). Once this model is established, forecasted environmental conditions under climate change are used as input into the model to predict the genetic variation that would match forecasted climate conditions for a population. The Euclidean distance between a population’s current genetic variation and the predicted variation that would track climate change is the genomic offset.

**Explanation:** A genomic offset reflects how much genetic change would be required for a population to track climate change at a location under the assumption that adaptation will proceed in similar ways to the local environmental adaptation present in the species currently. These estimates provide insights on the relative capacity for adaptation to environmental change among populations as a higher offset suggests that more adaptive genetic change would be required for that population to respond to the impacts of climate change [@layton2021; @tigano2023].

**Information Source:** The genomic offsets presented in this assessment were determined from the same low-coverage whole-genome resequencing of Pacific salmon DNA as those used to determine heterozygosity. See the Heterozygosity section above for details on sequence generation and genotype calling. Genotypes were used to calculate population-specific allele frequencies at all SNPs in the dataset for each species, and genotype-environmental associations were determined by redundancy analysis [@capblancq2021] using the allele frequencies and environmental data obtained from the WorldClim database (four non-redundant bioclim variables: bio1, bio5, bio15 and bio16). Outlier SNPs consistent with involvement in environmental adaptation were determined by identifying SNPs with loadings on the first three RDA axes that exceeded three standard deviations from the mean of all SNP loadings on the axis. Allele frequencies for the outlier loci were then used as the response variables in a gradient forest model with all 19 bioclim variables as predictors. Genomic offsets were determined by the Euclidean distance between the model fit with current environmental conditions, and the model fit under forecasted conditions for CMIP6 SSP8.5 (no mitigation scenario).

**Standardization Method:** Genomic offset is standardized using a linear normalization according to the formula: $$z_{i} = \frac{x_{i} - Qs_{0.025}}{Qs_{0.975} - Qs_{0.025}}$$

Where $z_{i}$ is the normalized value, $x_{i}$ is the original value, and $Qs_{0.025}$ and $Qs_{0.975}$ represent the 2.5^th^ quantile and 97.5^th^ quantiles within the species' range of values. Measures of genomic offset are not comparable across species, therefore ranges used for standardization were restricted to observations within the same species. The 2.5th and 97.5th percentile values were used instead of the minimum and maximum values to avoid undue influence from CUs with extreme values.

**Limitations:** Genomic offsets are not a direct estimate of climate change vulnerability, and their predictive capability is highly contextual. For example, in some cases, environmental changes that result in higher genomic offsets can correspond to increases in population fitness, as adaptive states under current conditions may no longer be optimal under novel conditions, but may still experience an overall increase in fitness. Additionally, the gradient forest model fit assumes that current genetic variation and environmental conditions reflect adaptive states, and that future adaptation to climate conditions will reflect adaptative process that have produced extant patterns of variation. However, given what is known about environmental adaptation based on current variation, genomic offsets provide predictions on which populations of a species may require the most genetic change as a result of climate change, which can be considered as a component of estimating risk. Genomic offsets are relative measures that are dependent on the genomic loci considered in the analysis, and are not directly comparable across separate analyses, studies or species.

# References
