---
output:
  html_document:
    toc: false
    number-sections: false
    toc_float: true
    toc_collapsed: false
    toc_depth: 3
    theme: flatly
params:
  FULL_CU_IN: "CK-03"
  default_rcp: "45"
  default_period: "3"
title: CVIS Data Report - `r cu_run_i$CU_NAME` 
---

```{r custom theme, echo = FALSE, message = F, warning = F}
knitr::opts_chunk$set(message = F, echo = F, warning = F)
library(magrittr)
library(reactable)
library(htmltools)

```

```{r load data, echo=FALSE, message=FALSE, warning=FALSE}
# Load new scoring results and sensitivity data if not already in environment
if (!exists("all_std_long")) {
  load(file.path(here(), "output", "scoring_results.Rdata"))
}
if (!exists("fw_all")) {
  # Load stream indicator data (adjust filename if different)
  load(file.path(here(), "output", "fw_rearing_indicators.Rdata"))
}
if (!exists("overall_sensitivity")) {
  load(file.path(here(), "output", "sensitivity_analysis.Rdata"))
}
``` 


```{r subset CU}
#params$FULL_CU_IN <- "SEL-05-02"
cu_i <- params$FULL_CU_IN
cu_run_i <- cu_run[cu_run$FULL_CU_IN == params$FULL_CU_IN,]
sp_pick <- cu_run$SPECIES_NAME[cu_run$FULL_CU_IN == cu_i] #species abbr
sp_pick_bcfp <- spp_lookup$spp_abr_bcfp[spp_lookup$SPECIES_NAME == sp_pick] 
sp_pick_ENM  <- str_to_lower(sp_pick)[1]

#default RCP and period values
rcp_pick <- params$default_rcp
period_pick <- params$default_period

```


```{r spatial subset, echo = FALSE, message = F, warning = F}

#indicator dataframe
cu_std <- get_CU_indicators(all_std_long,
  cu_i = cu_i,
  use_standardized = FALSE,
  period_pick = period_pick,
  RCP_pick = rcp_pick)

# Stream indicator summary rows (fw_all format)
fw_cu <- ss_all %>%
  filter(FULL_CU_IN == cu_i) %>%
  distinct(FULL_CU_IN, .keep_all = TRUE)

stream_cu_sub <- stream_cu_picks[,colnames(stream_cu_picks) == cu_i]

cu_boundary_i <- cu_boundary[cu_boundary$FULL_CU_IN == cu_i,]

#subset nuseds observations
nuseds_cu <- nuseds_Fr[nuseds_Fr$FULL_CU_IN == cu_i,]

fw_sp_cu <- subset_fw_models(
  fw_models = fw_sp_ind,
  cu_i = cu_i,
  stream_cu_picks = stream_cu_picks,
  cu_run = cu_run,
  spp_lookup = spp_lookup,
  to_factor = TRUE
) %>%
  rename(favchange_45   = contains(paste0("favchange_", sp_pick_ENM, "_45_3")),
         favchange_85   = contains(paste0("favchange_", sp_pick_ENM, "_85_3")),
         fav_45   = contains(paste0("fav_", sp_pick_ENM, "_45_3")),
         fav_85   = contains(paste0("fav_", sp_pick_ENM, "_85_3")))

#make accessible only version for plotting some indicators
acc_sp_cu <- fw_sp_cu %>%
  filter(model_access_salmon %in% c("OBSERVED", "INFERRED"))

#subset in CU boundary
temp <- unlist(st_intersects(cu_boundary_i, watershed_flow))
watershed_flow_cu <- watershed_flow[temp,]

#subset in cu boundary
temp <- unlist(st_intersects(cu_boundary_i, stations_flow))
stations_cu <- stations_flow[temp,]


## Migration spatial layers
migr_cu <- migr_list[[cu_i]]


#Lakes within CU boundary (for plotting)
temp <- unlist(st_intersects(cu_boundary_i, lakes_Fr))
lakes_cu <- lakes_Fr[temp,]


## Timing data
cu_timing_long_i <- cu_timing_long[cu_timing_long$FULL_CU_IN == cu_i, ]
cu_timing_i <- cu_timing_Fr[cu_timing_Fr$FULL_CU_IN == cu_i, ]


#get long data frame of standardized indicator values for the CU
ind_cu <- get_CU_indicators(all_std_long,
  cu_i,
  indicators_choose = tbl_indicators$abbrev,
  RCP_pick = rcp_pick,
  period_pick = period_pick
) %>%
  dplyr::mutate(
    SPECIES_NAME = unique(all_std_long$SPECIES_NAME[all_std_long$FULL_CU_IN == cu_i])[1],
    CVIS_NAME = unique(all_std_long$CVIS_NAME[all_std_long$FULL_CU_IN == cu_i])[1]
  )


```


```{r helper function for subset tables}
# Function to create a reactable for a single indicator category
create_indicator_reactable <- function(df, indicator_codes) {
  # Filter to selected indicators
  tbl_view <- df[df[["Indicator Code"]] %in% indicator_codes, , drop = FALSE]

  if (nrow(tbl_view) == 0) {
    return(NULL)
  }

  # Precompute scaling for coloring
  diff_max <- tryCatch(
    {
      x <- tbl_view[["CU score"]] - tbl_view[["Species score"]]
      m <- max(abs(x), na.rm = TRUE)
      if (is.infinite(m) || is.na(m) || m == 0) 1 else m
    },
    error = function(e) 1)

  # Color function for CU score vs Species score
  cu_z_cell <- function(value, index, name) {
    sp <- tbl_view[["Species score"]][index]
    if (is.na(value) || is.na(sp)) return("—")
    diff <- value - sp
    intensity <- min(1, abs(diff) / diff_max)
    alpha <- 0.15 + 0.40 * intensity
    bg <- if (diff < 0) {
      sprintf("rgba(34, 197, 94, %.3f)", alpha)
    } else if (diff > 0) {
      sprintf("rgba(239, 68, 68, %.3f)", alpha)
    } else {
      "rgba(148, 163, 184, 0.20)"
    }
    div(
      style = list(
        backgroundColor = bg,
        borderRadius = "6px",
        padding = "1px 6px",
        display = "inline-block"
      ),
      sprintf("%.3f", value)
    )
  }

  # Create reactable
  reactable(
    tbl_view,
    filterable = FALSE,
    searchable = FALSE,
    highlight = TRUE,
    bordered = TRUE,
    striped = FALSE,
    compact = TRUE,
    pagination = FALSE,

    defaultColDef = colDef(
      minWidth = 70,
      headerVAlign = "center",
      vAlign = "center"
    ),

    theme = reactableTheme(
      borderColor = "#e5e7eb",
      stripedColor = "#f9fafb",
      highlightColor = "#eef2ff"
    ),

    columnGroups = list(
      colGroup(
        name = "Standardized Scores",
        columns = intersect(c("CU score", "Species score", "All Species score"), names(tbl_view))
      ),
      colGroup(
        name = "CU Raw Values",
        columns = intersect(c("Mean", "Q10", "Q90"), names(tbl_view))
      ),
      colGroup(
        name = "Raw Means",
        columns = intersect(c("Species Mean", "All Species Mean"), names(tbl_view))
      )
    ),

    columns = list(
      `Indicator Type` = colDef(width = 140),
      `Indicator Code` = colDef(width = 115),
      `CU score` = colDef(width = 95, format = colFormat(digits = 3), na = "—", cell = cu_z_cell),
      `Species score` = colDef(width = 95, format = colFormat(digits = 3), na = "—"),
      `All Species score` = colDef(width = 115, format = colFormat(digits = 3), na = "—"),
      `Mean` = colDef(width = 95, format = colFormat(digits = 2, separators = TRUE), na = "—"),
      `Q10` = colDef(width = 85, format = colFormat(digits = 2, separators = TRUE), na = "—"),
      `Q90` = colDef(width = 85, format = colFormat(digits = 2, separators = TRUE), na = "—"),
      `Species Mean` = colDef(width = 95, format = colFormat(digits = 2, separators = TRUE), na = "—"),
      `All Species Mean` = colDef(width = 95, format = colFormat(digits = 2, separators = TRUE), na = "—")
    )
  )
}
```

```{r prepare indicator table data}
# Prepare the indicator table data once for use in multiple tabs
cu_all_raw <- get_CU_indicators(all_std_long,
  cu_i = cu_i,
  use_standardized = FALSE,
  period_pick = "3",
  RCP_pick = "45")

cu_all_std <- get_CU_indicators(all_std_long,
  cu_i = cu_i,
  use_standardized = TRUE,
  period_pick = "3",
  RCP_pick = "45")

cu_all <- cu_all_raw %>%
  left_join(cu_all_std,
    by = c("FULL_CU_IN", "rcp", "period_code", "indicator"),
    suffix = c("_raw", "_std")) %>%
  left_join(select(tbl_indicators, abbrev, long_type), by = c("indicator" = "abbrev")) %>%
  mutate(
    across(where(is.numeric), ~ ifelse(is.nan(.), NA_real_, .))
  )

# Build display-only view with required columns
pick_col <- function(df, choices) {
  out <- choices[choices %in% names(df)]
  if (length(out) == 0) NA_character_ else out[1]
}

std_cu_col <- pick_col(cu_all, c("std_cu_value", "std_cu_mean", "cu_value_std", "cu_mean_std"))
std_sp_col <- pick_col(cu_all, c("std_sp_value", "std_sp_mean", "sp_value_std", "sp_mean_std"))
std_allcu_col <- pick_col(cu_all, c("std_allcu_value", "std_allcu_mean", "allcu_value_std", "allcu_mean_std"))
cu_col <- pick_col(cu_all, c("cu_value_raw", "cu_mean_raw", "cu_value", "cu_mean"))
q10_col <- pick_col(cu_all, c("cu_qlowgcm_raw", "cu_qlowgcm", "cu_qlowgcm_std"))
q90_col <- pick_col(cu_all, c("cu_qhighgcm_raw", "cu_qhighgcm", "cu_qhighgcm_std"))
sp_col <- pick_col(cu_all, c("sp_value_raw", "sp_mean_raw", "sp_value", "sp_mean"))
allcu_col <- pick_col(cu_all, c("allcu_value_raw", "allcu_mean_raw", "allcu_value", "allcu_mean"))

tbl_view <- cu_all %>%
  mutate(CU_Period = paste0(rcp, " • P", period_code)) %>%
  transmute(
    `Indicator Type` = long_type,
    `Indicator Code` = indicator,
    `CU score` = if (is.na(std_cu_col)) NA_real_ else .[[std_cu_col]],
    `Species score` = if (is.na(std_sp_col)) NA_real_ else .[[std_sp_col]],
    `All Species score` = if (is.na(std_allcu_col)) NA_real_ else .[[std_allcu_col]],
    `Mean` = if (is.na(cu_col)) NA_real_ else .[[cu_col]],
    `Q10` = if (is.na(q10_col)) NA_real_ else .[[q10_col]],
    `Q90` = if (is.na(q90_col)) NA_real_ else .[[q90_col]],
    `Species Mean` = if (is.na(sp_col)) NA_real_ else .[[sp_col]],
    `All Species Mean` = if (is.na(allcu_col)) NA_real_ else .[[allcu_col]]
  )

# Reorder display so CU block is together, species/all means at the end
display_order <- c(
  "Indicator Type",
  "Indicator Code",
  # z-scores
  "CU score", "Species score", "All Species score",
  # CU block
  "Mean", "Q10", "Q90",
  # tail
  "Species Mean", "All Species Mean"
)
tbl_view <- tbl_view[, intersect(display_order, names(tbl_view))]
```

# {.tabset .tabset-fade}

## About

This report explores climate vulnerability indicator data for the selected conservation unit (CU) in the Fraser Basin. The report displays maps and summaries of indicator data for the CU.

### Sections

- **Overview:** Summary of all climate vulnerability indicators for the selected conservation unit. 
- **Demographics:** Conservation status and recent spawner abundance
- **Timing:** Life stage timing for the CU and indicator time periods
- **Spawning and Rearing:** Environmental change to streams, cumulative threats to stream habitats, and freshwater residency times
- **Migration:** Changes to stream temperatures and flows during upstream migration, and distance to spawning sites
- **Marine:** Changes to sea surface temperature and cumulative impacts to nearshore marine habitat


### Default Parameters

All indicators are displayed using:

- **RCP 4.5:** Representative Concentration Pathway 4.5 (moderate emissions scenario)
- **Time Period:** Mid-Century (2041-2060)

### Data Quality

::: {style="background-color: #fff3cd; border: 1px solid #ffc107; padding: 15px; border-radius: 4px;"}
#### ⚠️ Draft Data Disclaimer

**DRAFT - Not for further distribution without permission of the authors**

This data and analysis are preliminary and subject to change. The indicators presented here are under active development and have not been peer-reviewed. Results should be interpreted with caution and are intended for exploratory analysis only. Please contact the authors before citing or using this information in any formal capacity.
:::

## Overview

### All Climate Vulnerability Indicators

Summary of all climate vulnerability indicators for the selected conservation unit. Green shading indicates values below species average (lower risk), while red indicates values above species average (higher risk).

```{r overview table}
table_id <- "cu_compact"

# Precompute scaling for coloring intensity (avoid division by zero)
diff_max <- tryCatch({
  x <- tbl_view[["CU score"]] - tbl_view[["Species score"]]
  m <- max(abs(x), na.rm = TRUE)
  if (is.infinite(m) || is.na(m) || m == 0) 1 else m
}, error = function(e) 1)

# Helpers
big_num_cols <- intersect(c("CU Mean", "Species Mean", "All Species Mean"), names(tbl_view))
big_num_def  <- function() colDef(format = colFormat(digits = 2, separators = TRUE), na = "—")
q_def        <- function() colDef(format = colFormat(digits = 2, separators = TRUE), na = "—")

# Color function for CU z vs Species z
cu_z_cell <- function(value, index, name) {
  sp <- tbl_view[["Species score"]][index]
  if (is.na(value) || is.na(sp)) return("—")
  diff <- value - sp
  # intensity scaled 0..1
  intensity <- min(1, abs(diff) / diff_max)
  # alpha between 0.15 and 0.55 for visibility
  alpha <- 0.15 + 0.40 * intensity
  bg <- if (diff < 0) {
    # greener for smaller CU z
    sprintf("rgba(34, 197, 94, %.3f)", alpha)     # Tailwind green-500 base
  } else if (diff > 0) {
    sprintf("rgba(239, 68, 68, %.3f)", alpha)     # Tailwind red-500 base
  } else {
    "rgba(148, 163, 184, 0.20)"                   # neutral (slate-400 @ 0.2)
  }
  div(
    style = list(
      backgroundColor = bg,
      borderRadius    = "6px",
      padding         = "1px 6px",
      display         = "inline-block"
    ),
    sprintf("%.3f", value)
  )
}
reactable(
  tbl_view,
  elementId = table_id,
  filterable = F,
  searchable = F,
  highlight = T,
  bordered = TRUE,
  striped = F,
  compact = TRUE,
  pagination = FALSE,
  defaultSorted = "Indicator Type",
  defaultSortOrder = "asc",

  # Slim defaults
  defaultColDef = colDef(
    minWidth = 70,
    headerVAlign = "center",
    vAlign = "center"
  ),

  theme = reactableTheme(
    borderColor = "#e5e7eb",
    stripedColor = "#f9fafb",
    highlightColor = "#eef2ff"
  ),

  # Column groups (short labels)
  columnGroups = list(
    colGroup(
      name = "Standardized Scores",
      columns = intersect(c("CU score","Species score","All Species score"), names(tbl_view))
    ),
    colGroup(
      name = "CU Raw Values",
      columns = intersect(c("Mean","Q10","Q90"), names(tbl_view))
    ),
    colGroup(
      name = "Raw Means",
      columns = intersect(c("Species Mean","All Species Mean"), names(tbl_view))
    )
  ),

  # Per-column formatting + widths
  columns = (function() {
    cols <- list()

    # Meta
    if ("Indicator Type" %in% names(tbl_view)) {
      cols[["Indicator Type"]] <- colDef(width = 100)
    }
    if ("Indicator Code" %in% names(tbl_view)) {
      cols[["Indicator Code"]] <- colDef(width = 115)
    }

    # z-scores: CU z gets comparative coloring; Species/All plain 3 decimals
    if ("CU score" %in% names(tbl_view)) {
      cols[["CU score"]] <- colDef(width = 95, format = colFormat(digits = 3), na = "—",
                               cell = cu_z_cell)
    }
    if ("Species score" %in% names(tbl_view)) {
      cols[["Species score"]] <- colDef(width = 95, format = colFormat(digits = 3), na = "—")
    }
    if ("All Species score" %in% names(tbl_view)) {
      cols[["All Species score"]] <- colDef(width = 115, format = colFormat(digits = 3), na = "—")
    }

    # CU block: mean (0 decimals), quantiles (2 decimals)
    if ("Mean" %in% names(tbl_view)) {
      cols[["Mean"]] <- big_num_def(); cols[["Mean"]]$width <- 95
    }
    if ("Q10" %in% names(tbl_view)) {
      cols[["Q10"]] <- q_def(); cols[["Q10"]]$width <- 85
    }
    if ("Q90" %in% names(tbl_view)) {
      cols[["Q90"]] <- q_def(); cols[["Q90"]]$width <- 85
    }

    # Species & All Species means at the end: 0 decimals
    if ("Species Mean" %in% names(tbl_view)) {
      cols[["Species Mean"]] <- big_num_def(); cols[["Species Mean"]]$width <- 95
    }
    if ("All Species Mean" %in% names(tbl_view)) {
      cols[["All Species Mean"]] <- big_num_def(); cols[["All Species Mean"]]$width <- 95
    }

    cols
  })(),

  # Row highlighting (kept)
  rowClass = function(index, name) {
    ind <- tbl_view[["Indicator Code"]][index]
    if (!is.null(ind) && ind %in% c("tw8proj","tw8rate","SSTproj","SSTrate")) {
      "climate-row"
    } else if (!is.null(ind) && ind %in% c("migrTproj","migrQpdelta","migrdist")) {
      "migration-row"
    } else {
      ""
    }
  }
) %>%
  htmlwidgets::prependContent(
    htmltools::tags$h4(
      paste0(cu_i, " Indicator Summary, RCP ",rcp_pick, ", Period ", period_pick),
      style = "margin: 0 0 6px; font-size: 14px; font-weight: 600;"
    )
  ) %>%
  htmlwidgets::prependContent(tags$style(
    HTML(sprintf("
      /* Compact font + padding for this table only */
      #%s .rt-table { font-size: 11.5px; }
      #%s .rt-th, #%s .rt-td { padding: 3px 6px; }
      #%s .climate-row { background-color: #fcffeb; }
      #%s .migration-row { background-color: #eff6ff; }
      #%s .rt-search { max-width: 240px; }
      #%s .rt-td div { line-height: 1.15; }
      #%s .rt-col-group > div { padding: 4px 6px; }
    ", table_id, table_id, table_id, table_id, table_id, table_id, table_id, table_id))
  ))
```

## Demographics

### Demographic Indicators Summary

Recent conservation status and number of mature individuals indicator values.

```{r demographics plot}
plot_cu_indicators_lollipop(
  ind_cu,
  indicators_choose = c("CUstatus", "CUnmat")
)
```

### Spawner Abundance Time Series

Time series of annual spawner abundance (blue line) and generational geometric average (red line). Points at bottom indicate Wild Salmon Policy status assessment (Red/Amber/Green) with shape indicating data quality confidence.

```{r abundance plot}

if(nrow(status_data[status_data$FULL_CU_IN == cu_i,]) > 0)  
  {
  aplot <- abundance_status_plot(status_data, 
                      cu_i)
  aplot 
}

```

## Timing

### Life History Timing

Timing by life history stage is taken from Wilson and Peacock (2025) [<https://cdnsciencepub.com/doi/10.1139/cjfas-2024-0213>]. Data quality is rated on a scale of 1 to 6 based on number of observations and their recentness, with 1 being the highest quality data.

Life stage timing throughout the year showing spawning, upstream migration, ocean entry, and juvenile freshwater migration periods. Shaded areas indicate periods used for calculating specific vulnerability indicators. Point size represents data quality (larger = LOWER quality).

```{r life stage timing}

tp <- cu_timing_plot(cu_timing_long_i)

tp
```

## Spawning & Rearing {.tabset .tabset-pills}

### CU Location

Geographic location of the selected conservation unit (green) relative to all Fraser River basin conservation units (grey outlines).

```{r cu boundary}
### boundary location
p1 <- cu_boundary_highlight(cu_boundary, 
                      cu_pick = cu_i)

p1
```

### Stream Access

#### BC FishPass Statistics

Summary statistics for accessible stream habitat within the conservation unit boundary based on BC Fishpass habitat potential models.

```{r data tables}
fw_cu %>%
  select(c(total_length_acc, n_streams, proportion_rear, proportion_spawn, proportion_rs,
           avg_elevation, avg_lat, avg_lon)) %>%
  mutate(
    total_length_acc = round(total_length_acc / 1000, 1),
    proportion_rear = round(proportion_rear, 3),
    proportion_spawn = round(proportion_spawn, 3),
    proportion_rs = round(proportion_rs, 3),
    avg_elevation = round(avg_elevation, 1),
    avg_lat = round(avg_lat, 5),
    avg_lon = round(avg_lon, 5)
  ) %>%
  gt() %>%
  tab_header(
    title = md("**CU Boundary Stream Statistics**")
  ) %>%
  cols_label(
    total_length_acc = "Accessible Stream Length (km)",
    n_streams = "Number of Stream Segments",
    avg_elevation = "Avg. Elevation (m)",
    proportion_rear = "Proportion Rearing",
    proportion_spawn = "Proportion Spawning",
    proportion_rs = "Proportion Rearing/Spawning"
  ) %>%
  tab_spanner(
    label = "Habitat Model",
    columns = c(proportion_rear, proportion_spawn, proportion_rs)
  ) %>%
  tab_options(
    table.font.size = "small",           # Reduce overall font size
    data_row.padding = px(1),           # Minimize padding between rows
  )
```

#### Accessible Stream Network

Accessible streams only with coloring indicating BC Fishpass modelled habitat potential and NuSEDS spawning site locations within the conservation unit boundary.

```{r accessibility plot}
acc_plot <- stream_accessible_plot(fw_sp_cu,
  nuseds_cu,
  cu_boundary_i,
  lakes_cu)

acc_plot
```

### Hydrology

#### Hydrologic Regime Coverage

Proportion of CU boundary classified by hydrologic regime type using method described in Ruzzante et al 2025. prop_coverage indicates proportion of CU boundary within gauged watersheds.

```{r hydrology}
fw_cu %>%
  select(contains("prop_")) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  gt() %>%
  tab_header(
    title = md("**Hydrologic Regime Coverage Proportions**")
  ) %>%
  cols_label(
    prop_snow = "Snow-Dominated",
    prop_rain = "Rain-Dominated",
    prop_hybrid = "Hybrid",
    prop_glacial = "Glacial",
    prop_coverage = "Gauged Coverage"
  ) %>%
  fmt_number(
    columns = where(is.numeric),
    decimals = 3
  ) %>%
  tab_options(
    table.font.size = "small",           # Reduce overall font size
    data_row.padding = px(1),           # Minimize padding between rows
  )
```

#### Hydrologic Regime Map

Map of hydrologic regime zones and flow gauge locations within the CU boundary using method described in Ruzzante et al 2025.

```{r regime plot}
# hydrologic gauge stations and regime plot
regime_plot <- cu_hydrologic_regime(cu_boundary_i,
                                    stream_data = fw_sp_cu,
                                 watershed_flow_cu,
                                 stations_cu,
                                 fw_sp_cu)

regime_plot
```

### Flow

#### Flow Indicators Summary

Projected changes in stream flow during the month of August (flow8pdelta) and Nov-Jan (flow18pdelta) for RCP 4.5, 2041-2060.

```{r flow indicators table}
# Create subset table for flow indicators
create_indicator_reactable(tbl_view, c("flow8pdelta", "flow18pdelta"))
```

#### Change in August Flow

Projected proportional change in August stream flows. Negative values indicate flow reductions relative to historic period.

```{r flow indicator plots}
Q8_plot <- stream_indicator_plot(acc_sp_cu, 
                                 cu_boundary_i,
                                 lakes_cu,
                                variable = "flow8pdelta_9_45_3", 
                                plot_title = "Change in August flow - 2041-2060",
                                unit_label = "Proportion change",
                                xlim = c(-1,0),
                                scico_palette = "lapaz",
                                palette_direction = -1)

Q8_plot
```

#### Change in Nov-Jan Flow

Projected proportional change in winter flows (November-January). Positive values indicate flow increases relative to historic period.

```{r winter flow plot}
QNDJ_plot <- stream_indicator_plot(acc_sp_cu, 
                                 cu_boundary_i,
                                 lakes_cu,
                                variable = "flow18pdelta_9_45_3", 
                                plot_title = "Change in Nov_Jan flow - 2041-2060",
                                unit_label = "Proportion change",
                                scico_palette = "devon",
                                xlim = c(0,1),
                                palette_direction = -1)
  

QNDJ_plot
```

### Temperature

#### Temperature Indicators Summary

Stream temperature indicators: projected August temperature (tw8proj) and rate of change in August temperature (tw8rate).

```{r temperature table}
# Create subset table for temperature indicators
create_indicator_reactable(tbl_view, c("tw8proj", "tw8rate"))
```

#### Projected August Stream Temperature

Projected mean August stream temperature (2041-2060, RCP 4.5).

```{r temperature}
#temperature plot
t_plot <- stream_indicator_plot(acc_sp_cu, 
                                cu_boundary_i, 
                                lakes_cu,
                                Tw_stations,
                                variable = "tw8proj_9_45_3", 
                                temp_stations = TRUE,
                                plot_title = "August Mean Temperature - 2041-2060",
                                unit_label = "Degrees C",
                                scico_palette = "roma",
                                palette_direction = -1)

t_plot
```

#### Rate of Temperature Change

Rate of change in August stream temperature from baseline (1981-2000) to mid-century (2041-2060).

```{r temperature rate plot}
tRate_plot <- stream_indicator_plot(acc_sp_cu, 
                                cu_boundary_i, 
                                lakes_cu,
                                Tw_stations,
                                variable = "tw8rate_9_45_3", 
                                temp_stations = TRUE,
                                plot_title = "Rate of Change in August Mean Temperature - 1981-2000 to 2041-2060",
                                unit_label = "Degrees C",
                                scico_palette = "roma",
                                palette_direction = -1)

tRate_plot
```

### ENM

#### ENM Indicators Summary

Indicator value of changes in habitat favourability (favchange) from Environmental Niche Models.

```{r enm table}
# Create subset table for ENM indicators
create_indicator_reactable(tbl_view, c("favchange"))
```


#### Change in Favourability

Change in habitat favourability from baseline to mid-century. Negative values indicate decreased favourability, positive values indicate improving conditions.

```{r enm diff plot}
ENMdiff_plot <- stream_indicator_plot(fw_sp_cu, 
                                 cu_boundary_i,
                                 lakes_cu,
                                variable = "favchange_45", 
                                plot_title = "ENM favourability change, 1981-2000 to 2041-2060",
                                histogram_fill = "model_access_salmon",
                                scico_palette = "berlin",
                                xlim = c(-1,1),
                                palette_direction = -1)
  

ENMdiff_plot 
```

### Threats

#### Cumulative Threats Indicator Summary

Indicator of cumulative threat score (cthr) representing anthropogenic stressors to stream habitat.

```{r threats table}
# Create subset table for cumulative threats indicators
create_indicator_reactable(tbl_view, c("cthr_anad"))
```

#### Cumulative Threats to Stream Habitat

Cumulative threat score combining nine anthropogenic stressor categories. Higher values represent higher threats.

```{r Cumulative Threats}

ct_plot <- stream_indicator_plot(acc_sp_cu, 
                                 cu_boundary_i,
                                 lakes_cu,
                                variable = "cthr_anad", 
                                plot_title = "Cumulative stressors",
                                unit_label = "Cumulative Threat Score",
                                scico_palette = "lajolla",
                                palette_direction = -1)


ct_plot

```

## Migration {.tabset .tabset-pills}

### Path

#### Migration Indicators Summary

Indicators representing migration temperature (migrTproj), migration flow change (migrQpdelta), and migration distance from ocean to spawning grounds.

```{r migration table}
# Create subset table for migration indicators
create_indicator_reactable(tbl_view, c("migrTproj", "migrQpdelta", "migrdist"))
```

#### Migration Route

Migration route from river mouth to NuSEDS spawning sites used to calculate indicators, coloured by channel width.

```{r MIgration path}

mplot <- migration_path_plot(migr_cu,
  nuseds_cu,
  cu_boundary_i,
  colour_var = "channel_width",
colour_label = "Channel width (m)",
plot_title = paste0("Migration path length = ", cu_std$cu_value[cu_std$indicator == "migrdist"][1], "km") 
)

mplot
```

### Timing - Migration

#### Migration Timing and Temperature

Daily projected stream temperatures during upstream migration period. Vertical dashed lines show run timing start/end (blue) and spawn timing start/peak (red). Shaded areas show 10th-90th percentile range across climate models.

```{r Migration temp}

timeplot <- migr_timing_plot(migr_daily_all,
                             cu_i,
                             cu_timing_i,
                             rcp = "45",
                             period_choose = c("1981-2010", "2041-2060"))

timeplot

```

## Marine {.tabset .tabset-pills}

### Marine Adaptive Zone

Marine Adaptive Zone (green) representing the nearshore rearing area used for marine indicator calculations.

```{r Marine}

### subset indicator values
cu_MAZ <- cvis_cu_list$MAZ[cvis_cu_list$FULL_CU_IN == cu_i][1]

start_month <- month.name[unique(cu_timing_i$ns_start_month)]
end_month <- month.name[unique(cu_timing_i$ns_end_month)]

MAZ_pick <- cu_MAZ

MAZ_boundary_highlight(MAZ,
                          MAZ_pick = cu_MAZ)
```


### SST

#### Sea Surface Temperature Indicators Summary

Nearshore marine indicators related to sea surface temperature, the projected SST during the ocean entry period (SSTproj), and the rate of change in SST (SSTrate) relative to the baseline period.

```{r sst table}
# Create subset table for SST indicators
create_indicator_reactable(tbl_view, c("SSTproj", "SSTrate"))
```

#### Projected Sea Surface Temperature

Projected sea surface temperature during the period surrounding peak ocean entry (2041-2060, RCP 4.5).

```{r SST}

# SST months
months_include <- seq(
    from = cu_timing_i$ns_start_month[1],
    to = cu_timing_i$ns_end_month[1],
    by = 1
  )

rcp_pick <- "45"
period_pick <- 3

#subset SST time and locations spatial object
SST_cu_sp <- get_spatial_var(CMIP6_SST,
                             months = months_include,
                             period_pick = period_pick,
                             rcp_pick = rcp_pick)

p <- marine_indicator_plot(SST_cu_sp,
                           MAZ_sp = MAZ,
                           var = "SST_oe",
                           unit_label = "Degrees C",
                           plot_title = paste0("Projected SST, 2041-2060, RCP ", rcp_pick),
                           palette_limits = c(9,16))
                           
p
```

### Impacts

#### Marine Cumulative Impacts Indicator Summary

Indicator for cumulative impact score (CImpact) for nearshore marine habitats within the marine adaptive zone.

```{r impacts table}
# Create subset table for marine impacts indicators
create_indicator_reactable(tbl_view, c("CImpact"))
```

#### Cumulative Impacts on Marine Habitat

Cumulative impact score combining stressors including aquaculture, fishing, invasive species, pollution, and shipping. Higher values indicate a higher impact score.

```{r Cumulative impacts}

p2 <- marine_indicator_plot(filter(CImpact_points, MAZ_Acrony == cu_MAZ),
                      MAZ_sp = MAZ,
                      var = "Cumul_Impact_ALL",
                      unit_label = "CI Score",
                      palette_limits = c(NA,NA))

p2
```
