---
output:
  html_document:
    mathjax: null
    toc: false
    number-sections: false
    theme: flatly
    css: "scientific_style.css"
params:
  FULL_CU_IN: "CK-03"
  default_rcp: "45"
  default_period: "3"
title: "`r local({ load(here::here('processed_data', 'CU', 'cu_run.Rds')); paste0(params$FULL_CU_IN, ': ', cu_run$CVIS_LABEL[cu_run$FULL_CU_IN == params$FULL_CU_IN], ' (', cu_run$SPECIES_NAME[cu_run$FULL_CU_IN == params$FULL_CU_IN], ')') })`"
---

```{r custom theme, echo = FALSE, message = F, warning = F}
knitr::opts_chunk$set(message = F, echo = F, warning = F, dev = "jpeg", jpegquality = 75, dpi = 90)
library(magrittr)
library(reactable)
library(htmltools)
library(sf)
library(tidyverse)
library(gt)
source(here::here("code", "5d_table_summaries.R"))
source(here::here("code", "5b_plots_compare.R"))
```

```{r load data, echo=FALSE, message=FALSE, warning=FALSE}
# Helper to find the latest date-prefixed file matching a pattern
get_latest_file <- function(path, pattern) {
  files <- list.files(path, pattern = pattern, full.names = TRUE)
  if (length(files) == 0) {
    files <- list.files(here::here("output"), pattern = pattern, full.names = TRUE)
  }
  if (length(files) == 0) {
    stop("No files matching pattern '", pattern, "' found in ", path)
  }
  sorted_files <- sort(files, decreasing = TRUE)
  return(sorted_files[1])
}

# Ensure setup and scoring results are loaded
if (!exists("paths")) {
  source(here::here("code", "0_setup.R"))
}
if (!exists("all_std_long") || !exists("all_std_long_baseline")) {
  load(file.path(here::here(), "output", "scoring_results.Rdata"))
}
if (!exists("fw_all") || !exists("ss_all")) {
  load(get_latest_file(paths$fw, "fw_rearing_indicators.Rdata"))
}
if (!exists("overall_sensitivity")) {
  load(file.path(here::here(), "output", "sensitivity_analysis.Rdata"))
}
if (!exists("stream_cu_picks")) {
  load(file.path(paths$fw, "fw_streampicks_tscapes.Rdata"))
}
if (!exists("migr_list")) {
  load(file.path(paths$fw, "fw_upstream_paths.Rdata"))
}
if (!exists("fw_sp_ind")) {
  load(file.path(paths$fw, "fw_stream_indicators_sp.Rds"))
}
if (!exists("cu_lakes_precut")) {
  cu_lakes_precut <- readRDS(file.path(here::here(), "output", "cu_lakes_precut.Rds"))
}
if (!exists("watershed_flow")) {
  load(file.path(paths$fw, "flow_gauge_data.Rdata"))
}
if (!exists("cu_timing_long")) {
  load(file.path(paths$CU, "cu_timing_data.Rdata"))
}
if (!exists("status_data")) {
  load(file.path(paths$CU, "cu_status_data.Rds"))
}
if (!exists("cu_run")) {
  load(file.path(paths$CU, "cu_run.Rds"))
}
if (!exists("cu_boundary")) {
  load(file.path(paths$fw, "cu_boundary.Rds"))
  cu_boundary <- cu_boundary %>% 
    left_join(select(spp_lookup, PSF_species, SPECIES_NAME), join_by(Species == PSF_species))
}
if (!exists("migr_daily_all")) {
  load(get_latest_file(paths$fw, "migr_stats.Rdata"))
}
if (!exists("MAZ")) {
  load(file.path(paths$marine, "MAZ.Rds"))
}

# Setup spatial background
Fr_basin <- filter(basins, BASIN == "FRASER")

# Helper functions for SMU calculations and coloring
get_CU_indicators_smu <- function(data, cu_i, use_standardized = TRUE, RCP_pick = "45", period_pick = "3") {
  data_sub <- data %>%
    filter((rcp == RCP_pick & period_code == period_pick) | (rcp == "0" & period_code == "0")) %>%
    group_by(FULL_CU_IN, indicator, stat) %>%
    mutate(has_projection = any(rcp == RCP_pick & period_code == period_pick)) %>%
    filter(!has_projection | (rcp == RCP_pick & period_code == period_pick)) %>%
    ungroup() %>%
    select(-has_projection)

  val_col <- if (use_standardized) "std_value" else "value"

  if ("gcm" %in% names(data_sub)) {
    if ("gcm_name" %in% names(data_sub)) {
      data_sub <- data_sub %>% filter(gcm == "9" | gcm == "0" | (is.na(gcm) & is.na(gcm_name)) | gcm_name == "ensemble")
    } else {
      data_sub <- data_sub %>% filter(gcm == "9" | gcm == "0" | is.na(gcm))
    }
  }

  if (exists("tbl_standardize")) {
    if ("dsmodel" %in% names(data_sub)) {
      data_sub <- data_sub %>%
        left_join(dplyr::select(tbl_standardize, abbrev, dsmodel_baseline_ind = dsmodel_baseline), by = c("indicator" = "abbrev")) %>%
        filter(is.na(dsmodel) | dsmodel == dsmodel_baseline_ind | is.na(dsmodel_baseline_ind)) %>%
        dplyr::select(-dsmodel_baseline_ind)
    }
    if ("std_method" %in% names(data_sub)) {
      data_sub <- data_sub %>%
        left_join(dplyr::select(tbl_standardize, abbrev, std_fun), by = c("indicator" = "abbrev")) %>%
        mutate(
          std_fun_base = case_when(
            std_fun %in% c("linear_std", "invlinear_std") ~ "linear",
            TRUE ~ "exponential"
          )
        ) %>%
        filter(is.na(std_method) | std_method == std_fun_base) %>%
        dplyr::select(-std_fun, -std_fun_base)
    }
  }

  data_sub <- data_sub %>%
    filter(
      indicator %in% tbl_indicators$abbrev,
      stat %in% c("mean", "qlowgcm", "qhighgcm")
    ) %>%
    select(FULL_CU_IN, SPECIES_NAME, SMU_SIMPLE, indicator, stat, !!sym(val_col)) %>%
    rename(val = !!sym(val_col))

  # Get current CU's SMU
  cu_row <- cu_run[cu_run$FULL_CU_IN == cu_i, ]
  smu_pick <- cu_row$SMU_SIMPLE[1]

  # Calculate SMU averages
  smu_avgs <- data_sub %>%
    filter(SMU_SIMPLE == smu_pick) %>%
    group_by(indicator, stat) %>%
    summarise(val = mean(val, na.rm = TRUE), .groups = "drop") %>%
    rename(smu = val)

  # Calculate All Species averages
  all_cu_avgs <- data_sub %>%
    group_by(indicator, stat) %>%
    summarise(val = mean(val, na.rm = TRUE), .groups = "drop") %>%
    rename(allcu = val)

  # Get CU values
  data_CU <- data_sub %>%
    filter(FULL_CU_IN == cu_i) %>%
    select(indicator, stat, val) %>%
    rename(cu = val)

  # Join all together
  out_data <- data_CU %>%
    left_join(smu_avgs, by = c("indicator", "stat")) %>%
    left_join(all_cu_avgs, by = c("indicator", "stat")) %>%
    mutate(FULL_CU_IN = cu_i, rcp = RCP_pick, period_code = period_pick) %>%
    relocate(FULL_CU_IN, rcp, period_code)

  # Rename and pivot
  out_data <- out_data %>%
    pivot_wider(
      names_from = stat,
      values_from = c("cu", "smu", "allcu"),
      names_glue = "{.value}_{stat}"
    ) %>%
    rename_with(~ str_replace(., "_mean$", "_value"), ends_with("_mean"))

  return(out_data)
}

# Pastel RdYlBu interpolator (soft/easy-to-read backgrounds)
get_cvis_table_color <- function(val) {
  sapply(val, function(v) {
    if (is.na(v) || is.nan(v)) return("#ffffff")
    v <- max(0, min(1, v))
    pal <- colorRamp(c("#e0f3f8", "#ffffeb", "#fde8e8"), space = "rgb")
    rgb_val <- pal(v)
    rgb(rgb_val[1,1], rgb_val[1,2], rgb_val[1,3], maxColorValue = 255)
  })
}
```

```{r subset CU}
cu_i <- params$FULL_CU_IN
cu_run_i <- cu_run[cu_run$FULL_CU_IN == params$FULL_CU_IN,]
sp_pick <- cu_run$SPECIES_NAME[cu_run$FULL_CU_IN == cu_i] # species abbr
sp_pick_bcfp <- spp_lookup$spp_abr_bcfp[spp_lookup$SPECIES_NAME == sp_pick] 
sp_pick_ENM  <- str_to_lower(sp_pick)[1]

# default RCP and period values
rcp_pick <- params$default_rcp
period_pick <- params$default_period

# Assign to global environment so lexical-scoping functions in plots can access
assign("cu_boundary_i", cu_boundary[cu_boundary$FULL_CU_IN == cu_i,], envir = .GlobalEnv)
assign("nuseds_cu", nuseds_Fr[nuseds_Fr$FULL_CU_IN == cu_i,], envir = .GlobalEnv)
```

```{r spatial subset, echo = FALSE, message = F, warning = F}
# indicator dataframe
cu_std <- get_CU_indicators(all_std_long,
  cu_i = cu_i,
  use_standardized = FALSE,
  period_pick = period_pick,
  RCP_pick = rcp_pick)

# Stream indicator summary rows (fw_all format)
fw_cu <- ss_all %>%
  filter(FULL_CU_IN == cu_i) %>%
  distinct(FULL_CU_IN, .keep_all = TRUE)

fw_sp_cu <- subset_fw_models(
  fw_models = fw_sp_ind,
  cu_i = cu_i,
  stream_cu_picks = stream_acc_cu_picks,
  cu_run = cu_run,
  spp_lookup = spp_lookup,
  to_factor = TRUE
) %>%
  rename(favchange_45   = contains(paste0("favchange_", sp_pick_ENM, "_45_3")),
         favchange_85   = contains(paste0("favchange_", sp_pick_ENM, "_85_3")),
         fav_45   = contains(paste0("fav_", sp_pick_ENM, "_45_3")),
         fav_85   = contains(paste0("fav_", sp_pick_ENM, "_85_3")))

# subset in CU boundary
temp <- unlist(st_intersects(cu_boundary_i, watershed_flow))
watershed_flow_cu <- watershed_flow[temp,]

# subset in cu boundary
temp <- unlist(st_intersects(cu_boundary_i, stations_flow))
stations_cu <- stations_flow[temp,]

# Migration spatial layers
migr_cu <- migr_list[[cu_i]]

# Lakes within CU boundary (from pre-cut dataset)
lakes_cu <- cu_lakes_precut[[cu_i]]

# Timing data
cu_timing_long_i <- cu_timing_long[cu_timing_long$FULL_CU_IN == cu_i, ]
cu_timing_i <- cu_timing_Fr[cu_timing_Fr$FULL_CU_IN == cu_i, ]

# get long data frame of standardized indicator values for the CU
ind_cu <- get_CU_indicators(all_std_long,
  cu_i,
  indicators_choose = tbl_indicators$abbrev,
  RCP_pick = rcp_pick,
  period_pick = period_pick
) %>%
  dplyr::mutate(
    SPECIES_NAME = unique(all_std_long$SPECIES_NAME[all_std_long$FULL_CU_IN == cu_i])[1],
    CVIS_LABEL = unique(all_std_long$CVIS_LABEL[all_std_long$FULL_CU_IN == cu_i])[1]
  )
```

```{r prepare indicator table data}
# Prepare the indicator table data
cu_all_raw <- get_CU_indicators_smu(all_std_long, cu_i = cu_i, use_standardized = FALSE, RCP_pick = rcp_pick, period_pick = period_pick)
cu_all_std <- get_CU_indicators_smu(all_std_long, cu_i = cu_i, use_standardized = TRUE, RCP_pick = rcp_pick, period_pick = period_pick)

cu_all <- cu_all_raw %>%
  left_join(cu_all_std,
    by = c("FULL_CU_IN", "rcp", "period_code", "indicator"),
    suffix = c("_raw", "_std")) %>%
  left_join(select(tbl_indicators, abbrev, long_type), by = c("indicator" = "abbrev")) %>%
  mutate(
    across(where(is.numeric), ~ ifelse(is.nan(.), NA_real_, .))
  )

tbl_view <- cu_all %>%
  transmute(
    `Indicator Type` = long_type,
    `Indicator Code` = indicator,
    `CU score` = cu_value_std,
    `SMU score` = smu_value_std,
    `All Species score` = allcu_value_std,
    `Mean` = cu_value_raw,
    `Q10` = cu_qlowgcm_raw,
    `Q90` = cu_qhighgcm_raw,
    `SMU Mean` = smu_value_raw,
    `All Species Mean` = allcu_value_raw
  )
```

# {.tabset .tabset-fade}

## Overview {.tabset .tabset-fade}

### Vulnerability Summary

```{r vulnerability profile plot, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 9.5, fig.width = 8.5, out.width = "100%"}
plot_cu_vulnerability_summary(
  cu_code = cu_i,
  scores_tidy = scores_tidy_baseline,
  all_std_long = all_std_long_baseline,
  indicators_metadata = tbl_indicators
)
```

**Fig. S2.1.** Summary of climate vulnerability for the selected Conservation Unit (CU). The top panel shows the Overall Vulnerability score (calculated using the category average methodology) and category-specific scores (scaled 0 to 100). The bottom panel shows the standardized values of individual indicators (scaled 0 to 1) grouped by category. Background violins represent the distribution of baseline means across all CUs in the entire dataset. Plotted points and text labels indicate the standardized and raw values for the selected CU. 

### All Climate Vulnerability Indicators

```{r raw baseline violins plot, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 7, fig.width = 11, out.width = "100%"}
plot_raw_baseline_violins(
  all_std_long_baseline = all_std_long_baseline,
  tbl_indicators = tbl_indicators,
  cu_code = cu_i
) + plot_annotation(title = NULL, subtitle = NULL)
```

**Fig. S2.2.** Distribution of raw indicator values under the default scenario across all Conservation Units (CUs). Violins show the distributions across all CUs, and the orange diamond highlights the value for the selected CU. 

**Table S2.1.** Summary of all climate vulnerability indicators for the selected Conservation Unit (CU). The table shows default scenario mean and variation in raw units across all CUs, along with means by species and the values for the current CU. Vulnerability risk is evaluated across five core categories: Demographics, Freshwater Spawning & Rearing, Genetics, Marine Exposure, and Upstream Migration.

```{r overview table}
generate_cvis_summary_table(
  all_std_long_baseline = all_std_long_baseline,
  cu_code = cu_i,
  group_by = "species"
)
```

### CU Boundary & MAZ

```{r cu and maz boundary highlights, fig.height = 4, fig.width = 10, dev = "jpeg", jpegquality = 75, dpi = 90}
# CU locator plot
p_cu <- cu_boundary_highlight(cu_boundary, cu_pick = cu_i) + labs(tag = "a")

# MAZ locator plot
cu_MAZ <- cvis_cu_list$MAZ[cvis_cu_list$FULL_CU_IN == cu_i][1]
if (!is.na(cu_MAZ) && cu_MAZ != "") {
  p_maz <- MAZ_boundary_highlight(MAZ, MAZ_pick = cu_MAZ) + labs(tag = "b")
  
  # Combine side-by-side
  p_combined <- p_cu + p_maz
  p_combined
} else {
  p_cu
}
```

**Fig. S2.3.** Geographic location of the selected Conservation Unit (CU) boundary (a) and its associated Marine Adaptive Zone (MAZ) (b).

### Life History Timing

```{r life stage timing, fig.height=10, fig.width=9.5, out.width="100%"}
if (nrow(cu_timing_long_i) > 0) {
  tp <- plot_timing_comparison(cu_timing_long_i) + labs(title = NULL, subtitle = NULL)
  tp
}
```

**Fig. S2.4.** Life history stage timing (spawning, adult upstream migration, ocean entry, and juvenile freshwater migration) throughout the year for the selected Conservation Unit (CU) relative to all Fraser River basin CUs. Bars represent 5th-95th percentiles of observed timing, points represent peak timing. Bar opacity and point shapes represent data quality (Circle = High 1-2, Triangle = Medium 3-4, Square = Low 5-6). Data are adapted from Wilson and Peacock (2025) [<https://cdnsciencepub.com/doi/10.1139/cjfas-2024-0213>].

## Demographics

### Demographic Indicators Summary

**Table S2.2.** Demographic indicator values for the selected Conservation Unit (CU), compared against all CUs and species. Demographics evaluate Wild Salmon Policy (WSP) status (standardized as Green = 0, Amber = 0.5, Red = 1) and the geometric average of natural-origin spawners from the most recent generation (CUnmat).

```{r demographics table}
generate_cvis_summary_table(
  all_std_long_baseline = all_std_long_baseline,
  cu_code = cu_i,
  group_by = "species",
  categories = "dem")
```

### Spawner Abundance Time Series

```{r abundance plot, fig.height=4, fig.width=6.5}
if(nrow(status_data[status_data$FULL_CU_IN == cu_i,]) > 0)  
{
  aplot <- abundance_status_plot(status_data, cu_i)
  aplot 
}
```

**Fig. S2.5.** Annual spawner abundance time series (blue line) and generational geometric average (red line) for the selected Conservation Unit (CU). Points at the bottom indicate Wild Salmon Policy (WSP) status assessments (Green, Amber, Red).



## Spawning & Rearing {.tabset .tabset-pills}

### Spawning & Rearing Indicators


```{r rearing multipanel plot, fig.height = 9, fig.width = 8.5, out.width = "100%", dev = "jpeg", jpegquality = 75, dpi = 90}
variables_rearing <- c("favchange_45", "cthr_anad", "tw8proj_9_45_3", "tw8rate_9_45_3", "flow8pdelta_9_45_3", "flow18pdelta_9_45_3")
titles_rearing <- c("Change in ENM Favourability", "Cumulative Threat Score", "August Mean Temperature", "Rate of Temp. Change", "Change in August Flow", "Change in Nov-Jan Flow")
palettes_rearing <- cvis_risk_palette
directions_rearing <- c(1, -1, -1, -1, 1, -1)

f2 <- stream_indicator_multipanel_plot(
  fw_sp_cu,
  cu_boundary_i,
  lakes_cu,
  variables = variables_rearing,
  plot_titles = titles_rearing,
  risk_palette = palettes_rearing,
  palette_directions = directions_rearing
)
f2
```

**Fig. S2.6.** Projected values of spawning and rearing indicators across accessible stream segments within the selected CU boundary. Stream segments shown represent all accessible streams, however, only those modelled as potential rearing/spawning habitat are included when calculating mean indicator values by CU. Inset histograms show the distribution of values across accessible stream segments.

**Table S2.3.** Spawning and rearing indicator values for the selected Conservation Unit (CU), compared against all CUs and species under baseline and climate projection scenarios. Includes August mean temperature (tw8proj) and warming rate (tw8rate), proportional flow changes in August (flow8pdelta) and winter (flow18pdelta), ENM habitat suitability changes (favchange), and watershed cumulative threats (cthr).

```{r spawning rearing indicators table}
generate_cvis_summary_table(
  all_std_long_baseline = all_std_long_baseline,
  cu_code = cu_i,
  group_by = "species",
  categories = "fwrs",
  title = "Spawning & Rearing Indicators Summary",
  subtitle = paste0("Baseline scenario mean and variation, and comparison against all CUs and species (RCP ", rcp_pick, ", Period ", period_pick, ")")
)
```

### Stream Access (BC Fishpass)

**Table S2.4.** Summary of the stream network and modelled rearing and spawning habitat within the Conservation Unit (CU) boundary using the BC Freshwater Atlas and the BC Fishpass access model and linear spawning/rearing habitat model [<https://smnorris.github.io/bcfishpass/>].

```{r stream stats table}
if (nrow(fw_cu) > 0) {
  fw_cu %>%
    select(c(total_length_acc, n_streams, proportion_rear, proportion_spawn, proportion_rs,
             avg_elevation, avg_lat, avg_lon)) %>%
    mutate(
      total_length_acc = round(total_length_acc / 1000, 1),
      proportion_rear = round(proportion_rear, 3),
      proportion_spawn = round(proportion_spawn, 3),
      proportion_rs = round(proportion_rs, 3),
      avg_elevation = round(avg_elevation, 1),
      avg_lat = round(avg_lat, 5),
      avg_lon = round(avg_lon, 5)
    ) %>%
    gt() %>%
    tab_header(
      title = md("**CU Boundary Stream Statistics**")
    ) %>%
    cols_label(
      total_length_acc = "Accessible Stream Length (km)",
      n_streams = "Number of Stream Segments",
      avg_elevation = "Avg. Elevation (m)",
      proportion_rear = "Proportion Rearing",
      proportion_spawn = "Proportion Spawning",
      proportion_rs = "Proportion Rearing/Spawning"
    ) %>%
    tab_spanner(
      label = "Habitat Model",
      columns = c(proportion_rear, proportion_spawn, proportion_rs)
    ) %>%
    tab_options(
      table.font.size = "small",
      data_row.padding = px(1)
    )
}
```

```{r accessibility plot, fig.height = 5.5, fig.width = 6.5, dev = "jpeg", jpegquality = 75, dpi = 90}
acc_plot <- stream_accessible_plot(fw_sp_cu,
  nuseds_cu,
  cu_boundary_i,
  lakes_cu)
acc_plot
```

**Fig. S2.7.** Accessible streams within the CU boundary coloured by BC Fishpass modelled linear/spawning habitat. NuSEDS spawning sites are also shown. Stream segments that are not accessible are not shown.

## Migration

### Migration Indicators Summary
**Table S2.5.** Upstream migration indicator values for the selected Conservation Unit (CU), compared against all CUs and species. Indicators include projected migration temperature (migrTproj) and flow changes (migrQpdelta) averaged over a moving window corresponding to migration timing (Wilson & Peacock 2025) using VIC-GL model projections, and total upstream migration distance (migrdist) reflecting energetics.

```{r migration indicators table}
generate_cvis_summary_table(
  all_std_long_baseline = all_std_long_baseline,
  cu_code = cu_i,
  group_by = "species",
  categories = "migr"
)
```

### Migration Path and Temperature

```{r migration path timing plot, fig.height = 5, fig.width = 9.5, dev = "jpeg", jpegquality = 75, dpi = 90}
if (!is.null(migr_cu) && nrow(nuseds_cu) > 0 && nrow(migr_daily_all[migr_daily_all$FULL_CU_IN == cu_i, ]) > 0) {
  mplot <- migration_path_timing_plot(
    migr_path = migr_cu,
    nuseds_data = nuseds_cu,
    cu_boundary = cu_boundary_i,
    migr_daily_all = migr_daily_all,
    cu_i = cu_i,
    timing = cu_timing_i,
    rcp = rcp_pick,
    period_choose = c("1981-2010", "2041-2060")
  )
  mplot
}
```

**Fig. S2.8.** Projected daily stream temperatures overlaid with upstream run timing and spawn timing (a), and upstream migration route from the river mouth to NuSEDS spawning sites (b). Uncertainty bands around projected daily temperatures represent 10-90% quantile ranges across global climate models, while the bolded line represents the ensemble mean across global climate models.

## Marine

### Marine Indicators Summary

**Table S2.6.** Marine indicator values for the selected Conservation Unit (CU), compared against all CUs and species. Includes projected sea surface temperature (SSTproj) and warming rate (SSTrate) in nearshore marine areas during the 6-month ocean entry period (Wilson & Peacock 2025) averaged across the CU's Marine Adaptive Zone (MAZ), and coastal cumulative human impacts (CImpact) derived from CMIP6 projections and ROMS datasets.

```{r marine indicators table}
generate_cvis_summary_table(
  all_std_long_baseline = all_std_long_baseline,
  cu_code = cu_i,
  group_by = "species",
  categories = "mar"
)
```

### Marine Indicators Values for Marine Adaptive Zones

```{r marine plot, echo=FALSE, out.width="100%", fig.width=12, fig.height=9}
if (!exists("shared_marine_plot", envir = .GlobalEnv)) {
  # Load only when generating the plot for the first time
  if (!exists("maz_all")) {
    load(get_latest_file(paths$marine, "marine_stats.Rdata"))
  }
  if (!exists("MAZ")) {
    load(file.path(paths$marine, "MAZ.Rds"))
  }
  
  # Generate the combined marine plot
  p_mar <- combined_maz_marine_plot(
    maz_all = maz_all,
    MAZ = MAZ
  )
  assign("shared_marine_plot", p_mar, envir = .GlobalEnv)
}

shared_marine_plot
```

**Fig. S2.9.** Mean indicator values for projected sea surface temperature, rate of sea surface temperature change, and cumulative habitat impacts for Marine Adaptive Zones (MAZs) of British Columbia (top panes) and the spatial distribution of indicator values within the Georgia Strait MAZ (bottom pane).

## Sensitivity Analysis {.tabset .tabset-fade}

### Score Sensitivity

```{r sensitivity plot scores, fig.height = 8, fig.width = 11, out.width="100%"}
plot_cu_sensitivity_scores(
  overall_sensitivity = overall_sensitivity,
  cu_code = cu_i
) + labs(title = NULL, subtitle = NULL)
```

**Fig. S2.10.** Sensitivity of vulnerability scores (scaled from 0 to 100) for the selected Conservation Unit (CU) across different scoring methodologies, emission scenarios, and global climate models. The dotted line represents the value for the selected CU under the default scenario, while orange diamonds represent change in values by scenario. The violins show the distributions of score deviations across all Fraser CUs.

### Indicator Sensitivity

```{r sensitivity plot indicators, fig.height = 7.5, fig.width = 11, out.width="100%"}
plot_cu_sensitivity_indicators(
  overall_sensitivity = overall_sensitivity,
  tbl_indicators = tbl_indicators,
  cu_code = cu_i
) + labs(title = NULL, subtitle = NULL)
```

**Fig. S2.11.** Sensitivity of raw climate indicator values across climate projection scenarios and global climate models for the selected Conservation Unit (CU). The dotted line represents the value for the selected CU under the default scenario, while orange diamonds represent change in values by scenario. Violins show the distribution of values for that scenario across all Fraser CUs.
