---
title: "Climate Vulnerability Indicators for Salmon: Supplemental Data Report"
subtitle: "Interactive Results by Indicator, Category, and Overall Scoring"
author: "DFO Science"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: 
  html_document:
    toc: false
    theme: flatly
    css: "CVIS_style.css"
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = "asis")

# Sourcing CVIS setup environment
library(here)
source(here("code", "0_setup.R"))

# Define global variables for scoping fix in subset_ind_table / plot_std_vs_raw
std_value <- "std_value"
value <- "value"

# Load outputs containing all_std_long and scores_tidy
load(file.path(paths$output, "scoring_results.Rdata"))
load(file.path(paths$params, "indicator_tables.Rdata"))

# Load watershed basins and filter to Fraser River basin for spatial maps
load(file.path(paths$fw, "basins_shp.Rds"))
Fr_basin <- filter(basins, BASIN == "FRASER")

# Load status and timing data if they do not exist
if (!exists("cu_timing_long") || !exists("cu_timing_Fr")) {
  load(file.path(paths$CU, "cu_timing_data.Rdata"))
}
if (!exists("status_data")) {
  load(file.path(paths$CU, "cu_status_data.Rds"))
}


# Clean up datasets to ensure consistency across plots
# Join with standardized table to filter to the baseline downscaling models (dsmodel)
# and filter to keep only standardizing methods that correspond to the baseline functions,
# resolving scoping/duplicate list-column errors.
report_dat <- all_std_long %>%
  left_join(tbl_standardize %>% select(abbrev, dsmodel_baseline_ind = dsmodel_baseline, std_fun), by = c("indicator" = "abbrev")) %>%
  filter(dsmodel == dsmodel_baseline_ind) %>%
  mutate(
    std_fun_base = case_when(
      std_fun %in% c("linear_std", "invlinear_std") ~ "linear",
      TRUE ~ "exponential"
    )
  ) %>%
  filter(is.na(std_method) | std_method == std_fun_base) %>%
  select(-std_fun, -std_fun_base) %>%
  # Fix: Change gcm == "0" to "9" for static & baseline indicators so they are selected by subset_ind_table (which filters for gcm == "9")
  mutate(gcm = if_else(gcm == "0", "9", gcm))
```

<div class="row">
  <div class="col-md-12">
    <p class="lead">
      This interactive supplemental report captures the Climate Vulnerability Indicators for Salmon (CVIS) results for all Conservation Units (CUs) in the Fraser River basin. It presents results by individual indicator, category averages, and overall vulnerability scores.
    </p>
    <p>
      Use the control panel below to select the **Representative Concentration Pathway (RCP)** climate scenario and the **projection Time Period** to dynamically update the heatmaps, maps, and lollipop charts. You can also explore different overall scoring aggregation methods and choose which species to include on the spatial maps.
    </p>
  </div>
</div>

```{r controls}
# Top control panel for interactive inputs
inputPanel(
  selectInput("rcp_pick", label = "Representative Concentration Pathway (RCP):",
              choices = list("RCP 4.5 (Moderate Emissions)" = "45", 
                             "RCP 8.5 (High Emissions)" = "85"), 
              selected = "45"),
  selectInput("period_pick", label = "Time Period:",
              choices = list("Mid-Century (2041-2060)" = "3", 
                             "End-of-Century (2081-2100)" = "5"), 
              selected = "3"),
  selectInput("method_pick", label = "Overall Scoring Method (Detail Tabs):",
              choices = list("Sum-of-Averages (Equal Category Weighting)" = "catavg",
                             "Average of All Indicators (Equal Indicator Weighting)" = "avgall",
                             "Sum-of-Cubic-Means (Emphasizes High Risks)" = "avgcube"),
              selected = "catavg"),
  checkboxGroupInput("species_map_pick", label = "Selected Species (Maps, Tables & Timelines):",
                     choices = c("Chinook", "Coho", "Sockeye", "Chum", "Pink"),
                     selected = c("Chinook", "Coho", "Sockeye"),
                     inline = TRUE)
)
```

# CVIS Results {.tabset .tabset-fade}

## Overall Vulnerability (Tiles)

This tab displays vulnerability score heatmaps. The aggregate view shows full CU scores across all categories and indicators. Individual species tabs display granular indicators alongside category and overall scores.

```{r overall_tiles_ui}
# Nested tabset panel for Aggregate and Species details
tabsetPanel(
  tabPanel("Aggregate (All Species)",
    tags$br(),
    p(strong("Figure 1:"), "Comprehensive vulnerability scores and indicators across all Pacific Salmon Conservation Units (CUs) in the Fraser River basin."),
    renderPlot({
      indicator_cu_tile_plot(
        all_std_long = report_dat %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        scores_tidy = scores_tidy %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        indicators_metadata = tbl_indicators,
        dsmodel_pick = dsmodel_baseline
      )
    }, height = 900, res = 100)
  ),
  tabPanel("Chinook Detail",
    tags$br(),
    renderPlot({
      species_category_tile_plot_from_scores(
        all_std_long = all_std_long %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        scores_tidy = scores_tidy %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        species_pick = "Chinook",
        species_palette = species_palette,
        indicators_metadata = tbl_indicators,
        dsmodel_pick = dsmodel_baseline,
        rank_method = input$method_pick
      )
    }, height = 700, res = 100)
  ),
  tabPanel("Sockeye Detail",
    tags$br(),
    renderPlot({
      species_category_tile_plot_from_scores(
        all_std_long = all_std_long %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        scores_tidy = scores_tidy %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        species_pick = "Sockeye",
        species_palette = species_palette,
        indicators_metadata = tbl_indicators,
        dsmodel_pick = dsmodel_baseline,
        rank_method = input$method_pick
      )
    }, height = 800, res = 100)
  ),
  tabPanel("Coho Detail",
    tags$br(),
    renderPlot({
      species_category_tile_plot_from_scores(
        all_std_long = all_std_long %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        scores_tidy = scores_tidy %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        species_pick = "Coho",
        species_palette = species_palette,
        indicators_metadata = tbl_indicators,
        dsmodel_pick = dsmodel_baseline,
        rank_method = input$method_pick
      )
    }, height = 500, res = 100)
  ),
  tabPanel("Chum Detail",
    tags$br(),
    renderPlot({
      species_category_tile_plot_from_scores(
        all_std_long = all_std_long %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        scores_tidy = scores_tidy %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        species_pick = "Chum",
        species_palette = species_palette,
        indicators_metadata = tbl_indicators,
        dsmodel_pick = dsmodel_baseline,
        rank_method = input$method_pick
      )
    }, height = 400, res = 100)
  ),
  tabPanel("Pink Detail",
    tags$br(),
    renderPlot({
      species_category_tile_plot_from_scores(
        all_std_long = all_std_long %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        scores_tidy = scores_tidy %>% filter(rcp == input$rcp_pick, period_code == input$period_pick),
        species_pick = "Pink",
        species_palette = species_palette,
        indicators_metadata = tbl_indicators,
        dsmodel_pick = dsmodel_baseline,
        rank_method = input$method_pick
      )
    }, height = 400, res = 100)
  )
)
```


```{r helpers}
# Helper function to render reactive lollipop plots for any indicator code
render_cvis_lollipop <- function(ind_code) {
  req(input$rcp_pick, input$period_pick)
  
  # Resolve rcp and period pick based on indicator characteristics (data availability)
  static_inds <- c("cthr", "fwres", "migrdist", "CImpact", "CUstatus", "CUnmat", "hetzyg")
  
  if (ind_code %in% static_inds) {
    rcp_pick <- "0"
    period_pick <- "0"
  } else if (ind_code == "genoff") {
    # Genomic offset is available under RCP 4.5 and 8.5 but only has period code 3
    rcp_pick <- input$rcp_pick
    period_pick <- "3"
  } else {
    rcp_pick <- input$rcp_pick
    period_pick <- input$period_pick
  }
  
  # Retrieve metadata
  ind_row <- tbl_indicators %>% filter(abbrev == ind_code)
  unit_pick <- ind_row$unit[1]
  name_pick <- ind_row$name[1]
  
  # CUstatus is a categorical indicator; plot_lollipop requires use_standardized = TRUE
  use_std <- (ind_code == "CUstatus")
  
  plot_dat_filtered <- report_dat %>%
    filter(rcp == rcp_pick, period_code == period_pick)
  
  p <- plot_lollipop(
    data = plot_dat_filtered,
    indicator_pick = ind_code,
    indicator_name = name_pick,
    indicator_unit = unit_pick,
    use_standardized = use_std
  )
  
  # Add custom subtitle note if static or partially static
  if (ind_code %in% static_inds) {
    p <- p + labs(caption = "Note: This indicator is static (historical status/traits) and does not vary across future scenarios.")
  } else if (ind_code == "genoff") {
    p <- p + labs(caption = "Note: Genomic offset calculations are fixed to Mid-Century (2041-2060) and do not vary by projection period.")
  }
  
  print(p)
}

# Helper function to render reactive spatial maps for freshwater indicators
render_cvis_spatial <- function(ind_code) {
  req(input$rcp_pick, input$period_pick, input$species_map_pick)
  
  # Defensive check to make sure Fr_basin is loaded in the reactive/shiny context
  if (!exists("Fr_basin", envir = .GlobalEnv)) {
    if (!exists("basins", envir = .GlobalEnv)) {
      load(file.path(paths$fw, "basins_shp.Rds"), envir = .GlobalEnv)
    }
    assign("Fr_basin", dplyr::filter(get("basins", envir = .GlobalEnv), BASIN == "FRASER"), envir = .GlobalEnv)
  }
  
  # Defensive check to make sure cu_boundary column renaming is applied in the global environment
  if (exists("cu_boundary", envir = .GlobalEnv)) {
    boundary_temp <- get("cu_boundary", envir = .GlobalEnv)
    if ("Species" %in% names(boundary_temp)) {
      assign("cu_boundary", boundary_temp %>% rename(SPECIES_NAME = Species), envir = .GlobalEnv)
    }
  }
  
  # Resolve rcp and period pick based on data availability
  if (ind_code == "cthr") {
    rcp_pick <- "0"
    period_pick <- "0"
  } else {
    rcp_pick <- input$rcp_pick
    period_pick <- input$period_pick
  }
  
  # Retrieve metadata
  ind_row <- tbl_indicators %>% filter(abbrev == ind_code)
  name_pick <- ind_row$name[1]
  
  # Resolve palette direction (1 for flow8pdelta and favchange, -1 for others)
  pal_dir <- -1
  if (ind_code %in% c("flow8pdelta", "favchange")) {
    pal_dir <- 1
  }
  
  plot_dat_filtered <- report_dat %>%
    filter(rcp == rcp_pick, period_code == period_pick)
  
  p <- spatial_indicator_plot(
    data = plot_dat_filtered,
    cu_boundary = get("cu_boundary", envir = .GlobalEnv),
    outline = get("Fr_basin", envir = .GlobalEnv),
    sp_pick = input$species_map_pick,
    indicator_pick = ind_code,
    indicator_name = name_pick,
    use_standardized = FALSE,
    palette_direction = pal_dir
  )
  
  print(p)
}

# Helper function to render reactive standardized vs raw scatter plots
render_cvis_std_vs_raw <- function(ind_code) {
  req(input$rcp_pick, input$period_pick)
  
  # Resolve rcp and period pick based on indicator characteristics (data availability)
  static_inds <- c("cthr", "fwres", "migrdist", "CImpact", "CUstatus", "CUnmat", "hetzyg")
  
  if (ind_code %in% static_inds) {
    rcp_pick <- "0"
    period_pick <- "0"
  } else if (ind_code == "genoff") {
    rcp_pick <- input$rcp_pick
    period_pick <- "3"
  } else {
    rcp_pick <- input$rcp_pick
    period_pick <- input$period_pick
  }
  
  # Filter data for standard vs raw plot
  plot_dat_filtered <- report_dat %>%
    filter(rcp == rcp_pick, period_code == period_pick)
  
  p <- plot_std_vs_raw(
    data = plot_dat_filtered,
    indicator_pick = ind_code,
    include_histogram = FALSE
  )
  
  print(p)
}
```

## Freshwater Spawning & Rearing (Lollipops & Maps)

Freshwater spawning and rearing indicators represent projected climate changes to streams (temperature and flow), cumulative anthropogenic threats to habitats, and juvenile residency time. Spatial maps show raw indicator values across CU boundaries for selected species.

```{r fw_lollipop_ui}
tabsetPanel(
  tabPanel("Projected August Temp (tw8proj)",
    tags$br(),
    p(strong("Metric:"), " The average of August stream temperatures across all years of the projection period over all accessible streams with modelled habitat potential contained in the CU boundary, weighted by segment length."),
    fluidRow(
      column(5, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("tw8proj") }, height = 800, res = 100)),
      column(4, p(strong("Spatial Distribution:")), renderPlot({ render_cvis_spatial("tw8proj") }, height = 650, res = 100)),
      column(3, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("tw8proj") }, height = 600, res = 100))
    )
  ),
  tabPanel("Rate of Temp Change (tw8rate)",
    tags$br(),
    p(strong("Metric:"), " The decadal rate of increase in August mean stream temperature from a baseline period of 1981-2000 to the projection period averaged over all accessible streams with modelled habitat potential, weighted by segment length."),
    fluidRow(
      column(5, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("tw8rate") }, height = 800, res = 100)),
      column(4, p(strong("Spatial Distribution:")), renderPlot({ render_cvis_spatial("tw8rate") }, height = 650, res = 100)),
      column(3, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("tw8rate") }, height = 600, res = 100))
    )
  ),
  tabPanel("August Flow Decrease (flow8pdelta)",
    tags$br(),
    p(strong("Metric:"), " The change in average August flow from a baseline period of 1981-2010 to the projection period, expressed as a proportion of the average historical flow, weighted by segment length."),
    fluidRow(
      column(5, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("flow8pdelta") }, height = 800, res = 100)),
      column(4, p(strong("Spatial Distribution:")), renderPlot({ render_cvis_spatial("flow8pdelta") }, height = 650, res = 100)),
      column(3, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("flow8pdelta") }, height = 600, res = 100))
    )
  ),
  tabPanel("Winter Flow Increase (flow18pdelta)",
    tags$br(),
    p(strong("Metric:"), " The change in average flow for the months of November, December, and January from a baseline period of 1981-2010 to the projection period, expressed as a proportion of the average historical flow, weighted by segment length."),
    fluidRow(
      column(5, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("flow18pdelta") }, height = 800, res = 100)),
      column(4, p(strong("Spatial Distribution:")), renderPlot({ render_cvis_spatial("flow18pdelta") }, height = 650, res = 100)),
      column(3, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("flow18pdelta") }, height = 600, res = 100))
    )
  ),
  tabPanel("ENM Habitat Change (favchange)",
    tags$br(),
    p(strong("Metric:"), " The change in habitat favourability from the baseline period (1981-2000) to the projection period, calculated as the difference between future and baseline favourability values derived from Environmental Niche Models (ENMs)."),
    fluidRow(
      column(5, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("favchange") }, height = 800, res = 100)),
      column(4, p(strong("Spatial Distribution:")), renderPlot({ render_cvis_spatial("favchange") }, height = 650, res = 100)),
      column(3, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("favchange") }, height = 600, res = 100))
    )
  ),
  tabPanel("Cumulative Threats (cthr)",
    tags$br(),
    p(strong("Metric:"), " A composite index of threats from human activity to stream habitat, calculated using a normalized additive model combining nine threat categories including flow alteration, riparian disturbance, pollution, and sedimentation, weighted by segment length."),
    fluidRow(
      column(5, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("cthr") }, height = 800, res = 100)),
      column(4, p(strong("Spatial Distribution:")), renderPlot({ render_cvis_spatial("cthr") }, height = 650, res = 100)),
      column(3, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("cthr") }, height = 600, res = 100))
    )
  ),
  tabPanel("Freshwater Residency (fwres)",
    tags$br(),
    p(strong("Metric:"), " The duration in days that juvenile salmon reside in freshwater from the start of egg incubation to ocean entry, based on peak spawn timing to peak ocean entry timing plus freshwater age."),
    fluidRow(
      column(8, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("fwres") }, height = 800, res = 100)),
      column(4, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("fwres") }, height = 600, res = 100))
    )
  )
)
```

## Upstream Migration (Lollipops)

Upstream migration indicators evaluate projected changes in stream temperature and discharge during the river migration window for each CU, alongside physical migration distance.

```{r migration_lollipop_ui}
tabsetPanel(
  tabPanel("Projected Migration Temp (migrTproj)",
    tags$br(),
    p(strong("Metric:"), " The average of water temperatures during upstream migration estimated along the migration path from river entry to spawning sites using a moving migration window model matched to CU-specific timing."),
    fluidRow(
      column(8, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("migrTproj") }, height = 800, res = 100)),
      column(4, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("migrTproj") }, height = 600, res = 100))
    )
  ),
  tabPanel("Migration Flow Change (migrQpdelta)",
    tags$br(),
    p(strong("Metric:"), " The proportional change in flow during upstream migration from the baseline period to the projection period, estimated along the migration path from river entry to spawning sites using the moving migration window model."),
    fluidRow(
      column(8, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("migrQpdelta") }, height = 800, res = 100)),
      column(4, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("migrQpdelta") }, height = 600, res = 100))
    )
  ),
  tabPanel("Migration Distance (migrdist)",
    tags$br(),
    p(strong("Metric:"), " The average distance of the migration path from the river mouth to NuSEDS spawning sites across all spawning locations within the CU."),
    fluidRow(
      column(8, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("migrdist") }, height = 800, res = 100)),
      column(4, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("migrdist") }, height = 600, res = 100))
    )
  )
)
```

## Nearshore Marine (Lollipops)

Nearshore marine indicators represent climate changes to sea surface temperature (SST) and cumulative habitat impacts within the Marine Adaptive Zone (MAZ) during the critical ocean entry period.

```{r marine_lollipop_ui}
tabsetPanel(
  tabPanel("Projected SST (SSTproj)",
    tags$br(),
    p(strong("Metric:"), " Projected sea surface temperature throughout the marine adaptive zone (MAZ) for the CU over a 6-month period surrounding the peak month of ocean entry."),
    fluidRow(
      column(8, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("SSTproj") }, height = 800, res = 100)),
      column(4, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("SSTproj") }, height = 600, res = 100))
    )
  ),
  tabPanel("Rate of SST Change (SSTrate)",
    tags$br(),
    p(strong("Metric:"), " The decadal rate of increase in SST from a baseline period to the projection period throughout the marine adaptive zone (MAZ) for the CU over the 6-month ocean entry window."),
    fluidRow(
      column(8, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("SSTrate") }, height = 800, res = 100)),
      column(4, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("SSTrate") }, height = 600, res = 100))
    )
  ),
  tabPanel("Cumulative Marine Impacts (CImpact)",
    tags$br(),
    p(strong("Metric:"), " The average cumulative impact score of nearshore habitats throughout the MAZ for a CU, combining stressors such as aquaculture, fishing, invasive species, nutrient inputs, pollution, and shipping."),
    fluidRow(
      column(8, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("CImpact") }, height = 800, res = 100)),
      column(4, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("CImpact") }, height = 600, res = 100))
    )
  )
)
```

## Demographics (Lollipops, Tables & Timelines)

Demographic indicators capture the baseline conservation status and recent spawner abundance, representing demographic vulnerability. This tab also includes a summary status table and a comparative timing timeline for the selected species.

```{r dem_lollipop_ui}
tabsetPanel(
  tabPanel("WSP Conservation Status (CUstatus)",
    tags$br(),
    p(strong("Metric:"), " Recent Wild Salmon Policy (WSP) conservation status of a CU (Red, Amber, or Green) assessed using decision trees on spawner abundance, trends, and data quality (Pestal et al. 2023)."),
    fluidRow(
      column(8, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("CUstatus") }, height = 800, res = 100)),
      column(4, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("CUstatus") }, height = 600, res = 100))
    )
  ),
  tabPanel("Generational Spawners (CUnmat)",
    tags$br(),
    p(strong("Metric:"), " The geometric average of natural origin spawners over the most recent generation, providing a measure of population scale and demographic risk."),
    fluidRow(
      column(8, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("CUnmat") }, height = 800, res = 100)),
      column(4, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("CUnmat") }, height = 600, res = 100))
    )
  ),
  tabPanel("CU Status Table (5b)",
    tags$br(),
    p("This table displays recent Wild Salmon Policy (WSP) conservation status, spawner abundance, and data types for the selected species."),
    render_gt({
      cu_status_table(status_data, species_filter = input$species_map_pick)
    })
  ),
  tabPanel("Life History Timing Comparison (5b)",
    tags$br(),
    p("This plot compares run timing, spawning, ocean entry, and freshwater migration timing across the Conservation Units of the selected species."),
    renderPlot({
      plot_timing_comparison(
        cu_timing_long = cu_timing_long,
        species_select = input$species_map_pick,
        species_palette = species_palette
      )
    }, height = 850, res = 100)
  )
)
```

## Genetics (Lollipops)

Genetic indicators assess adaptive capacity based on genome-wide genetic diversity (heterozygosity) and the mismatch between current genetics and future environmental needs (genomic offset). Genetics data are available for Chinook, Sockeye, and Coho.

```{r gen_lollipop_ui}
tabsetPanel(
  tabPanel("Genetic Heterozygosity (hetzyg)",
    tags$br(),
    p(strong("Metric:"), " A measure of genome-wide genetic diversity within populations (average proportion of heterozygous sites). Higher values suggest greater raw material for adaptation to climate change."),
    fluidRow(
      column(8, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("hetzyg") }, height = 800, res = 100)),
      column(4, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("hetzyg") }, height = 600, res = 100))
    )
  ),
  tabPanel("Genomic Offset (genoff)",
    tags$br(),
    p(strong("Metric:"), " The Euclidean distance between current population allele frequencies and the frequencies predicted to be optimal under future climate conditions. Represents the genetic change required to track future environmental shifts."),
    fluidRow(
      column(8, p(strong("Relative CU Vulnerability:")), renderPlot({ render_cvis_lollipop("genoff") }, height = 800, res = 100)),
      column(4, p(strong("Standardized vs Raw:")), renderPlot({ render_cvis_std_vs_raw("genoff") }, height = 600, res = 100))
    )
  )
)
```
