---
title: "Climate Vulnerability Indicators for Salmon: Visual Results Overview"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:  
    toc: true  
    number-sections: true  
    toc_float: true  
    toc_collapsed: true
    toc_depth: 3
    theme: flatly
    css: "scientific_style.css"
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = "asis")

# Base setup
library(here)
source(here("code", "0_setup.R"))

# Load outputs containing all_std_long, scores_tidy, and sensitivity results
load(file.path(paths$output, "scoring_results.Rdata"))
load(file.path(paths$params, "indicator_tables.Rdata"))
load(file.path(paths$output, "sensitivity_analysis.Rdata"))

# Load spatial boundaries
load(file.path(paths$marine, "MAZ.Rds"))
load(file.path(paths$fw, "basins_shp.Rds"))
Fr_basin <- dplyr::filter(basins, BASIN == "FRASER")

# Load marine statistics for MAZ plotting
mar_file <- get_latest_file(paths$marine, "marine_stats.Rdata")
load(mar_file) # loads maz_all
```

# Overview

This report provides a visual overview of key results from the Climate Vulnerability Indicators for Salmon (CVIS) framework. It visualizes the cumulative vulnerability profiles across all evaluated Conservation Units (CUs), maps the spatial distribution of risk in freshwater rearing and marine adaptive zones, and evaluates the sensitivity of vulnerability rankings to mathematical assumptions and modeling options.

---

# Integrated Vulnerability Portfolio

The heatmap below illustrates the standardized scores (ranging from 0.0 to 1.0) for individual indicators, category-level vulnerability scores, and the integrated overall score (baseline average) across all 50 CUs. CUs are grouped by species and Stock Assessment Group/SMU (Salmon Management Unit) identifiers, sorted by their overall vulnerability.

```{r portfolio_heatmap, fig.height = 11, fig.width = 10, out.width = "100%", fig.align = "center"}
indicator_cu_tile_plot(
  all_std_long = all_std_long_baseline,
  scores_tidy = scores_tidy_baseline
)
```

**Figure 1.** Comparative portfolio heatmap of climate vulnerability indicators across all Fraser River basin Conservation Units. Cells represent standardized scores where 0.0 (cool blue) indicates lowest vulnerability/risk and 1.0 (warm red) indicates highest vulnerability/risk. Category scores (Demographics, Spawning & Rearing, Genetics, Nearshore Marine, Upstream Migration) and overall vulnerability scores are shown alongside individual indicators.

---

# Interspecific Vulnerability Distributions

We compare the density distribution and median values of the final integrated spawning and rearing vulnerability scores under the baseline, grouped by Pacific salmon species (Chinook, Sockeye, Coho, Chum, and Pink). This comparison shows how species life history traits (such as freshwater rearing residency duration) influence relative vulnerability.

```{r species_violins, fig.height = 5, fig.width = 8.5, out.width = "90%", fig.align = "center"}
plot_cvis_vulnerability_violins(scores_tidy_baseline)
```

**Figure 2.** Integrated spawning and rearing vulnerability scores across evaluated species. Violins visualize the density distribution and spread of scores within each species group under the baseline scenario.

---

# Spatial Distribution of Freshwater Spawning & Rearing Risk

The map below illustrates the spatial distribution of the aggregated freshwater spawning and rearing category score for Chinook salmon CUs across the Fraser River basin. Spawning and rearing stream networks within each CU boundary are colored based on their standardized score (the length-weighted average of the underlying freshwater indicators).

```{r freshwater_risk_map, fig.height = 9, fig.width = 10, out.width = "100%", fig.align = "center"}
spatial_fw_rearing_indicators_plot(
  data = all_std_long_baseline,
  cu_boundary = cu_boundary,
  outline = Fr_basin,
  species_pick = "Chinook"
)
```

**Figure 3.** Basin-wide spatial distribution of standardized freshwater spawning and rearing vulnerability scores for Chinook salmon. Higher scores (red/orange) represent stream networks with elevated cumulative vulnerability, while lower scores (blue) represent potential physical and thermal refugia.

---

# Marine Exposure Profiles by Marine Adaptive Zone

This section displays the regional marine vulnerability profiles across Marine Adaptive Zones (MAZs) representing oceanographic domains utilized by juvenile salmon during post-entry migration. 

```{r marine_maz_profiles, fig.height = 9, fig.width = 11, out.width = "100%", fig.align = "center"}
combined_maz_marine_plot(maz_all, MAZ)
```

**Figure 4.** Geographic boundaries of British Columbia Marine Adaptive Zones (left) and lollipop plots (right) of standardized marine indicators (sea surface temperature projections [SSTproj], SST warming rates [SSTrate], and cumulative human impacts [CImpact]) across each MAZ. Points represent MAZ-level averages, and error bars show spatial variation.

---

# Vulnerability Score Sensitivity Analysis

Vulnerability rankings and overall scores depend on mathematical assumptions and modeling options. This section evaluates how scores respond to alternative mathematical aggregation methods and details overall score deviations across sources of modeling variation.

## Comparison of Mathematical Aggregation Methods

The heatmap below compares overall and category vulnerability scores calculated using three mathematical aggregation methods: average (arithmetic mean), cube-mean (power mean with $p=3$ to weight extreme hazard scores), and a red flag count (sum of indicators exceeding a standardized score of 0.8).

```{r methods_comparison, fig.height = 10, fig.width = 9.5, out.width = "100%", fig.align = "center"}
plot_methods_compare_tile(scores_tidy_baseline)
```

**Figure 5.** Comparison of category-level and overall vulnerability scores across average, cube-mean, and red flag count aggregation methods. Shows how ranking structures respond to mathematical weighting assumptions.

## Sources of Modeling Variation

The boxplots below illustrate deviations in overall vulnerability scores for each CU resulting from four primary sources of model variation: Global Climate Model selection, emissions scenario, standardization curves, and indicator weighting schemes.

```{r score_deviations, fig.height = 8.5, fig.width = 9, out.width = "95%", fig.align = "center"}
plot_score_deviations(overall_sensitivity$deviations)
```

**Figure 6.** Quantitative sensitivity analysis of overall vulnerability scores across sources of modeling variation. Relative spread indicates which modeling choice contributes the greatest score variance.
