---
title: "Climate Vulnerability Indicator System - CU Explorer"
author: "CVIS Project"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: false
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load-packages, include=FALSE}
library(tidyverse)
library(reactable)
library(htmltools)
library(here)
```

```{r load-data, include=FALSE}
################################################################################
# LOAD DATA
################################################################################

# Find most recent standardized indicators CSV
csv_files <- list.files(
  here("processed_data"),
  pattern = "\\d{4}-\\d{2}-\\d{2}_standardized_indicators\\.csv$",
  full.names = TRUE
)

if (length(csv_files) == 0) {
  stop("No standardized indicators CSV found in processed_data/")
}

# Get most recent by date in filename
csv_dates <- basename(csv_files) %>%
  str_extract("^\\d{4}-\\d{2}-\\d{2}") %>%
  as.Date()

most_recent_csv <- csv_files[which.max(csv_dates)]
cat("Loading:", basename(most_recent_csv), "\n")

all_flat_std <- read_csv(most_recent_csv, show_col_types = FALSE)
```

```{r prepare-data, include=FALSE}
################################################################################
# PREPARE DATA
################################################################################

# Get unique species and CUs (avoid duplicates)
species_list <- sort(unique(all_flat_std$SPECIES_NAME))

cu_list <- all_flat_std %>%
  filter(rcp == "45", period_code == 3) %>%  # Filter first
  distinct(FULL_CU_IN, CVIS_NAME, SPECIES_NAME) %>%  # Then get unique
  arrange(SPECIES_NAME, CVIS_NAME)

# Identify indicator columns
raw_indicators <- c("favchange_mean", "CT_mean", "tw8rate_mean", "tw8proj_mean",
                    "lowQpdelta_mean", "highQpdelta_mean", "fwres",
                    "migrT_mean", "migrQ_mean", "migrdist",
                    "SSTproj_mean", "SSTrate_mean", "CImpact_mean",
                    "CUstatus", "CUnmat", "hetzyg_mean", "genoff_mean")

std_indicators <- paste0("std_", raw_indicators)

# Filter to available indicators
raw_indicators <- raw_indicators[raw_indicators %in% names(all_flat_std)]
std_indicators <- std_indicators[std_indicators %in% names(all_flat_std)]

# Create indicator names
indicator_names <- c(
  "favchange_mean" = "ENM Favourability Change",
  "CT_mean" = "Cumulative Threats (FW)",
  "tw8rate_mean" = "Aug Temperature Rate",
  "tw8proj_mean" = "Aug Temperature Projection",
  "lowQpdelta_mean" = "Aug Flow Change",
  "highQpdelta_mean" = "Nov-Jan Flow Change",
  "fwres" = "FW Residency Time",
  "migrT_mean" = "Migration Temperature",
  "migrQ_mean" = "Migration Discharge Change",
  "migrdist" = "Migration Distance",
  "SSTproj_mean" = "SST Projection",
  "SSTrate_mean" = "SST Rate",
  "CImpact_mean" = "Cumulative Impacts (Marine)",
  "CUstatus" = "WSP Status",
  "CUnmat" = "Mature Individuals",
  "hetzyg_mean" = "Heterozygosity",
  "genoff_mean" = "Genomic Offset"
)

cat("✓ Data prepared for", nrow(cu_list), "unique CUs\n")
```

## Conservation Unit Explorer

Select a Conservation Unit to view its climate vulnerability indicators compared to species-wide and all-CU averages.

**Table Structure:**
- **CU Values**: Raw indicator values for the selected CU (mean and GCM range)
- **Species Average**: Average across all CUs of the same species (mean and GCM range)
- **All-CU Average**: Average across all Conservation Units (mean and GCM range)
- **Standardized (0-1)**: Vulnerability scores for CU, species, and all-CU comparisons

**Color coding:** <span style="color: #28a745; font-weight: bold;">● Green (0.0-0.33)</span> = Low vulnerability | <span style="color: #ffc107; font-weight: bold;">● Yellow (0.33-0.67)</span> = Medium | <span style="color: #dc3545; font-weight: bold;">● Red (0.67-1.0)</span> = High

```{r filter-controls, echo=FALSE}
################################################################################
# SPECIES FILTER BUTTONS AND CU DROPDOWN
################################################################################

tags$div(
  style = "margin: 20px 0; padding: 20px; background-color: #f8f9fa; border-radius: 5px;",
  
  # Species buttons
  tags$div(
    style = "margin-bottom: 20px;",
    tags$strong("Filter by Species:", style = "display: block; margin-bottom: 10px; font-size: 16px;"),
    tags$div(
      id = "species-buttons",
      style = "display: flex; flex-wrap: wrap; gap: 10px;",
      
      tags$button(
        "All Species",
        class = "species-btn active-btn",
        onclick = "filterBySpecies(''); return false;",
        style = "padding: 10px 20px; background-color: #007bff; color: white; border: none; border-radius: 4px; cursor: pointer; font-size: 14px; font-weight: 600;"
      ),
      
      lapply(species_list, function(sp) {
        tags$button(
          sp,
          class = "species-btn",
          `data-species` = sp,
          onclick = sprintf("filterBySpecies('%s'); return false;", sp),
          style = "padding: 10px 20px; background-color: #6c757d; color: white; border: none; border-radius: 4px; cursor: pointer; font-size: 14px; font-weight: 500;"
        )
      })
    )
  ),
  
  # CU dropdown
  tags$div(
    tags$label(
      "Select Conservation Unit:",
      `for` = "cu-selector",
      style = "display: block; margin-bottom: 8px; font-weight: 600; font-size: 16px;"
    ),
    tags$select(
      id = "cu-selector",
      onchange = "updateIndicatorTable(); return false;",
      style = "width: 100%; max-width: 600px; padding: 10px; border: 1px solid #ced4da; border-radius: 4px; font-size: 15px; cursor: pointer;",
      tags$option("-- Select a Conservation Unit --", value = "", selected = "selected"),
      lapply(seq_len(nrow(cu_list)), function(i) {
        tags$option(
          sprintf("%s - %s", cu_list$FULL_CU_IN[i], cu_list$CVIS_NAME[i]),
          value = cu_list$FULL_CU_IN[i],
          `data-species` = cu_list$SPECIES_NAME[i],
          `data-cvis` = cu_list$CVIS_NAME[i]
        )
      })
    )
  )
)
```

```{r prepare-indicator-table, include=FALSE}
################################################################################
# PREPARE INDICATOR TABLE DATA (ONE ROW PER INDICATOR PER CU)
################################################################################

# STEP 1: Aggressively filter and deduplicate - ONE row per CU only
base_data <- all_flat_std %>%
  filter(rcp == "45", period_code == 3) %>%
  group_by(FULL_CU_IN) %>%
  slice(1) %>%  # Take only first row per CU
  ungroup()

cat("✓ Base data:", nrow(base_data), "unique CUs (RCP 45, Period 3)\n")

# Verify no duplicates in base data
dup_check <- base_data %>% 
  group_by(FULL_CU_IN) %>% 
  filter(n() > 1)

if (nrow(dup_check) > 0) {
  stop("Still have duplicates in base data!")
} else {
  cat("✓ Confirmed: No duplicate CUs in base data\n")
}

# STEP 2: Calculate species-wide stats (one row per species)
species_stats <- base_data %>%
  group_by(SPECIES_NAME) %>%
  summarise(
    across(
      c(all_of(raw_indicators), 
        matches("_qlowgcm$"), 
        matches("_qhighgcm$"),
        all_of(std_indicators)),
      ~mean(.x, na.rm = TRUE),
      .names = "sp_{.col}"
    ),
    .groups = "drop"
  )

# STEP 3: Calculate all-CU stats (one row total)
allcu_stats <- base_data %>%
  summarise(
    across(
      c(all_of(raw_indicators), 
        matches("_qlowgcm$"), 
        matches("_qhighgcm$"),
        all_of(std_indicators)),
      ~mean(.x, na.rm = TRUE),
      .names = "allcu_{.col}"
    )
  )

# STEP 4: Join CU data with species and all-CU stats
joined_data <- base_data %>%
  left_join(species_stats, by = "SPECIES_NAME") %>%
  bind_cols(allcu_stats)

cat("✓ Joined species and all-CU statistics\n")

# STEP 5: Pivot to long format - one row per CU per indicator
indicator_rows <- list()

for (i in seq_along(raw_indicators)) {
  raw_col <- raw_indicators[i]
  std_col <- std_indicators[i]
  indicator_name <- indicator_names[[raw_col]] %||% raw_col
  
  # Column names for this indicator
  cu_low_col <- paste0(raw_col, "_qlowgcm")
  cu_high_col <- paste0(raw_col, "_qhighgcm")
  sp_mean_col <- paste0("sp_", raw_col)
  sp_low_col <- paste0("sp_", raw_col, "_qlowgcm")
  sp_high_col <- paste0("sp_", raw_col, "_qhighgcm")
  allcu_mean_col <- paste0("allcu_", raw_col)
  allcu_low_col <- paste0("allcu_", raw_col, "_qlowgcm")
  allcu_high_col <- paste0("allcu_", raw_col, "_qhighgcm")
  sp_std_col <- paste0("sp_", std_col)
  allcu_std_col <- paste0("allcu_", std_col)
  
  # Extract this indicator for all CUs - convert everything to character for consistency
  indicator_df <- joined_data %>%
    transmute(
      FULL_CU_IN = FULL_CU_IN,
      indicator = indicator_name,
      cu_mean = as.character(.data[[raw_col]]),
      cu_qlowgcm = if (cu_low_col %in% names(joined_data)) as.character(.data[[cu_low_col]]) else NA_character_,
      cu_qhighgcm = if (cu_high_col %in% names(joined_data)) as.character(.data[[cu_high_col]]) else NA_character_,
      sp_mean = if (sp_mean_col %in% names(joined_data)) as.character(.data[[sp_mean_col]]) else NA_character_,
      sp_qlowgcm = if (sp_low_col %in% names(joined_data)) as.character(.data[[sp_low_col]]) else NA_character_,
      sp_qhighgcm = if (sp_high_col %in% names(joined_data)) as.character(.data[[sp_high_col]]) else NA_character_,
      allcu_mean = if (allcu_mean_col %in% names(joined_data)) as.character(.data[[allcu_mean_col]]) else NA_character_,
      allcu_qlowgcm = if (allcu_low_col %in% names(joined_data)) as.character(.data[[allcu_low_col]]) else NA_character_,
      allcu_qhighgcm = if (allcu_high_col %in% names(joined_data)) as.character(.data[[allcu_high_col]]) else NA_character_,
      std_cu_mean = as.numeric(.data[[std_col]]),
      std_sp_mean = if (sp_std_col %in% names(joined_data)) as.numeric(.data[[sp_std_col]]) else NA_real_,
      std_allcu_mean = if (allcu_std_col %in% names(joined_data)) as.numeric(.data[[allcu_std_col]]) else NA_real_
    )
  
  indicator_rows[[i]] <- indicator_df
}

# Combine all indicators - now all have consistent types
indicator_table_data <- bind_rows(indicator_rows)

# Final deduplication check
indicator_table_data <- indicator_table_data %>%
  group_by(FULL_CU_IN, indicator) %>%
  slice(1) %>%
  ungroup()

cat("✓ Final table:", nrow(indicator_table_data), "total rows\n")
cat("  (", length(unique(indicator_table_data$FULL_CU_IN)), "CUs ×", 
    length(unique(indicator_table_data$indicator)), "indicators)\n")

# Verify exactly one row per CU-indicator combination
duplicates <- indicator_table_data %>%
  group_by(FULL_CU_IN, indicator) %>%
  filter(n() > 1)

if (nrow(duplicates) > 0) {
  stop("ERROR: Found ", nrow(duplicates), " duplicate CU-indicator combinations!")
} else {
  cat("✓ SUCCESS: Each CU has exactly one row per indicator\n")
}

# For testing: show sample for one CU
sample_cu <- indicator_table_data %>% 
  filter(FULL_CU_IN == "CK-09") %>% 
  nrow()
if (sample_cu > 0) {
  cat("  Example: CK-09 has", sample_cu, "indicator rows\n")
}
```

```{r indicator-display, echo=FALSE}
################################################################################
# CU INFO AND INDICATOR TABLE
################################################################################

tags$div(
  id = "cu-display-container",
  style = "margin-top: 30px; display: none;",
  
  # CU header info
  tags$div(
    style = "background-color: #e3f2fd; padding: 20px; border-left: 4px solid #2196F3; border-radius: 4px; margin-bottom: 25px;",
    tags$h3(id = "cu-title", style = "margin: 0 0 10px 0; color: #1976D2;"),
    tags$p(id = "cu-subtitle", style = "margin: 0; font-size: 15px; color: #555;")
  ),
  
  # Indicator table - pre-rendered with all data
  reactable(
    indicator_table_data,
    elementId = "indicator-table",
    
    columns = list(
      FULL_CU_IN = colDef(show = FALSE),  # Hidden filter column
      
      indicator = colDef(
        name = "Indicator",
        minWidth = 220,
        sticky = "left",
        style = list(fontWeight = 600, backgroundColor = "#f8f9fa")
      ),
      
      # CU values group
      cu_mean = colDef(
        name = "Mean",
        minWidth = 120,
        cell = function(value) {
          if (is.na(value) || value == "NA") return("—")
          # Try to format as number, otherwise show as-is (for CUstatus)
          num_val <- suppressWarnings(as.numeric(value))
          if (!is.na(num_val)) {
            sprintf("%.3f", num_val)
          } else {
            value
          }
        }
      ),
      cu_qlowgcm = colDef(
        name = "Low GCM",
        minWidth = 100,
        cell = function(value) {
          if (is.na(value) || value == "NA") return("—")
          num_val <- suppressWarnings(as.numeric(value))
          if (!is.na(num_val)) sprintf("%.3f", num_val) else "—"
        },
        style = list(color = "#666", fontSize = "13px")
      ),
      cu_qhighgcm = colDef(
        name = "High GCM",
        minWidth = 100,
        cell = function(value) {
          if (is.na(value) || value == "NA") return("—")
          num_val <- suppressWarnings(as.numeric(value))
          if (!is.na(num_val)) sprintf("%.3f", num_val) else "—"
        },
        style = list(color = "#666", fontSize = "13px")
      ),
      
      # Species average group
      sp_mean = colDef(
        name = "Mean",
        minWidth = 120,
        cell = function(value) {
          if (is.na(value) || value == "NA") return("—")
          num_val <- suppressWarnings(as.numeric(value))
          if (!is.na(num_val)) sprintf("%.3f", num_val) else "—"
        }
      ),
      sp_qlowgcm = colDef(
        name = "Low GCM",
        minWidth = 100,
        cell = function(value) {
          if (is.na(value) || value == "NA") return("—")
          num_val <- suppressWarnings(as.numeric(value))
          if (!is.na(num_val)) sprintf("%.3f", num_val) else "—"
        },
        style = list(color = "#666", fontSize = "13px")
      ),
      sp_qhighgcm = colDef(
        name = "High GCM",
        minWidth = 100,
        cell = function(value) {
          if (is.na(value) || value == "NA") return("—")
          num_val <- suppressWarnings(as.numeric(value))
          if (!is.na(num_val)) sprintf("%.3f", num_val) else "—"
        },
        style = list(color = "#666", fontSize = "13px")
      ),
      
      # All-CU average group
      allcu_mean = colDef(
        name = "Mean",
        minWidth = 120,
        cell = function(value) {
          if (is.na(value) || value == "NA") return("—")
          num_val <- suppressWarnings(as.numeric(value))
          if (!is.na(num_val)) sprintf("%.3f", num_val) else "—"
        }
      ),
      allcu_qlowgcm = colDef(
        name = "Low GCM",
        minWidth = 100,
        cell = function(value) {
          if (is.na(value) || value == "NA") return("—")
          num_val <- suppressWarnings(as.numeric(value))
          if (!is.na(num_val)) sprintf("%.3f", num_val) else "—"
        },
        style = list(color = "#666", fontSize = "13px")
      ),
      allcu_qhighgcm = colDef(
        name = "High GCM",
        minWidth = 100,
        cell = function(value) {
          if (is.na(value) || value == "NA") return("—")
          num_val <- suppressWarnings(as.numeric(value))
          if (!is.na(num_val)) sprintf("%.3f", num_val) else "—"
        },
        style = list(color = "#666", fontSize = "13px")
      ),
      
      # Standardized values
      std_cu_mean = colDef(
        name = "CU",
        minWidth = 90,
        format = colFormat(digits = 3),
        na = "—",
        style = function(value) {
          if (is.na(value)) return(list(color = "#999"))
          
          color <- if (value <= 0.33) {
            "#28a745"  # Green
          } else if (value <= 0.67) {
            "#ffc107"  # Yellow
          } else {
            "#dc3545"  # Red
          }
          
          list(color = color, fontWeight = "bold")
        }
      ),
      std_sp_mean = colDef(
        name = "Species",
        minWidth = 90,
        format = colFormat(digits = 3),
        na = "—",
        style = function(value) {
          if (is.na(value)) return(list(color = "#999"))
          
          color <- if (value <= 0.33) {
            "#28a745"
          } else if (value <= 0.67) {
            "#ffc107"
          } else {
            "#dc3545"
          }
          
          list(color = color, fontWeight = "bold")
        }
      ),
      std_allcu_mean = colDef(
        name = "All CUs",
        minWidth = 90,
        format = colFormat(digits = 3),
        na = "—",
        style = function(value) {
          if (is.na(value)) return(list(color = "#999"))
          
          color <- if (value <= 0.33) {
            "#28a745"
          } else if (value <= 0.67) {
            "#ffc107"
          } else {
            "#dc3545"
          }
          
          list(color = color, fontWeight = "bold")
        }
      )
    ),
    
    # Column groups
    columnGroups = list(
      colGroup(name = "CU Values", columns = c("cu_mean", "cu_qlowgcm", "cu_qhighgcm")),
      colGroup(name = "Species Average", columns = c("sp_mean", "sp_qlowgcm", "sp_qhighgcm")),
      colGroup(name = "All-CU Average", columns = c("allcu_mean", "allcu_qlowgcm", "allcu_qhighgcm")),
      colGroup(name = "Standardized (0-1)", columns = c("std_cu_mean", "std_sp_mean", "std_allcu_mean"))
    ),
    
    defaultPageSize = 20,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(10, 20, 50),
    
    striped = TRUE,
    bordered = TRUE,
    highlight = TRUE,
    searchable = TRUE,
    
    theme = reactableTheme(
      borderColor = "#dfe2e5",
      stripedColor = "#f6f8fa",
      highlightColor = "#e3f2fd",
      cellPadding = "8px 12px",
      style = list(
        fontFamily = "-apple-system, BlinkMacSystemFont, 'Segoe UI', Roboto, Oxygen, Ubuntu, sans-serif",
        fontSize = "14px"
      ),
      headerStyle = list(
        backgroundColor = "#f8f9fa",
        fontWeight = 600
      )
    )
  )
)
```

```{js filter-script}
// Initialize when DOM loads
document.addEventListener('DOMContentLoaded', function() {
  console.log('CU Explorer initialized');
});

/**
 * Filter CU dropdown by species
 */
function filterBySpecies(species) {
  var dropdown = document.getElementById('cu-selector');
  if (!dropdown) return;
  
  // Update button styling
  document.querySelectorAll('.species-btn').forEach(function(btn) {
    btn.classList.remove('active-btn');
    if ((species === '' && btn.textContent === 'All Species') ||
        (species !== '' && btn.getAttribute('data-species') === species)) {
      btn.classList.add('active-btn');
    }
  });
  
  // Filter dropdown options
  Array.from(dropdown.options).forEach(function(opt) {
    if (opt.value === '') {
      opt.style.display = '';  // Always show placeholder
    } else {
      var optSpecies = opt.getAttribute('data-species');
      var isVisible = (species === '' || optSpecies === species);
      opt.style.display = isVisible ? '' : 'none';
    }
  });
  
  // Reset dropdown selection
  dropdown.value = '';
  
  // Hide table when filtering
  document.getElementById('cu-display-container').style.display = 'none';
}

/**
 * Update indicator table when CU is selected
 */
function updateIndicatorTable() {
  var dropdown = document.getElementById('cu-selector');
  var cuId = dropdown.value;
  
  var container = document.getElementById('cu-display-container');
  var title = document.getElementById('cu-title');
  var subtitle = document.getElementById('cu-subtitle');
  
  // Hide if no CU selected
  if (!cuId) {
    container.style.display = 'none';
    return;
  }
  
  // Get selected option details
  var selectedOption = dropdown.options[dropdown.selectedIndex];
  var cvisName = selectedOption.getAttribute('data-cvis');
  var species = selectedOption.getAttribute('data-species');
  
  // Update header
  title.textContent = cvisName;
  subtitle.innerHTML = '<strong>CU ID:</strong> ' + cuId + ' | <strong>Species:</strong> ' + species + 
                       ' | <strong>Scenario:</strong> RCP 4.5, Period 2041-2060';
  
  // Filter the reactable by CU ID
  var table = document.getElementById('indicator-table');
  if (table) {
    var widget = HTMLWidgets.find('#indicator-table');
    if (widget && widget.setFilter) {
      widget.setFilter('FULL_CU_IN', cuId);
    }
  }
  
  // Show container
  container.style.display = 'block';
  
  // Smooth scroll
  setTimeout(function() {
    container.scrollIntoView({ behavior: 'smooth', block: 'start' });
  }, 100);
}
```

```{css custom-styles}
/* Button styles */
.species-btn {
  transition: all 0.2s ease;
}

.species-btn:hover {
  opacity: 0.85;
  transform: translateY(-1px);
  box-shadow: 0 2px 5px rgba(0,0,0,0.2);
}

.species-btn:active {
  transform: translateY(0);
}

.active-btn {
  background-color: #007bff !important;
  font-weight: 600 !important;
}

/* Dropdown styles */
#cu-selector {
  transition: border-color 0.2s;
}

#cu-selector:focus {
  outline: none;
  border-color: #007bff;
  box-shadow: 0 0 0 0.2rem rgba(0,123,255,0.25);
}

#cu-selector option {
  padding: 10px;
}

/* Display container */
#cu-display-container {
  background: white;
  border-radius: 4px;
  padding: 20px;
  box-shadow: 0 1px 3px rgba(0,0,0,0.1);
}

/* Reactable styles */
.reactable {
  font-family: -apple-system, BlinkMacSystemFont, 'Segoe UI', Roboto, Oxygen, Ubuntu, sans-serif;
  font-size: 14px;
}

.reactable .rt-thead {
  background-color: #f8f9fa;
  font-weight: 600;
}

.reactable .rt-th {
  padding: 12px 8px;
  border-bottom: 2px solid #dee2e6;
}
```

---

<div style="margin-top: 40px; padding: 20px; background-color: #f8f9fa; border-radius: 5px;">

### About This Tool

This interactive tool displays climate vulnerability indicators for individual Conservation Units (CUs).

**Data Summary:**
- **Total CUs**: `r nrow(cu_list)`
- **Species**: `r paste(species_list, collapse = ", ")`
- **Indicators**: `r length(raw_indicators)`
- **Scenario**: RCP 4.5, Period 2041-2060 (period code 3)

**Column Group Descriptions:**
- **CU Values**: Raw measurements for the selected Conservation Unit
  - Mean: Average value across GCM models
  - Low/High GCM: Range of values from different climate models
- **Species Average**: Mean values across all CUs of this species
- **All-CU Average**: Mean values across all Conservation Units (baseline for comparison)
- **Standardized (0-1)**: Normalized vulnerability scores where 0 = low vulnerability, 1 = high vulnerability

**Note:** GCM ranges represent uncertainty from different Global Climate Models. NA values indicate data not available for that indicator.

**Data Source:** `r basename(most_recent_csv)`

</div>

---

**Report generated**: `r Sys.time()`
