---
title: "CVIS Indicator Standardization Methods"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

# Functions for standardization of indices, scoring and plotting

# raw standardization function between 0 and 1 using min and max values
linear_std <- function(x, xmin = NA, xmax = NA) {
  if(is.na(xmax)) xmax <- max(x, na.rm = T)
  if(is.na(xmin)) xmin <- min(x, na.rm = T)
  
  x <- ifelse(x < xmin, xmin, x)
  x <- ifelse(x > xmax, xmax, x)
  
  return((x - xmin) / (xmax - xmin))
}

#inverse raw standardization function with 1 corresponding to lowest value
invlinear_std <- function(x, xmin = NA, xmax = NA) {
  if(is.na(xmax)) xmax <- max(x, na.rm = T)
  if(is.na(xmin)) xmin <- min(x, na.rm = T)
  
  y <- rep(NA, length(x))
  
  for(i in 1:length(x)) {
    if(is.na(x[i])) next
    z <- ifelse(x[i] < xmin, xmin, ifelse(x[i] > xmax, xmax, x[i]))
    y[i] <- 1 - (z - xmin) / (xmax - xmin)
  }
  
  return(y)
  # 
  # x <- ifelse(x < xmin, xmin, x)
  # x <- ifelse(x > xmax, xmax, x)
  # 
  # return(1 - (x - xmin) / (xmax - xmin))
}

## asymptotic standardization function for indicators where higher values = greater risk
logarithmic_std <- function(x, lambda = 0.6, xmin = NA, xmax = NA) {
  # Standardize a score
  # x = raw score
  # lambda = rate parameter
  # returns standardized score
  if(is.na(xmax)) xmax <- max(x, na.rm = T)
  if(is.na(xmin)) xmin <- min(x, na.rm = T)
  
  x_std <- (x - xmin) / (xmax - xmin)
  
  return(1 - exp(-lambda * x_std) + exp(-lambda) * x_std)
}

#exponential increase function where higher values = greater risk
exponential_std <- function(x, lambda = 1, xmin = NA, xmax = NA) {

  if(is.na(xmax)) xmax <- max(x, na.rm = T)
  if(is.na(xmin)) xmin <- min(x, na.rm = T)
  
  x_std <- (x - xmin) / (xmax - xmin)
  
  return(exp(lambda * x_std)/exp(lambda) - exp(-lambda)*(1-x_std))
}

# exponential decay standardization function for indicators where higher values = lower risk
decay_std <- function(x, lambda = 0.03, xmin = NA, xmax = NA) {
  # Standardize a score
  # x = raw score
  # lambda = rate parameter
  # returns standardized score
  
  if(is.na(xmax)) xmax <- max(x, na.rm = T)
  if(is.na(xmin)) xmin <- min(x, na.rm = T)
  
  x_std <- (x - xmin) / (xmax - xmin)
  return(exp(-lambda * x_std) - exp(-lambda)*x_std)
}

step_std <- function(x, x1 = 200, x2 = 300) {
  # Standardize a score into 3 categories
  # x = raw score
  # x1 = lowest threshold
  # x2 = high threshold
  y <- rep(NA, length(x))
  
  for(i in 1:length(x)) {
    if(x[i] < x1) y[i] <- 0.33
    if(x[i] >= x1 && x[i] < x2) y[i] <- 0.66
    if(x[i] >= x2) y[i] <- 1
  }
  
  return(y)
}


cat_std <- function(x, x1 = "G", x2 = "A/G", x3 = "A", x4 = "R/A", x5 = "R") {
  # Standardize a score between 0 and 1 based on category values
  y <- rep(NA, length(x))

  for(i in 1:length(x)) {
    if(is.na(x[i])) next
    if(x[i] == x1) y[i] <- 0
    if(x[i] == x2) y[i] <- 0.25
    if(x[i] == x3) y[i] <- 0.5
    if(x[i] == x4) y[i] <- 0.75
    if(x[i] == x5) y[i] <- 1
  }
  return(y)
}

enh_std <- function(x, z) {
  # Standardize a score between 0 and 1 based on category values
  y <- rep(NA, length(x))
  
  for(i in 1:length(x)) {
    if(x[i] == 0) y[i] <- 0
    if(x[i] > 0 && !grepl("Harvest", z[i])) y[i] <- 0.5
    if(x[i] > 0 && grepl("Harvest", z[i]))  y[i] <- 1
  }
  return(y)
}

```


Climate vulnerability indicators are standardized between a scale of 0 and 1 to improve comparability, facilitate interpretation and represent risk based on mechanistic relationships and thresholds. This is done by applying a standardization function to map the raw values to a standardized scale.

The standardization methods were adapted from those developed by Boyce et al (2023) [https://www.nature.com/articles/s41558-022-01437-y].

All standardized values are between 0 and 1, with 0 indicating the lowest risk and 1 indicating the highest risk. There are several different standardization functions that are applied depending on the raw indicator values and the mechanisms affecting risk. Functions can standardize risk based on the range of observed raw values, or using thresholds to represent physiological and other mechanisms.

The functions are:

**Positive standardization functions -** higher raw indicator value corresponds to higher standardized risk

**1. Linear positive standardization function -** raw value is rescaled between 0 to 1 using a linear relationship based on minimum and maximum values.  
equation 1 : y = (x - xmin) / (xmax - xmin)

**2. Logarithmic standardization function -** logarithmic increase in the standardized value up to an asymptote of 1.   
equation 2: y = 1 - exp(-lambda * x_std) + exp(-lambda) * x_std   where x_std represents x rescaled between 0 and 1 using equation 1.

**3. Exponential increase standardization function - ** exponential increase in the standardized value from 0 to 1. 
equation 3: y = exp(lambda * x_std)/exp(lambda) - exp(-lambda)*(1-x_std)   where x_std represents x rescaled between 0 and 1 using equation 1.

**4. Stepwise standardization function -** stepwise increase in the standardized value based on thresholds.

**5. Categorical standardization function -** standardized value based on categories of raw values.

**Negative standardization functions -** higher raw indicator value corresponds to lower standardized risk

**6. Linear negative standardization function -** raw value is rescaled between 0 and 1 using an inverse linear relationship based on minimum and maximum values.
equation 4: y = 1 - (z - xmin) / (xmax - xmin)

**7. Exponential decay standardization function -** exponential decrease in the standardized value from 1 to 0. The inverse of the logarithmic growth standardization function. 
equation 5: y = exp(-lambda * x_std) - exp(-lambda) * x_std   where x_std represents x rescaled between 0 and 1 using equation 4.



## Plots of standardization functions

```{r, echo=FALSE}
## simulation to test lambda values and plot them

std_sim <- as_tibble(c(seq(0, 1, 0.1), seq(2:100))) %>%
  mutate(lambda_1 = logarithmic_std(value, lambda = 1),
         lambda_5 = logarithmic_std(value, lambda = 5),
         lambda_3 = logarithmic_std(value, lambda = 3),
         lambda_10 = logarithmic_std(value, lambda = 10),
         linear = linear_std(value)) %>%
  pivot_longer(cols = c(lambda_1, lambda_3, lambda_5, lambda_10, linear),
               names_to = "function_type", values_to = "std")

stdinv_sim <- as_tibble(c(seq(0, 1, 0.1), seq(2:100))) %>%
  mutate(lambda_10 = decay_std(value, lambda = 10),
         lambda_5 = decay_std(value, lambda = 5),
         lambda_2 = decay_std(value, lambda = 2),
         lambda_1 = decay_std(value, lambda = 1),
         linear_negative = invlinear_std(value)) %>%
  pivot_longer(cols = c(lambda_10, lambda_5, lambda_2, lambda_1, linear_negative),
               names_to = "function_type", values_to = "std")

neginv_sim <- as_tibble(seq(-1,0, 0.01)) %>%
  mutate(lambda_5 = decay_std(value, lambda = 5, xmax = 0),
         lambda_1 = decay_std(value, lambda = 1, xmax = 0),
         lambda_3 = decay_std(value, lambda = 3, xmax = 0),
         lambda_10 = decay_std(value, lambda = 10, xmax = 0),
         linear_negative = invlinear_std(value)) %>%
  pivot_longer(cols = c(lambda_5, lambda_1, lambda_10, lambda_3, linear_negative),
               names_to = "function_type", values_to = "std")

exponential_sim <- as_tibble(c(seq(0, 1, 0.1), seq(2:100))) %>%
  mutate(lambda_1 = exponential_std(value, lambda = 1),
         lambda_3 = exponential_std(value, lambda = 3),
         lambda_10 = exponential_std(value, lambda = 10),
         linear = linear_std(value)) %>%
  pivot_longer(cols = c(lambda_1, lambda_3, lambda_10, linear),
               names_to = "function_type", values_to = "std")



step_sim <- as_tibble(seq(0:400)) %>%
  mutate(step_200_300 = step_std(value, x1 = 200, x2 = 300))


log_p <- ggplot() +
  geom_point(data = std_sim, aes(x = value, y = std, color = function_type)) +
  labs(title = "Positive linear and logarithmic standardization functions") +
  labs(x = "Raw value", y = "Standardized value")

exp_p <- ggplot(exponential_sim) +
   geom_point(aes(x = value, y = std, color = function_type)) +
   labs(title = "Positive linear and exponential increase standardization functions") +
   labs(x = "Raw value", y = "Standardized value")

decay_p <- ggplot() +
  geom_point(data = stdinv_sim, aes(x = value, y = std, color = function_type)) +
  labs(title = "Linear negative and exponential decay standardization functions") +
  labs(x = "Raw value", y = "Standardized value")

decayneg_p <- ggplot() +
  geom_point(data = neginv_sim, aes(x = value, y = std, color = function_type)) +
  labs(title = "Negative values with exponential decay functions") +
  labs(x = "Raw value", y = "Standardized value")

step_p <- ggplot() +
  geom_point(data = step_sim, aes(x = value, y = step_200_300), color = "blue") +
  labs(title = "Stepwise standardization function") +
  labs(x = "Raw value", y = "Standardized value") +
  ylim(0,1)


log_p
exp_p
decay_p
decayneg_p
step_p

```

