---
title: "Climate Vulnerability Indicators for Salmon: Migration Indicators"
output:
  html_document:
    df_print: paged
    number-sections: true
    toc_float: true
    toc_depth: 3
date: today
---

```{r custom theme, echo = FALSE, message = F, warning = F}

library(viridis)
library(gt)        # nice tables
library(patchwork) # plot layouts
library(ggtext)  # plot text colors
library(corrplot)

knitr::opts_chunk$set(message = F, echo = F, warning = F)

theme_cvis <- function(base_size = 14) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = rel(0.85), face = "bold"),
      axis.text = element_text(size = rel(0.70), face = "bold"),
      axis.line = element_line(color = "black"),
      # La légende
      legend.title = element_text(size = rel(0.85), face = "bold"),
      legend.text = element_text(size = rel(0.70), face = "bold"),
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", colour = NA),
      # Les étiquettes dans le cas d'un facetting
      strip.background = element_rect(fill = "#17252D", color = "#17252D"),
      strip.text = element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
    )
}
theme_set(theme_cvis())
```

## Methods

**Spatial analysis** 

## Plots

The following plots show the mean and range of values for environmental change indicators within CU boundaries.For each CU and indicator, the plots show the ensemble mean value for all accessible streams that are modelled as potential spawning or rearing habitat. 

Two types of variation in environmental change are plotted where available:  
**Climate model variation** (thin red bars) - estimated using the minimum and maximum of global climate model means for streams within the CU boundary.

The upstream migration indicators are:

```{r}


migr_indicators <- tribble(
  ~abbrev, ~std_fun, ~name,
  "migrT",          "linear_std",     "Projected temperature during upstream migration",
  "migrQ",          "linear_std",    "Projected discharge during upstream migration",
  "migrTthr19",    "exponential_std", "Average proportion of path above 19 degrees during upstream migration"
)

migr_indicators %>%
  gt()

```



```{r plot function}


# Function to create a lollipop chart
plot_lollipop <- function(data, indicator_pick) {
  # Define suffixes
  stat_suffix <- "wmean"
  sp_suffix <- c("wqlow_sp", "wqhigh_sp")
  gcm_range_suffix <- c("qlow_gcm", "qhigh_gcm")
  
  #take column names that contain prefix with model type
  cols_sub <- names(data)[str_detect(names(data), indicator_pick$abbrev)]
  
  #take names with prefix that also contain stat suffix
  stat_col <- cols_sub[str_detect(cols_sub, paste0(stat_suffix, "$"))]  #must end with wmean
  min_col  <- cols_sub[str_detect(cols_sub,sp_suffix[1])]
  max_col  <- cols_sub[str_detect(cols_sub,sp_suffix[2])]
  min_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[1], collapse = "|"))]
  max_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[2], collapse = "|"))]
  
  id_col <- "CU_NAME"
  sp_col <- "CU_Species"
  
  # Check if mean and ID columns exist
  if (!all(c(id_col, stat_col) %in% names(data))) {
    stop("One or more required columns are missing in the data.")
  }
  
  # Prepare data for plotting
  plot_data <- data %>%
    select(all_of(c(id_col, stat_col, sp_col))) %>%
    rename(mean = !!stat_col, id = !!id_col, sp = !!sp_col)
  
  #palette_colors <- pal_npg("nrc")(5) # 4 colors from Nature Publishing Group palette
  
  # Create a named color palette using unique categories
  unique_sp <- unique(plot_data$sp)
  palette_colors <- pal_npg("nrc")(length(unique_sp))
  names(palette_colors) <- unique_sp
  
  plot_data$color <- palette_colors[plot_data$sp]
  plot_data$label <- paste0("<span style='color:", plot_data$color, "'>", plot_data$sp, "</span>")
  
  
  # Create a unique ordering of id by sp
  ordered_ids <- plot_data %>%
    arrange(sp, id) %>%
    distinct(id) %>%
    pull(id)
  
  # Apply the ordering
  plot_data$id <- factor(plot_data$id, levels = ordered_ids)
  
  # Add min and max if available
  if (length(min_col) == 1 && length(max_col) == 1) {
    plot_data <- plot_data %>%
      mutate(
        min = data[[min_col]],
        max = data[[max_col]]
      )
    has_range <- TRUE
  } else {
    has_range <- FALSE
  }
  
  # add GCM range if available
  if (length(min_gcmcol) == 1 && length(max_gcmcol) == 1) {
    plot_data <- plot_data %>%
      mutate(
        mingcm = data[[min_gcmcol]],
        maxgcm = data[[max_gcmcol]]
      )
    has_gcm <- TRUE
  } else {
    has_gcm <- FALSE
  }
  
  # Create the lollipop chart
  p <- ggplot(plot_data, aes(x = id, color = sp)) +
    scale_fill_manual(values = palette_colors) 
  
  if (has_range) {
    p <- p + geom_segment(aes(xend = id, y = min, yend = max), 
                          color = "darkgrey",
                          linewidth = 2.5)
  }
  
  if (has_gcm) {
    p <- p + geom_segment(aes(xend = id, y = mingcm, yend = maxgcm), 
                          color = "darkred",
                          linewidth = 1)
  }
  
  p +   geom_point(aes(y = mean), size = 2.5) +
    labs(
      title = paste(indicator_pick$name),
      x = "CU",
      y = stat_col,
      color = "Species"
    ) +
    coord_flip() +
    theme_minimal() +
    theme(
      axis.text.y =element_text(size=7)
      
    )
  
}
```


```{r}


spatial_indicator_plot <- function(data,
                                   outline = Fr_basin,
                                   sp_pick = "Chinook", 
                                   indicator_pick, 
                                   palette = "Zissou1",
                                   reversed_pal = FALSE) 
{
  
  data_sp <- filter(data, Species == sp_pick)

  #take column names that contain prefix with model type
  cols_sub <- names(data_sp)[str_detect(names(data_sp), indicator_pick$abbrev)]
  #take names with prefix that also contain stat suffix
  stat_col <- cols_sub[str_detect(cols_sub, paste0(stat_suffix, "$"))]
  
  data_sp <- mutate(data_sp, plot_col = .data[[stat_col]]) 
  
  
# Generate the palette
  palette <- wes_palette("Zissou1", 100, type = "continuous")
  if (reversed_pal) {
    palette <- rev(palette)
  }

 p <- ggplot() +
  geom_sf(data = data_sp, aes(fill = plot_col), alpha = 0.3) +
  scale_fill_gradientn(colours = palette) +
  #geom_label(data = cu_boundary_show, aes(label = CUID), size = 2) +
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) + 
   labs(subtitle = paste(sp_pick, "-", indicator_pick$name),
        legend = indicator_pick$abbrev)
 
 return(p)

}



```




## RCP 45 - 2041 to 2060

```{r}

data <- filter(migr_all_flat, period == "2041-2060", RCP == "45")

### Low flow comparison plot
stat_suffix <- "mean"
gcm_range_suffix <- c("q10", "q90")
#take column names that contain prefix with model type
cols_sub <- names(data)[str_detect(names(data), "migrT_")]

stat_col <- cols_sub[str_detect(cols_sub, paste0(stat_suffix, "$"))]  #must end with wmean
min_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[1], collapse = "|"))]
max_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[2], collapse = "|"))]
  
id_col <- "CU_NAME"
sp_col <- "CU_Species"
  
  # Check if mean and ID columns exist
  if (!all(c(id_col, stat_col) %in% names(data))) {
    stop("One or more required columns are missing in the data.")
  }
  
  # Prepare data for plotting
  plot_data <- data %>%
    select(all_of(c(id_col, stat_col, sp_col, min_gcmcol, max_gcmcol)))
  

```

```{r standardized indicators}

#### plotting standardized indicators

# lollipop plot of all rearing indicators
data <- filter(fwR_all_std, period == "3", RCP == "45")

#get standardized indicator value columns
cols_sub <- names(data)[str_starts(names(data), "std")]

stat_suffix <- "wmean"
gcm_range_suffix <- c("qlow_gcm","qhigh_gcm")

#take names with prefix that also contain stat suffix
stat_col <- cols_sub[str_detect(cols_sub, paste0(stat_suffix, "$"))]  #must end with wmean

min_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[1], collapse = "|"))]
max_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[2], collapse = "|"))]

id_col <- "CU_NAME"
sp_col <- "CU_Species"

# Check if mean and ID columns exist
if (!all(c(id_col, stat_col) %in% names(data))) {
  stop("One or more required columns are missing in the data.")
}



# Prepare data for plotting
plot_data <- data %>%
  select(all_of(c(id_col, stat_col, sp_col, min_gcmcol, max_gcmcol))) %>%
  pivot_longer(cols = starts_with("std"),
               names_to = c("prefix", "indicator", "stat"),
               names_pattern = "^([^_]+)_([^_]+)_(.+)$") %>%
  pivot_wider(names_from = "stat", 
              values_from = "value") 
  
# Create a unique ordering of id by sp
ordered_ids <- plot_data %>%
  arrange(CU_Species, CU_NAME) %>%
  distinct(CU_NAME) %>%
  pull(CU_NAME)

# Apply the ordering
plot_data$CU_NAME <- factor(plot_data$CU_NAME, levels = ordered_ids)
  
# Create a named color palette using unique categories
unique_sp <- unique(plot_data$CU_Species)
palette_colors <- pal_npg("nrc")(length(unique_sp))
names(palette_colors) <- unique_sp

# Create the lollipop chart
p <- ggplot(plot_data, aes(x = CU_NAME)) +
  geom_segment(aes(xend = CU_NAME, y = wmean_qlow_gcm, yend = wmean_qhigh_gcm), 
               color = "darkgrey",
               linewidth = 1) +
  geom_point(aes(y = wmean, color = CU_Species), size = 3) +
  coord_flip() +
  theme_minimal() +
  theme(
       axis.text.y =element_text(size=8),
       axis.text.x =element_text(size=7),
       legend.position = "none"
     ) +
  scale_y_continuous(breaks = c(0,0.5,1), labels = c(0, 0.5, 1)) + 
  facet_grid(.~ indicator) +
  labs(title = "Standardized indicator values for Fraser CUs",
       y = "Standardized value",
       x = NULL)


p

```


```{r period 3 indicator plots}

data <- filter(fwR_all_flat, period == "3", RCP == "45")

cu_boundary_plot <- cu_boundary %>%
  filter(FULL_CU_IN %in% cu_run$FULL_CU_IN) %>%
  left_join(data, by = c("FULL_CU_IN"))
  
for(i in 1:nrow(indicators)) {
  
  p <- plot_lollipop(data, indicators[i,])
  
  p2 <- spatial_indicator_plot(cu_boundary_plot, 
                               sp_pick = "Chinook",
                               indicator_pick = indicators[i,])
  # p3 <- spatial_indicator_plot(cu_boundary_plot, 
  #                              sp_pick = "Coho",
  #                              indicator_pick =indicators[i,])
  # p4 <- spatial_indicator_plot(cu_boundary_plot, 
  #                              sp_pick = "Sockeye-Lake",
  #                              indicator_pick =indicators[i,])

  
  print(p)
  print(p2)
  
}


```

## RCP 45 - 2081 to 2100

```{r period 5}

data <- filter(fwR_all_flat, period == "5", RCP == "45")

for(i in 1:nrow(indicators)) {
  p <- plot_lollipop(data, indicators[i,])
  
  print(p)
}

```

