---
title: "CU Boundary Stream Rearing/Spawning Indicators"
output: html_document
date: "2025-06-06"
---





```{r custom theme, echo = FALSE, message = F, warning = F}

library(viridis)
library(gt)        # nice tables
library(patchwork) # plot layouts

knitr::opts_chunk$set(message = F, echo = F, warning = F)

theme_cvis <- function(base_size = 14) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = rel(0.85), face = "bold"),
      axis.text = element_text(size = rel(0.70), face = "bold"),
      axis.line = element_line(color = "black"),
      # La légende
      legend.title = element_text(size = rel(0.85), face = "bold"),
      legend.text = element_text(size = rel(0.70), face = "bold"),
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", colour = NA),
      # Les étiquettes dans le cas d'un facetting
      strip.background = element_rect(fill = "#17252D", color = "#17252D"),
      strip.text = element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
    )
}
theme_set(theme_cvis())
```


The following plots show the mean and range of values for environmental change indicators within CU boundaries.
For each CU and indicator, the plots show the ensemble mean value for all **accessible** streams. 

Two types of variation in environmental change are plotted where available:  
**Climate model variation** (thin red bars) - estimated using the minimum and maximum of global climate model means for streams within the CU boundary.
**Spatial variation** (thick grey bars) -estimated using the 95% interval of ensemble means for stream segments within the CU boundary. 

The indicators currently plotted are:

```{r}
indicators <- tribble(
  ~abbrev, ~name,
  "ENM_Fav_change", "ENM Change in Favourability",
                      "ct", "Cumulative threats",
                      "Tw8rate", "Rate of change in August Temperature",
                      "Tw8proj", "Projected August Temperature",
                      "Q8pdelta", "Proportional change in August flow (stream model)",
                      "QmonthH", "Historic month of maximum flow",
                      "st8pdelta",  "Proportional change in August flow (station model)")


indicators %>%
  gt()

```





```{r plot function}


indicators <- tribble(
  ~abbrev, ~name,
  "ENM_Fav_change", "ENM Change in Favourability",
                      "ct", "Cumulative threats",
                      "Tw8rate", "Rate of change in August Temperature",
                      "Tw8proj", "Projected August Temperature",
                      "Q8pdelta", "Proportional change in August flow (stream model)",
                      "QmonthH", "Historic month of maximum flow",
                      "st8pdelta",  "Proportional change in August flow (station model)")


# Function to create a lollipop chart
plot_lollipop <- function(data, indicator_pick) {
  # Define suffixes
  stat_suffix <- "wmean"
  sp_suffix <- c("wq025_sp", "wq975_sp")
  gcm_range_suffix <- c("q025_gcm", "min_gcm", "q975_gcm", "max_gcm")
  
  #take column names that contain prefix with model type
  cols_sub <- names(data)[str_detect(names(data), indicator_pick$abbrev)]
  
  #take names with prefix that also contain stat suffix
  stat_col <- cols_sub[str_detect(cols_sub, paste0(stat_suffix, "$"))]  #must end with wmean
  min_col  <- cols_sub[str_detect(cols_sub,sp_suffix[1])]
  max_col  <- cols_sub[str_detect(cols_sub,sp_suffix[2])]
  min_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[1:2], collapse = "|"))]
  max_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[3:4], collapse = "|"))]
  
  id_col <- "FULL_CU_IN"
  sp_col <- "CU_Species"
  
  # Check if mean and ID columns exist
  if (!all(c(id_col, stat_col) %in% names(data))) {
    stop("One or more required columns are missing in the data.")
  }
  
  # Prepare data for plotting
  plot_data <- data %>%
    select(all_of(c(id_col, stat_col, sp_col))) %>%
    rename(mean = !!stat_col, id = !!id_col, sp = !!sp_col)

  #palette_colors <- pal_npg("nrc")(5) # 4 colors from Nature Publishing Group palette


  # Create a named color palette using unique categories
  unique_sp <- unique(plot_data$sp)
  palette_colors <- pal_npg("nrc")(length(unique_sp))
  names(palette_colors) <- unique_sp
  
  plot_data$color <- palette_colors[plot_data$sp]
  plot_data$label <- paste0("<span style='color:", plot_data$color, "'>", plot_data$sp, "</span>")
  
  # Add min and max if available
  if (length(min_col) == 1 && length(max_col) == 1) {
    plot_data <- plot_data %>%
      mutate(
        min = data[[min_col]],
        max = data[[max_col]]
      )
    has_range <- TRUE
  } else {
    has_range <- FALSE
  }
  
  # add GCM range if available
  if (length(min_gcmcol) == 1 && length(max_gcmcol) == 1) {
    plot_data <- plot_data %>%
      mutate(
        mingcm = data[[min_gcmcol]],
        maxgcm = data[[max_gcmcol]]
      )
    has_gcm <- TRUE
  } else {
    has_gcm <- FALSE
  }
  
  # Create the lollipop chart
  p <- ggplot(plot_data, aes(x = reorder(id, sp), color = sp)) +
    scale_fill_manual(values = palette_colors) 
  
  if (has_range) {
    p <- p + geom_segment(aes(xend = id, y = min, yend = max), 
                          color = "darkgrey",
                          linewidth = 2.5)
  }
  
  if (has_gcm) {
    p <- p + geom_segment(aes(xend = id, y = mingcm, yend = maxgcm), 
                          color = "darkred",
                            linewidth = 1)
  }
  
   p +   geom_point(aes(y = mean), size = 3) +
     labs(
      title = paste(indicator_pick$name),
      x = "CU",
      y = stat_col,
      color = "Species"
    ) +
    coord_flip() +
    theme_minimal() +
    theme(
       axis.text.x = element_markdown() # enables HTML rendering
     )

}
```





## RCP 45 - 2041 to 2060

```{r period 3}

fwR_45_3 <- filter(fwR_all_flat, period == "3", RCP == "45")

for(i in 1:nrow(indicators)) {
  p <- plot_lollipop(fwR_45_3, indicators[i,])
  
  print(p)
}


```

## RCP 45 - 2081 to 2100

```{r period 5}

fwR_45_5 <- filter(fwR_all_flat, period == "5", RCP == "45")

for(i in 1:nrow(indicators)) {
  p <- plot_lollipop(fwR_45_5, indicators[i,])
  
  print(p)
}

```

