---
title: "Climate Vulnerability Indicators for Salmon: Rearing/Spawning Indicators"
output:
  html_document:
    df_print: paged
    number-sections: true
    toc_float: true
    toc_depth: 3
date: today
---

```{r custom theme, echo = FALSE, message = F, warning = F}

library(viridis)
library(gt)        # nice tables
library(patchwork) # plot layouts
library(ggtext)  # plot text colors
library(corrplot)

knitr::opts_chunk$set(message = F, echo = F, warning = F)

theme_cvis <- function(base_size = 14) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = rel(0.85), face = "bold"),
      axis.text = element_text(size = rel(0.70), face = "bold"),
      axis.line = element_line(color = "black"),
      # La légende
      legend.title = element_text(size = rel(0.85), face = "bold"),
      legend.text = element_text(size = rel(0.70), face = "bold"),
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", colour = NA),
      # Les étiquettes dans le cas d'un facetting
      strip.background = element_rect(fill = "#17252D", color = "#17252D"),
      strip.text = element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
    )
}
theme_set(theme_cvis())
```

## Methods

**Spatial analysis** 
Vulnerability indicators for freshwater spawning and rearing life stages represent environmental change to streams, as well as cumulative threats to stream habitats and freshwater residency times. The British Columbia Freshwater Atlas was used as the stream network with spatial delineation to Pacific salmon conservation units based upon conservation unit boundaries. The stream network was limited to accessible streams with channel width and gradients that meet criteria for intrinsic habitat potential for each Pacific salmon species using the [BC Fishpass accessibility model] (https://smnorris.github.io/bcfishpass/02_model_access.html) and [habitat model](https://smnorris.github.io/bcfishpass/03_model_habitat_linear.html). For sockeye salmon, the habitat model also includes criteria for proximity to lakes. This filtering process removed smaller and inaccessible streams while retaining all stream segments with observed occurrences and modelled habitat potential.  

For each indicator of environmental change, a historical baseline period beginning in 1980 was compared to projected conditions averaged over 20-year periods, from 2041 to 2060, 2061 to 2080, and 2081 to 2100. Indicators were also compared across two emissions scenarios or representative concentration pathways (RCPs),  RCP45 and RCP85, representing moderate and high trajectories for continued greenhouse gas emissions. We focus on the mid-century 2041 to 2060 period and RCP45 emissions scenario for comparison of vulnerability scores due to its relevancy for informing planning and decision-making, and lower uncertainty compared to end-of-century projections.  

Environmental change indicators included projected August stream temperatures, rate of change in August stream temperature, percent change in August flows, percent change in flows from November to January, and change in environmental niche model habitat suitability.  Each indicator was calculated as the mean value across all stream segments assigned to the conservation unit, weighted by stream segment length. Temperature indicators (projected and rate of change) were based on a statistical downscaling approach fitted to streams throughout British Columbia (Weller et al. 2024). Hydrologic indicators were based on the Pacific Climate Impacts Consortium gridded hydrological model that was further downscaled to the Fraser basin stream network (CSAS 2024). For the indicator of percent change in August flows, an alternative set of projections using a statistical downscaling approach fitted to hydrometric stations in BC (Ruzzante in prep) was also applied to compare variation in responses (to be further described in an appendix).  

Variation in environmental change indicators for each CU was summarized across global climate model projections and spatial dimensions using 80% quantile ranges. Spatial variation quantiles were weighted by stream length and characterized the range in responses for accessible streams with intrinsic habitat potential using the ensemble mean of climate model projections.  Global climate model quantiles represented the variation of the weighted mean for streams within the CU boundary across all included global climate models (generally 6).  

**Indicator Standardization** 
Raw indicator values were standardized between 0 and 1 to improve comparability, facilitate interpretation, and represent risk using mechanistic relationships and thresholds. A linear function was used for most indicators, with the lowest and highest values scaled to the minimum and maximum mean CU value across all GCMs. An exponential function was used for the projected temperature indicator to represent increasing risk at higher temperatures. Ecological niche model values were not standardized as the model predicted improved habitat favourability for most CUs.  
To represent overall risk across indicators for the spawning and rearing stages, an average of standardized risk scores was calculated.  Other methods for combining risk scores will be further tested. 

## Plots

The following plots show the mean and range of values for environmental change indicators within CU boundaries.For each CU and indicator, the plots show the ensemble mean value for all accessible streams that are modelled as potential spawning or rearing habitat. 

Two types of variation in environmental change are plotted where available:  
**Climate model variation** (thin red bars) - estimated using the minimum and maximum of global climate model means for streams within the CU boundary.
**Spatial variation** (thick grey bars) -estimated using the 95% range of ensemble means for stream segments within the CU boundary. 

The freshwater spawning/rearing indicators are:

```{r}

basins <- st_read(file.path(paths$spatial, "BC_Basins", "BC_Basins_GoogleMapPL.shp"), quiet = TRUE) %>%
  st_cast("POLYGON")
st_crs(basins) <- 4269 
basins <- st_transform(basins, crs = 3005)
Fr_basin <- filter(basins, BASIN == "FRASER") 

indicators <- tribble(
  ~abbrev, ~name,
  "Fav_change", "ENM Change in Favourability",
                      "ct", "Cumulative threats",
                      "Tw8rate", "Rate of change in August Temperature",
                      "Tw8proj", "Projected August Temperature",
                      "lowQpdelta", "Proportional change in August flow (stream model)",
                      "st8pdelta",  "Proportional change in August flow (station model)",
                      "highQpdelta",     "Proportional change in Nov-Jan flow (stream model)")


indicators %>%
  gt()

```



```{r plot function}


# Function to create a lollipop chart
plot_lollipop <- function(data, indicator_pick) {
  # Define suffixes
  stat_suffix <- "wmean"
  sp_suffix <- c("wqlow_sp", "wqhigh_sp")
  gcm_range_suffix <- c("qlow_gcm", "qhigh_gcm")
  
  #take column names that contain prefix with model type
  cols_sub <- names(data)[str_detect(names(data), indicator_pick$abbrev)]
  
  #take names with prefix that also contain stat suffix
  stat_col <- cols_sub[str_detect(cols_sub, paste0(stat_suffix, "$"))]  #must end with wmean
  min_col  <- cols_sub[str_detect(cols_sub,sp_suffix[1])]
  max_col  <- cols_sub[str_detect(cols_sub,sp_suffix[2])]
  min_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[1], collapse = "|"))]
  max_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[2], collapse = "|"))]
  
  id_col <- "CU_NAME"
  sp_col <- "CU_Species"
  
  # Check if mean and ID columns exist
  if (!all(c(id_col, stat_col) %in% names(data))) {
    stop("One or more required columns are missing in the data.")
  }
  
  # Prepare data for plotting
  plot_data <- data %>%
    select(all_of(c(id_col, stat_col, sp_col))) %>%
    rename(mean = !!stat_col, id = !!id_col, sp = !!sp_col)

  #palette_colors <- pal_npg("nrc")(5) # 4 colors from Nature Publishing Group palette

  # Create a named color palette using unique categories
  unique_sp <- unique(plot_data$sp)
  palette_colors <- pal_npg("nrc")(length(unique_sp))
  names(palette_colors) <- unique_sp
  
  plot_data$color <- palette_colors[plot_data$sp]
  plot_data$label <- paste0("<span style='color:", plot_data$color, "'>", plot_data$sp, "</span>")
  
  
# Create a unique ordering of id by sp
ordered_ids <- plot_data %>%
  arrange(sp, id) %>%
  distinct(id) %>%
  pull(id)

# Apply the ordering
plot_data$id <- factor(plot_data$id, levels = ordered_ids)
  
  # Add min and max if available
  if (length(min_col) == 1 && length(max_col) == 1) {
    plot_data <- plot_data %>%
      mutate(
        min = data[[min_col]],
        max = data[[max_col]]
      )
    has_range <- TRUE
  } else {
    has_range <- FALSE
  }
  
  # add GCM range if available
  if (length(min_gcmcol) == 1 && length(max_gcmcol) == 1) {
    plot_data <- plot_data %>%
      mutate(
        mingcm = data[[min_gcmcol]],
        maxgcm = data[[max_gcmcol]]
      )
    has_gcm <- TRUE
  } else {
    has_gcm <- FALSE
  }
  
  # Create the lollipop chart
  p <- ggplot(plot_data, aes(x = id, color = sp)) +
    scale_fill_manual(values = palette_colors) 
  
  if (has_range) {
    p <- p + geom_segment(aes(xend = id, y = min, yend = max), 
                          color = "darkgrey",
                          linewidth = 2.5)
  }
  
  if (has_gcm) {
    p <- p + geom_segment(aes(xend = id, y = mingcm, yend = maxgcm), 
                          color = "darkred",
                            linewidth = 1)
  }
  
   p +   geom_point(aes(y = mean), size = 2.5) +
     labs(
      title = paste(indicator_pick$name),
      x = "CU",
      y = stat_col,
      color = "Species"
    ) +
    coord_flip() +
    theme_minimal() +
    theme(
       axis.text.y =element_text(size=7)
       
     )

}
```


```{r}


spatial_indicator_plot <- function(data,
                                   outline = Fr_basin,
                                   sp_pick = "Chinook", 
                                   indicator_pick, 
                                   palette = "Zissou1") 
{
  
  data_sp <- filter(data, Species == sp_pick)

  #take column names that contain prefix with model type
  cols_sub <- names(data_sp)[str_detect(names(data_sp), indicator_pick$abbrev)]
  #take names with prefix that also contain stat suffix
  stat_col <- cols_sub[str_detect(cols_sub, paste0(stat_suffix, "$"))]
  
  data_sp <- mutate(data_sp, plot_col = .data[[stat_col]]) 
  
 p <- ggplot() +
  geom_sf(data = data_sp, aes(fill = plot_col), alpha = 0.3) +
  scale_fill_gradientn(colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous")) +
  #geom_label(data = cu_boundary_show, aes(label = CUID), size = 2) +
  geom_sf(data = Fr_basin, colour = "black", fill = NA, alpha = 0.3) + 
   labs(subtitle = paste(sp_pick, "-", indicator_pick$name),
        legend = indicator_pick$abbrev)
 
 #if(palette == "Zissou1") p 
 #if(palette == "Viridis") p + scale_fill_viridis_c()
 
 return(p)

}



```




## RCP 45 - 2041 to 2060

```{r}

data <- filter(fwR_all_std, period == "3", RCP == "45")

### Low flow comparison plot
stat_suffix <- "wmean"
gcm_range_suffix <- c("qlow_gcm", "qhigh_gcm")
#take column names that contain prefix with model type
cols_sub <- names(data)[str_detect(names(data), "lowQpdelta|st8pdelta")]

stat_col <- cols_sub[str_detect(cols_sub, paste0(stat_suffix, "$"))]  #must end with wmean
min_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[1], collapse = "|"))]
max_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[2], collapse = "|"))]
  
id_col <- "CU_NAME"
sp_col <- "CU_Species"
  
  # Check if mean and ID columns exist
  if (!all(c(id_col, stat_col) %in% names(data))) {
    stop("One or more required columns are missing in the data.")
  }
  
  # Prepare data for plotting
  plot_data <- data %>%
    select(all_of(c(id_col, stat_col, sp_col, min_gcmcol, max_gcmcol)))
  
  ggplot(plot_data) +
    geom_errorbar(aes(x = lowQpdelta_wmean, y = st8pdelta_wmean, ymin = st8pdelta_qlow_gcm, ymax = st8pdelta_qhigh_gcm), colour = "blue") +
    geom_errorbarh(aes(x = lowQpdelta_wmean, y = st8pdelta_wmean, xmin = lowQpdelta_wmean_qlow_gcm, xmax = lowQpdelta_wmean_qlow_gcm)) +
        geom_point(aes(x=lowQpdelta_wmean, y=st8pdelta_wmean, col = CU_Species), size = 3) +
    labs(subtitle = "statistical station model (st8) vs PCIC stream model",
         x = "PCIC stream model change in August flow",
         y = "Ruzzante station model change in August flow")
  

```

```{r standardized indicators}

#### plotting standardized indicators

# lollipop plot of all rearing indicators
data <- filter(fwR_all_std, period == "3", RCP == "45")

#get standardized indicator value columns
cols_sub <- names(data)[str_starts(names(data), "std")]

stat_suffix <- "wmean"
gcm_range_suffix <- c("qlow_gcm","qhigh_gcm")

#take names with prefix that also contain stat suffix
stat_col <- cols_sub[str_detect(cols_sub, paste0(stat_suffix, "$"))]  #must end with wmean

min_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[1], collapse = "|"))]
max_gcmcol <- cols_sub[str_detect(cols_sub, paste0(gcm_range_suffix[2], collapse = "|"))]

id_col <- "CU_NAME"
sp_col <- "CU_Species"

# Check if mean and ID columns exist
if (!all(c(id_col, stat_col) %in% names(data))) {
  stop("One or more required columns are missing in the data.")
}



# Prepare data for plotting
plot_data <- data %>%
  select(all_of(c(id_col, stat_col, sp_col, min_gcmcol, max_gcmcol))) %>%
  select(-contains(c("st8", "QmonthH"))) %>%
  pivot_longer(cols = starts_with("std"),
               names_to = c("prefix", "indicator", "stat"),
               names_pattern = "^([^_]+)_([^_]+)_(.+)$") %>%
  pivot_wider(names_from = "stat", 
              values_from = "value") 
  
# Create a unique ordering of id by sp
ordered_ids <- plot_data %>%
  arrange(CU_Species, CU_NAME) %>%
  distinct(CU_NAME) %>%
  pull(CU_NAME)

# Apply the ordering
plot_data$CU_NAME <- factor(plot_data$CU_NAME, levels = ordered_ids)
  
# Create a named color palette using unique categories
unique_sp <- unique(plot_data$CU_Species)
palette_colors <- pal_npg("nrc")(length(unique_sp))
names(palette_colors) <- unique_sp

# Create the lollipop chart
p <- ggplot(plot_data, aes(x = CU_NAME)) +
  geom_segment(aes(xend = CU_NAME, y = wmean_qlow_gcm, yend = wmean_qhigh_gcm), 
               color = "darkgrey",
               linewidth = 1) +
  geom_point(aes(y = wmean, color = CU_Species), size = 3) +
  coord_flip() +
  theme_minimal() +
  theme(
       axis.text.y =element_text(size=8),
       axis.text.x =element_text(size=7),
       legend.position = "none"
     ) +
  scale_y_continuous(breaks = c(0,0.5,1), labels = c(0, 0.5, 1)) + 
  facet_grid(.~ indicator) +
  labs(title = "Standardized indicator values for Fraser CUs",
       y = "Standardized value",
       x = NULL)


p

```


```{r period 3 indicator plots}

data <- filter(fwR_all_flat, period == "3", RCP == "45")

cu_boundary_plot <- cu_boundary %>%
  filter(FULL_CU_IN %in% cu_run$FULL_CU_IN) %>%
  left_join(data, by = c("FULL_CU_IN"))
  
for(i in 1:nrow(indicators)) {
  
  p <- plot_lollipop(data, indicators[i,])
  
  p2 <- spatial_indicator_plot(cu_boundary_plot, 
                               sp_pick = "Chinook",
                               indicator_pick = indicators[i,])
  # p3 <- spatial_indicator_plot(cu_boundary_plot, 
  #                              sp_pick = "Coho",
  #                              indicator_pick =indicators[i,])
  # p4 <- spatial_indicator_plot(cu_boundary_plot, 
  #                              sp_pick = "Sockeye-Lake",
  #                              indicator_pick =indicators[i,])

  
  print(p)
  print(p2)
  
}


```

## RCP 45 - 2081 to 2100

```{r period 5}

data <- filter(fwR_all_flat, period == "5", RCP == "45")

for(i in 1:nrow(indicators)) {
  p <- plot_lollipop(data, indicators[i,])
  
  print(p)
}

```

