---
output:
  html_document:
    toc: true
    number-sections: true
    toc_float: true
    toc_collapsed: true
params:
  FULL_CU_IN: "CK-3"
---


```{r custom theme, echo = FALSE, message = F, warning = F}

library(gt)        # nice tables


knitr::opts_chunk$set(message = F, echo = F, warning = F)


```



```{r subset CU}

#params$FULL_CU_IN <- "CK-17"
cu_i <- params$FULL_CU_IN
cu_run_i <- cu_run[cu_run$FULL_CU_IN == params$FULL_CU_IN,]
sp_pick <- cu_run$spp[cu_run$FULL_CU_IN == cu_i] #species abbr
sp_pick_bcfp <- spp_lookup$spp_abr_bcfp[spp_lookup$spp_abr == sp_pick] 


```


---
title: `r cu_run_i$CU_NAME` - Climate Vulnerability Indicator Report 
---




```{r spatial subset, echo = FALSE, message = F, warning = F}


# Stream spatial layers
fwR_cu <- fwR_all[[cu_i]] 

stream_cu_sub <- stream_cu_picks[,colnames(stream_cu_picks) == cu_i]
reaches_ENM_sub <- ENM_cu_picks[,colnames(ENM_cu_picks) == cu_i]

cu_boundary_i <- cu_boundary[cu_boundary$FULL_CU_IN == cu_i,]

#subset nuseds observations
nuseds_cu <- nuseds_Fr[nuseds_Fr$FULL_CU_IN == cu_i,]

fwModels_cu <- fwModels[stream_cu_sub,] %>%
  rename(keep_model_spawning = contains(paste0("model_spawning_", sp_pick_bcfp)),
         keep_model_rearing  = contains(paste0("model_rearing_", sp_pick_bcfp))) %>%
  mutate(model_rs = if_any(starts_with("keep_model"), ~ . == TRUE)) %>%#get boolean for model spawning and rearing 
  mutate(model_rs = factor(model_rs, levels = c(TRUE, FALSE), 
                           labels = c("1-SPAWNING/REARING", "2-ACCESSIBLE"))) %>%
  select(-starts_with(c("model_spawning","model_rearing", "known_rearing", "known_spawning")))

ENM_cu <- reaches_ENM_all[reaches_ENM_sub,]

#read Temperature gauge locations
Tw_stations <- st_read(file.path(paths$climate, "Tw_stations.gdb"), quiet = TRUE)

#subset in CU boundary
temp <- unlist(st_intersects(cu_boundary_i, watershed_flow))
watershed_flow_cu <- watershed_flow[temp,]

#flow stations
stations_flow <- st_read(file.path(paths$climate, "Ruzzante_low_flows", "stations.gpkg"), quiet = TRUE) %>%
  st_transform(3005) 
#subset in cu boundary
temp <- unlist(st_intersects(cu_boundary_i, stations_flow))
stations_cu <- stations_flow[temp,]


## Migration spatial layers
migr_cu <- migr_list[[cu_i]]

migr_flat_cu <- migr_all_flat[migr_all_flat$FULL_CU_IN == cu_i,]


## Timing data
cu_timing_long_i <- cu_timing_long[cu_timing_long$FULL_CU_IN == cu_i, ]
cu_timing_i <- cu_timing_Fr[cu_timing_Fr$FULL_CU_IN == cu_i, ]


#get long data frame of standardized indicator values for the CU
ind_cu <- get_CU_indicators(all_flat_std,
  cu_i,
  indicators_choose = tbl_indicators$abbrev,
  RCP_pick = "45",
  period_pick = "3"
)


```


```{r cu boundary}
### boundary location
p1 <- cu_boundary_highlight(cu_boundary, 
                      cu_pick = cu_i)

p1

```

## Demographic Overview
 
### Status and abundance
```{r demographics summary}

tbl_indicators %>% 
  filter(type == "dem") %>%
  gt()

dem_details <- cu_run_i %>%
  select(FULL_CU_IN, CU_NAME, 
         gen_length, 
         WSP_population_status, Most_Recent_Generational_Average)
         #SEP_avg_annual_releases_actual, SEP_primary_prod_objective)

dem_details %>%
  gt() %>%
  fmt_number(
    columns = c(gen_length, Most_Recent_Generational_Average),
    decimals = 2
  )

pdem <- plot_cu_lolli(ind_cu,
              indicators_choose = c("CUstatus", "CUnmat"))

pdem


```


## Life stage timing

Timing by life history stage is taken from Wilson and Peacock (2025) [https://cdnsciencepub.com/doi/10.1139/cjfas-2024-0213].

Data quality is rated on a scale of 1 to 5 based on number of observations and their recentness, with 1 being the highest quality data.
```{r life stage timing}

tp <- cu_timing_plot(cu_timing_long_i)

tp
```


## Spawning and Rearing

```{r}

tbl_indicators %>% 
  filter(type == "fwR") %>%
  gt()

pfw <- plot_cu_lolli(ind_cu,
              indicators_choose = c("highQpdelta", "lowQpdelta", 
                                    "Tw8rate", "Tw8proj", 
                                    "ct",
                                    "Favchange",
                                    "fwres"))
pfw

```


### Stream accessibility map
```{r data tables}


acc_plot <- stream_accessible_plot(fwModels_cu,
  nuseds_cu,
  cu_boundary_i)

acc_plot


fwR_cu$streams %>%
  select(c(total_length_acc, n_streams, proportion_rear, proportion_spawn, proportion_rs,
           avg_elevation, avg_lat, avg_lon)) %>%
  mutate(
    total_length_acc = round(total_length_acc / 1000, 1),
    proportion_rear = round(proportion_rear, 3),
    proportion_spawn = round(proportion_spawn, 3),
    proportion_rs = round(proportion_rs, 3),
    avg_elevation = round(avg_elevation, 1),
    avg_lat = round(avg_lat, 5),
    avg_lon = round(avg_lon, 5)
  ) %>%
  gt() %>%
  tab_header(
    title = md("**CU Boundary Stream Statistics**")
  ) %>%
  cols_label(
    total_length_acc = "Accessible Stream Length (km)",
    n_streams = "Number of Stream Segments",
    proportion_rear = "Proportion Rearing",
    proportion_spawn = "Proportion Spawning",
    proportion_rs = "Proportion Rearing/Spawning",
    avg_elevation = "Avg. Elevation (m)",
    avg_lat = "Avg. Latitude",
    avg_lon = "Avg. Longitude"
  ) %>%
  tab_spanner(
    label = "Habitat Model",
    columns = c(proportion_rear, proportion_spawn, proportion_rs)
  )


fwR_cu$streams %>%
  select(contains("nuseds")) %>%
  gt() %>%
  tab_header(title = "CU NUSEDS statistics")


# fwR_cu$wpQlow %>%
#   gt() %>%
#   tab_header(title = "August flows - statistical gauge model")

```

### Hydrologic Regime

```{r hydrology}

fwR_cu$streams %>%
  select(contains("prop_")) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  gt() %>%
  tab_header(
    title = md("**Hydrologic Regime Coverage Proportions**")
  ) %>%
  cols_label(
    prop_snow = "Snow-Dominated",
    prop_rain = "Rain-Dominated",
    prop_hybrid = "Hybrid",
    prop_glacial = "Glacial",
    prop_coverage = "Total Coverage"
  ) %>%
  fmt_number(
    columns = where(is.numeric),
    decimals = 3
  )


# hydrologic gauge stations and regime plot
regime_plot <- cu_hydrologic_regime(cu_boundary_i,
                                 watershed_flow_cu,
                                 stations_cu,
                                 fwModels_cu)

regime_plot


# fwR_cu$fwQlow %>%
#     select(c(GCM, RCP, period), contains("wmean"), contains("wqlow"), 
#          contains("wqhigh"), -contains("sd"), -contains("CV")) %>%
#   mutate(
#     across(where(is.numeric), ~ round(.x, 3))
#   ) %>%
#   gt() %>%
#   tab_header(
#     title = md("**August Flows – PCIC Stream Model**")
#   ) %>%
#   cols_label(
#     GCM = "GCM",
#     RCP = "RCP Scenario",
#     period = "Time Period",
#     Qproj_wmean = "Projected Flow Mean (m³/s)",
#     QMADhist_wmean = "Historical MAD Mean",
#     Qpdelta_wmean = "Δ Flow Mean",
#     Qproj_wmean_qlow_gcm = "GCM Q1 (Flow)",
#     Qproj_wmean_qhigh_gcm = "GCM Q3 (Flow)",
#     Qpdelta_wmean_qlow_gcm = "GCM Q1 (Δ Flow)",
#     Qpdelta_wmean_qhigh_gcm = "GCM Q3 (Δ Flow)",
#     Qpdelta_wqlow_sp = "Δ Flow Q1 (Spatial)",
#     Qpdelta_wqhigh_sp = "Δ Flow Q3 (Spatial)"
#   ) %>%
#   tab_spanner(
#     label = "Projected Flow (m³/s)",
#     columns = c(Qproj_wmean, Qproj_wmean_qlow_gcm, Qproj_wmean_qhigh_gcm)
#   ) %>%
#   tab_spanner(
#     label = "Change in Flow (Δ)",
#     columns = c(Qpdelta_wmean, Qpdelta_wmean_qlow_gcm, Qpdelta_wmean_qhigh_gcm, Qpdelta_wqlow_sp, Qpdelta_wqhigh_sp)
#   ) %>%
#   fmt_number(
#     columns = where(is.numeric),
#     decimals = 3
#   )

```


### Stream flow indicators
```{r flow indicator plots}
Q8_plot <- stream_indicator_plot(fwModels_cu, 
                                 cu_boundary_i,
                                variable = "Qpdelta_flow_8_45_3", 
                                plot_title = "Change in August flow - 2041-2060",
                                unit_label = "Proportion change",
                                scico_palette = "lapaz",
                                palette_direction = -1)

Q8_plot


QNDJ_plot <- stream_indicator_plot(fwModels_cu, 
                                 cu_boundary_i,
                                variable = "Qpdelta_flow_18_45_3", 
                                plot_title = "Change in Nov_Jan flow - 2041-2060",
                                unit_label = "Proportion change",
                                scico_palette = "devon",
                                palette_direction = -1)
  

QNDJ_plot




# fwR_cu$fwQhigh %>%
#     select(c(GCM, RCP, period), contains("wmean"), contains("wqlow"), 
#          contains("wqhigh"), -contains("sd"), -contains("CV")) %>%
#   gt() %>%
#   tab_header(
#     title = md("**Nov-Jan Flows – PCIC Stream Model**")
#   ) %>%
#   cols_label(
#     GCM = "GCM",
#     RCP = "RCP Scenario",
#     period = "Time Period",
#     Qproj_wmean = "Projected Flow Mean (m³/s)",
#     QMADhist_wmean = "Historical MAD Mean",
#     Qpdelta_wmean = "Δ Flow Mean",
#     Qproj_wmean_qlow_gcm = "GCM Q1 (Flow)",
#     Qproj_wmean_qhigh_gcm = "GCM Q3 (Flow)",
#     Qpdelta_wmean_qlow_gcm = "GCM Q1 (Δ Flow)",
#     Qpdelta_wmean_qhigh_gcm = "GCM Q3 (Δ Flow)",
#     Qpdelta_wqlow_sp = "Δ Flow Q1 (Spatial)",
#     Qpdelta_wqhigh_sp = "Δ Flow Q3 (Spatial)"
#   ) %>%
#   tab_spanner(
#     label = "Projected Flow (m³/s)",
#     columns = c(Qproj_wmean, Qproj_wmean_qlow_gcm, Qproj_wmean_qhigh_gcm)
#   ) %>%
#   tab_spanner(
#     label = "Change in Flow (Δ)",
#     columns = c(Qpdelta_wmean, Qpdelta_wmean_qlow_gcm, Qpdelta_wmean_qhigh_gcm, Qpdelta_wqlow_sp, Qpdelta_wqhigh_sp)
#   ) %>%
#   fmt_number(
#     columns = where(is.numeric),
#     decimals = 3
#   )




```




### Stream temperature indicators

```{r tempreature}


#temperature plot
t_plot <- stream_indicator_plot(fwModels_cu, 
                                cu_boundary_i, 
                                Tw_stations,
                                variable = "Tw8_9_45_3", 
                                temp_stations = TRUE,
                                plot_title = "August Mean Temperature - 2041-2060",
                                unit_label = "Degrees C",
                                scico_palette = "roma",
                                palette_direction = -1)


tRate_plot <- stream_indicator_plot(fwModels_cu, 
                                cu_boundary_i, 
                                Tw_stations,
                                variable = "delta_Tw8_9_45_3", 
                                temp_stations = TRUE,
                                plot_title = "Rate of Change in August Mean Temperature - 1981-2000 to 2041-2060",
                                unit_label = "Degrees C",
                                scico_palette = "roma",
                                palette_direction = -1)

t_plot
tRate_plot


# fwR_cu$fwT %>%
#   select(c(GCM, RCP, period), contains("wmean"), contains("wqlow"), 
#          contains("wqhigh"), -contains("sd"), -contains("CV")) %>%
#   mutate(
#     across(where(is.numeric), ~ round(.x, 3))
#   ) %>%
#   gt() %>%
#   tab_header(
#     title = md("**Projected Stream Temperature Statistics**")
#   ) %>%
#   cols_label(
#     GCM = "GCM",
#     RCP = "RCP Scenario",
#     period = "Time Period",
#     Tw8proj_wmean = "Projected Temp Mean (°C)",
#     Tw8proj_wqlow_sp = "Projected Temp Q1",
#     Tw8proj_wqhigh_sp = "Projected Temp Q3",
#     Tw8rate_wmean = "Temp Change Rate Mean (°C/yr)",
#     Tw8rate_wqlow_sp = "Temp Change Rate Q1",
#     Tw8rate_wqhigh_sp = "Temp Change Rate Q3",
#     Tw8proj_wmean_qlow_gcm = "GCM Q1 (Mean Temp)",
#     Tw8proj_wmean_qhigh_gcm = "GCM Q3 (Mean Temp)",
#     Tw8rate_wmean_qlow_gcm = "GCM Q1 (Rate)",
#     Tw8rate_wmean_qhigh_gcm = "GCM Q3 (Rate)",
#   ) %>%
#   tab_spanner(
#     label = "Projected Temperature (°C)",
#     columns = c(Tw8proj_wmean, Tw8proj_wqlow_sp, Tw8proj_wqhigh_sp)
#   ) %>%
#   tab_spanner(
#     label = "Temperature Change Rate (°C/decade)",
#     columns = c(Tw8rate_wmean, Tw8rate_wqlow_sp, Tw8rate_wqhigh_sp)
#   ) %>%
#   tab_spanner(
#     label = "GCM Variation - Projected Temp",
#     columns = c( Tw8proj_wmean_qlow_gcm, Tw8proj_wmean_qhigh_gcm)
#   ) %>%
#   tab_spanner(
#     label = "GCM Variation - Temp Rate",
#     columns = c(Tw8rate_wmean_qlow_gcm, Tw8rate_wmean_qhigh_gcm)
#   ) %>%
#   fmt_number(
#     columns = where(is.numeric),
#     decimals = 3
#   )




```

### Ecological Niche Model


```{r ENMs}

# fwR_cu$ENM %>%
#   mutate(
#     total_length = round(total_length/1000, 1),
#     across(starts_with("Fav"), round, 3)
#   ) %>%
#   gt() %>%
#   tab_header(
#     title = md("**ENM Favourability by RCP Scenario and Period**")
#   ) %>%
#   cols_label(
#     RCP = "RCP Scenario",
#     period = "Time Period",
#     n_streams = "Number of Streams",
#     total_length = "Total Length (km)",
#     Fav_wmean = "Favourability Mean",
#     Fav_wsd_sp = "Favourability SD",
#     Fav_wqlow_sp = "Favourability 10% Q",
#     Fav_wqhigh_sp = "Favourability 90% Q",
#     Fav_change_wmean = "Δ Favourability Mean",
#     Fav_change_wsd_sp = "Δ Favourability SD",
#     Fav_change_wqlow_sp = "Δ Favourability 10% Q",
#     Fav_change_wqhigh_sp = "Δ Favourability 90 %Q"
#   ) %>%
#   tab_spanner(
#     label = "Favourability Stats",
#     columns = c(Fav_wmean, Fav_wsd_sp, Fav_wqlow_sp, Fav_wqhigh_sp)
#   ) %>%
#   tab_spanner(
#     label = "Change in Favourability",
#     columns = c(Fav_change_wmean, Fav_change_wsd_sp, Fav_change_wqlow_sp, Fav_change_wqhigh_sp)
#   )


ENM_col <- names(select(ENM_cu, contains("Fav_f.9_45_3")))
ENM_col <- ENM_col[str_detect(ENM_col, sp_pick)]

ENM_values <- select(as.tibble(ENM_cu), !!ENM_col)

if(sum(!is.na(ENM_values) > 0)) { 
ENM_plot <- stream_indicator_plot(ENM_cu, 
                                  cu_boundary_i,
                                  variable = ENM_col,
                                  plot_title = "ENM favourability, 2041-2060",
                                  histogram_fill = "model_habitat_salmon",
                                  scico_palette = "berlin",
                                  palette_direction = -1)

ENM_plot
}

ENM_col <- names(select(ENM_cu, contains("Fav_change_9_45_5")))
ENM_col <- ENM_col[str_detect(ENM_col, sp_pick)]

ENM_values <- select(as.tibble(ENM_cu), !!ENM_col)

if(sum(!is.na(ENM_values) > 0)) { 
ENM_diff_plot <- stream_indicator_plot(ENM_cu, 
                                       cu_boundary_i,
                                  variable = ENM_col,
                                  plot_title = "ENM favourability change, 2081-2100",
                                  histogram_fill = "model_habitat_salmon",
                                  scico_palette = "berlin",
                                  palette_direction = -1)

ENM_diff_plot
}

```



### Cumulative threats

```{r Cumulative Threats}

# fwR_cu$CT %>%
#   select(-c(ct_n, ct_total_length)) %>%
#   mutate(
#     ct_wmean = round(ct_wmean, 3),
#     ct_wsd_sp = round(ct_wsd_sp, 3),
#     ct_wqlow_sp = round(ct_wqlow_sp, 3),
#     ct_wqhigh_sp = round(ct_wqhigh_sp, 3)
#   ) %>%
#   gt() %>%
#   tab_header(
#     title = md("**Cumulative Threats Summary**")
#   ) %>%
#   cols_label(
#     CT = "Threat Type",
#     ct_wmean = "Weighted Mean",
#     ct_wsd_sp = "Weighted SD",
#     ct_wqlow_sp = "Weighted 10% Q",
#     ct_wqhigh_sp = "Weighted 90% Q"
#   ) %>%
#   tab_spanner(
#     label = "Weighted Statistics",
#     columns = c(ct_wmean, ct_wsd_sp, ct_wqlow_sp, ct_wqhigh_sp)
#   ) %>%
#   fmt_number(
#     columns = c(ct_wmean, ct_wsd_sp, ct_wqlow_sp, ct_wqhigh_sp),
#     decimals = 3
#   )


ct_plot <- stream_indicator_plot(fwModels_cu, 
                                 cu_boundary_i,
                                variable = "CT_anad", 
                                plot_title = "Cumulative stressors",
                                unit_label = "Cumulative Threat Score",
                                scico_palette = "lajolla",
                                palette_direction = -1)


ct_plot

```


## Upstream Migration

```{r Upstream migr}

tbl_indicators %>% 
  filter(type == "migr") %>%
  gt()

pm <- plot_cu_lolli(ind_cu,
              indicators_choose = c("migrT",
                                    "migrQ",
                                    "migrA21",
                                    "migrdist"))
pm
```


### Migration path

```{r MIgration path}


mplot <- migration_path_plot(migr_cu,
  nuseds_cu,
  cu_boundary_i
)

mplot
  

```


### Migration timing and temperature

```{r Migration temp}


timeplot <- migr_timing_plot(migrT_rcps,
                             cu_i,
                             cu_timing_i,
                             rcp = "45",
                             period_choose = c("1981-2010", "2041-2060"))

timeplot

```


## Nearshore Marine

```{r Marine}

### subset indicator values
cu_mar <- mar_all_flat[mar_all_flat$FULL_CU_IN == cu_i, ]

start_month <- month.name[unique(cu_mar$ns_timing_start)]
  end_month <- month.name[unique(cu_mar$ns_timing_end)]

MAZ_pick <- unique(cu_mar$MAZ)


tbl_indicators %>% 
  filter(type == "mar") %>%
  gt()

pmar <- plot_cu_lolli(ind_cu,
              indicators_choose = c("SSTproj",
                                    "CI"))
pmar

```

### SST

```{r SST}

# subset SST time and locations spatial object
SST_cu_sp <- subset_and_mean_sst(sf_data = filter(SST_grid, MAZ_Acrony == MAZ_pick),
  timing_df = cu_mar,
  RCP_pick = "45")

p <- marine_indicator_plot(SST_cu_sp,
  MAZ = filter(MAZ, MAZ_Acrony == "GStr"),
  plot_title = paste0("Projected SST, RCP 45, ", start_month, "-", end_month))

p


```

